### 2020-04-06 Report/draft1.tex
Working video of greeks in static frame

%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS

\documentclass[11pt, a4paper,twocolumn]{article} % 10pt font size (11 and 12 also possible), A4 paper (letterpaper for US letter) and two column layout (remove for one column) Use additional titlepage argument to generate this
%\documentclass[12pt, a4paper,twocolumn,titlepage]{article}

\input{structure.tex} % Specifies the document structure and loads requires packages
\graphicspath{{"/Users/Kit/OneDrive/Documents/Computing/Trojan Asteroids/Report/Figures"}}

%	ARTICLE INFORMATION

\title{Modelling the Trojan Asteroids}
%\subtitle{Blue Organic Light Emitting Diodes using Thermally Activated Delayed Fluorescence}

%\author{
	%\authorstyle{Christopher Gallagher}}
\author{\authorstyle{Christopher Gallagher} 
	\institution{University of Cambridge}}
% Example of a one line author/institution relationship
%\author{\newauthor{John Marston} \newinstitution{Universidad Nacional Autónoma de México, Mexico City, Mexico}}

\date{\today} % Add a date here if you would like one to appear underneath the title block, use \today for the current date, leave empty for no date
\usepackage[english]{babel}
\usepackage{tabularx}
%\usepackage[backend=biber,doi=false]{biblatex}
\addbibresource{references.bib}
%\AtBeginBibliography{\small}
%----------------------------------------------------------------------------------------
\begin{document}

\maketitle % Print the title

\thispagestyle{firstpage} % Apply the page style for the first page (no headers and footers)

%	ABSTRACT

\lettrineabstract{Trojan asteroids are v important. I'm going to make that paragraph a little longer to fill out the space in case this is forming some kind of error. i really hope this is long enough.}

%	ARTICLE CONTENTS

\section{History}
This is a test with bib \ref{Ravindran2017}. Alternatively we consider \cite{Nakamura2008}.Nakamura2008

\printbibliography

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Wenneker Article
% LaTeX Template
% Version 2.0 (28/2/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Authors:
% Vel (vel@LaTeXTemplates.com)
% Frits Wenneker
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-06 Reference Scripts/Video Tests/video_test.py

@@ -1,0 +1,39 @@
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.animation import FuncAnimation
+
+
+plt.style.use("seaborn-pastel")
+
+
+# We simply create a figure window with a single axis in the figure.
+# Then we create our empty line object which is essentially the one to be modified in the animation.
+# The line object will be populated with data later.
+fig = plt.figure()
+ax = plt.axes(xlim=(0, 4), ylim=(-2, 2))
+(line,) = ax.plot([], [], lw=3)
+
+
+def init():
+    """ The init function initializes the data and also sets the axis limits."""
+    line.set_data([], [])
+    return (line,)
+
+
+def animate(i):
+    """Takes in the frame number(i) as the parameter and creates a function dependant on i 
+    
+    Returns a tuple of the plot objects which have been modified which 
+    tells the animation framework what parts of the plot should be animated.
+    """
+    x = np.linspace(0, 4, 1000)
+    y = np.sin(2 * np.pi * (x - 0.01 * i))
+    line.set_data(x, y)
+    return (line,)
+
+
+anim = FuncAnimation(fig, animate, init_func=init, frames=1000, interval=20, blit=True)
+# create actual animation object, blit parameter ensures that only changed parts of plot are re-drawn
+
+anim.save("sine_wave.gif", writer="imagemagick")

### 2020-04-06 Reference Scripts/Video Tests/video_test2.py

@@ -1,0 +1,28 @@
+import matplotlib.pyplot as plt
+from matplotlib.animation import FuncAnimation
+
+Acc_11 = [1, 2, 3, 4, 6, 7]
+Acc_12 = [2, 2, 2, 2, 2, 2]
+
+# Scatter plot
+fig = plt.figure(figsize=(5, 5))
+axes = fig.add_subplot(111)
+axes.set_xlim(min(Acc_11), max(Acc_11))
+axes.set_ylim(min(Acc_12), max(Acc_12))
+
+(point,) = axes.plot([Acc_11[0]], [Acc_12[0]], "go")
+
+
+def ani(coords):
+    point.set_data([coords[0]], [coords[1]])
+    return point
+
+
+def frames():
+    for acc_11_pos, acc_12_pos in zip(Acc_11, Acc_12):
+        yield acc_11_pos, acc_12_pos
+
+
+ani = FuncAnimation(fig, ani, frames=frames, interval=1000)
+
+plt.show()

### 2020-04-06 Reference Scripts/Video Tests/video_test3.py

@@ -1,0 +1,56 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.animation as animation
+from math import sin, cos
+
+
+def data_gen():
+    gen_list = ([cos(t), sin(t)] for t in np.arange(0, 1000, 0.1))
+    return gen_list
+
+
+print(np.arange(0, 1000, 0.1))
+print(data_gen())
+print([cos(t), sin(t)] for t in np.arange(0, 1000, 0.1))
+
+
+def init():
+    ax.set_ylim(-6, 6)
+    ax.set_xlim(-6, 6)
+    return point
+
+
+fig, ax = plt.subplots()
+(point,) = ax.plot([0], [0], "go")
+point.set_data(0, 0)
+
+
+def run(data):
+
+    t, y = data
+    # xmin, xmax = ax.get_xlim()
+    # ymin, ymax = ax.get_ylim()
+
+    # if t >= xmax:
+    #     ax.set_xlim(xmin, t + 1)
+    #     ax.figure.canvas.draw()
+
+    # if t <= xmin:
+    #     ax.set_xlim(t - 1, xmax)
+    #     ax.figure.canvas.draw()
+
+    # if y >= ymax:
+    #     ax.set_ylim(ymin, y + 1)
+    #     ax.figure.canvas.draw()
+
+    # if y <= ymin:
+    #     ax.set_ylim(y - 1, ymax)
+    #     ax.figure.canvas.draw()
+
+    point.set_data(t, y)
+
+    return point
+
+
+ani = animation.FuncAnimation(fig, run, data_gen, init_func=init, interval=10)
+plt.show()

### 2020-04-06 orbits.py

@@ -20,6 +20,7 @@ greek_v = omega * (greek_rad)
 
 lagrange = (planet_rad - R / 2, R * math.sqrt(3) / 2, 0)
 # are these necessary as well as initial conditions?
+# decide later which are most useful to import etc
 
 # Initial conditions for Greeks, given as equillibrium state but can be changed in XXXXX
 cos, sin = np.cos(greek_theta), np.sin(greek_theta)
@@ -58,9 +59,11 @@ def rot_derivatives(t, y):
     solar_dr3 = np.linalg.norm(solar_pos(0) - position) ** 3  # should this be zero?
     planet_dr3 = np.linalg.norm(planet_pos(0) - position) ** 3
 
-    virtual_force_x = position[0] * omega ** 2 + 2 * omega * velocity[1]
-    virtual_force_y = position[1] * omega ** 2 - 2 * omega * velocity[0]
-    virtual_force_z = 0  # change this!!
+    virtual_force = (
+        position[0] * omega ** 2 + 2 * omega * velocity[1],
+        position[1] * omega ** 2 - 2 * omega * velocity[0],
+        0,
+    )
 
     acceleration = (
         -G
@@ -68,11 +71,11 @@ def rot_derivatives(t, y):
             M_S * (position[0] - solar_pos(0)[0]) / solar_dr3
             + M_P * (position[0] - planet_pos(0)[0]) / planet_dr3
         )
-        + virtual_force_x,
+        + virtual_force[0],
         -G * (M_S * position[1] / solar_dr3 + M_P * position[1] / planet_dr3)
-        + virtual_force_y,
+        + virtual_force[1],
         -G * (M_S * position[2] / solar_dr3 + M_P * position[2] / planet_dr3)
-        + virtual_force_z,
+        + virtual_force[2],
     )
 
     return np.concatenate((velocity, acceleration))

### 2020-04-06 Reference Scripts/Video Tests/video_test4_loop.py
Pre-branch for loop plotting implementation

@@ -1,0 +1,34 @@
+import matplotlib
+
+
+from matplotlib import pyplot as plt
+from matplotlib import animation
+
+fig = plt.figure()
+
+ax = plt.axes(xlim=(0, 2), ylim=(0, 100))
+
+N = 4
+points = [plt.plot() for _ in range(N)]  # points to animate
+
+
+def init():
+    # init points
+    for point in points:
+        point.set_data()
+    return points
+
+
+def animate(i):
+    # animate points
+    for j, point in enumerate(points):
+        point.set_data(10 * j, i)
+
+    return points
+
+
+anim = animation.FuncAnimation(
+    fig, animate, init_func=init, frames=100, interval=20, blit=True
+)
+
+plt.show()

### 2020-04-06 Reference Scripts/Video Tests/video_test4_loop_points.py

@@ -1,0 +1,39 @@
+from matplotlib import pyplot as plt
+from matplotlib import animation
+import numpy as np
+
+fig = plt.figure()
+
+ax = plt.axes(xlim=(-10, 10), ylim=(0, 100))
+
+N = 4
+points2 = ax.plot(*([[], []] * N), marker="o")  # old version
+
+
+points = []
+for i in range(N):
+    point = ax.plot([], [], label=i, marker="o")
+    points.append(point)
+points = np.ndarray.tolist(np.squeeze(points))
+ax.legend(loc=4)
+
+
+def init():
+    for line in points:
+        line.set_data([], [])
+    return points
+
+
+def animate(i):
+    points[0].set_data([0], [i])
+    points[1].set_data([[1], [i + 1]])
+    points[2].set_data([[2], [i + 2]])
+    points[3].set_data([[3], [i + 3]])
+    return points
+
+
+anim = animation.FuncAnimation(
+    fig, animate, init_func=init, frames=100, interval=20, blit=True
+)
+
+plt.show()

### 2020-04-06 Reference Scripts/backupfile.py

@@ -141,3 +141,63 @@ def stationary_frame(y0_stat=(rcos, rsin, 0, -vsin, vcos, 0)):
         y0=y0_stat,
         t_eval=time_span,
     )
+
+
+# CURRENT VERSION OF STAT_FRAME_VIDEO BEFORE IMPLEMEMNTING LOOP 06/04/2020 15;22
+
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.animation as animation
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import lagrange, omega, time_span  # Derived constants
+
+# Set up orbits for each body
+orbit_greeks = orbits.stationary_frame(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+orbit_trojans = orbits.stationary_frame(
+    (lagrange[0], -lagrange[1], 0, -omega * lagrange[1], -omega * lagrange[0], 0)
+)
+orbit_list = [orbit_greeks, orbit_trojans]
+
+
+def data_gen(orbit_data):
+    gen_list = (
+        [orbit_data.y[0, t], orbit_data.y[1, t]]
+        for t in np.arange(
+            0, ORBIT_NUM * PRECISION, 100
+        )  # step size of 100 pulls out each 100th point to speed up animation
+    )
+    return gen_list
+
+
+def init_g():
+    return point_g
+
+
+def run_g(data):
+    t, y = data
+    point_g.set_data(t, y)
+    return point_g
+
+
+fig = plt.figure()
+ax = plt.axes(xlim=(-8, 8), ylim=(-8, 8))
+ax.set_aspect(aspect=1)
+
+(point_g,) = ax.plot([0], [0], "go", label="Greeks")
+point_g.set_data(0, 0)
+
+ani = animation.FuncAnimation(
+    fig, run_g, data_gen(orbit_greeks), init_func=init_g, interval=1
+)
+plt.title("Asteroid orbit in static frame")
+plt.xlabel("X distance/ AU")
+plt.ylabel("Y distance/ AU")
+plt.legend()
+plt.show()
+
+# ani.save("asteroid_orbit.gif", writer="imagemagick")

### 2020-04-06 constants.py

@@ -16,3 +16,4 @@ planet_rad = R * M_S / (M_S + M_P)
 period = math.sqrt(R ** 3 / (M_S + M_P))
 omega = 2 * np.pi / period  # angular velocity of frame
 time_span = np.linspace(0, ORBIT_NUM * period, int(ORBIT_NUM * PRECISION))
+lagrange = (planet_rad - R / 2, R * math.sqrt(3) / 2, 0)

### 2020-04-06 stat_frame_video.py

@@ -1,0 +1,56 @@
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.animation as animation
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import lagrange, omega, time_span  # Derived constants
+
+# Set up orbits for each body
+orbit_greeks = orbits.stationary_frame(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+orbit_trojans = orbits.stationary_frame(
+    (lagrange[0], -lagrange[1], 0, -omega * lagrange[1], -omega * lagrange[0], 0)
+)
+orbit_list = [orbit_greeks, orbit_trojans]
+
+
+def data_gen(orbit_data):
+    gen_list = (
+        [orbit_data.y[0, t], orbit_data.y[1, t]]
+        for t in np.arange(
+            0, ORBIT_NUM * PRECISION, 100
+        )  # step size of 100 pulls out each 100th point to speed up animation
+    )
+    return gen_list
+
+
+def init_g():
+    return point_g
+
+
+def run_g(data):
+    t, y = data
+    point_g.set_data(t, y)
+    return point_g
+
+
+fig = plt.figure()
+ax = plt.axes(xlim=(-8, 8), ylim=(-8, 8))
+ax.set_aspect(aspect=1)
+
+(point_g,) = ax.plot([0], [0], "go", label="Greeks")
+point_g.set_data(0, 0)
+
+ani = animation.FuncAnimation(
+    fig, run_g, data_gen(orbit_greeks), init_func=init_g, interval=1
+)
+plt.title("Asteroid orbit in static frame")
+plt.xlabel("X distance/ AU")
+plt.ylabel("Y distance/ AU")
+plt.legend()
+plt.show()
+
+# ani.save("asteroid_orbit.gif", writer="imagemagick")

### 2020-04-06 constants.py
Working animation with all bodies, asteroid drift

@@ -7,8 +7,8 @@ M_S = 1  # solar mass
 M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
-PRECISION = 5000  # evaluation points per orbit
-ORBIT_NUM = 10  # number of orbits
+PRECISION = 1000  # evaluation points per orbit
+ORBIT_NUM = 5  # number of orbits
 
 # DERIVED CONSTANTS
 solar_rad = R * M_P / (M_S + M_P)  # distance from origin to sun in CoM frame

### 2020-04-06 stat_frame_video.py

@@ -7,50 +7,89 @@ import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import lagrange, omega, time_span  # Derived constants
 
+SAMPLE = int(500)  # Sample every n points
+
 # Set up orbits for each body
+orbit_sun = orbits.solar_pos(time_span)
+orbit_jupiter = orbits.planet_pos(time_span)
 orbit_greeks = orbits.stationary_frame(
     (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
 )
 orbit_trojans = orbits.stationary_frame(
-    (lagrange[0], -lagrange[1], 0, -omega * lagrange[1], -omega * lagrange[0], 0)
+    (lagrange[0], -lagrange[1], 0, omega * lagrange[1], omega * lagrange[0], 0)
 )
-orbit_list = [orbit_greeks, orbit_trojans]
+orbit_list = [orbit_sun, orbit_jupiter, orbit_greeks, orbit_trojans]
+bodies = ["Sun", "Jupiter", "Greeks", "Trojans"]
+N = len(bodies)
+
+orbit_data = np.zeros(
+    (len(orbit_trojans.t[0::SAMPLE]), 2, int(N))
+)  # select time coordinate, spatial coordinate (x/y), and index of body in list 'bodies'
 
 
-def data_gen(orbit_data):
-    gen_list = (
-        [orbit_data.y[0, t], orbit_data.y[1, t]]
-        for t in np.arange(
-            0, ORBIT_NUM * PRECISION, 100
-        )  # step size of 100 pulls out each 100th point to speed up animation
-    )
-    return gen_list
+for i in range(N):
+    if type(orbit_list[i]) is np.ndarray:  # Different format of orbit results
+        orbit_data[:, 0, i] = orbit_list[i][0][0::SAMPLE]
+        orbit_data[:, 1, i] = orbit_list[i][1][0::SAMPLE]
+    else:
+        orbit_data[:, 0, i] = orbit_list[i].y[0, 0::SAMPLE]
+        orbit_data[:, 1, i] = orbit_list[i].y[1, 0::SAMPLE]
 
 
-def init_g():
-    return point_g
-
-
-def run_g(data):
-    t, y = data
-    point_g.set_data(t, y)
-    return point_g
-
-
-fig = plt.figure()
[…]

### 2020-04-06 stat_frame_video.py
Text reader added

@@ -7,7 +7,7 @@ import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import lagrange, omega, time_span  # Derived constants
 
-SAMPLE = int(500)  # Sample every n points
+SAMPLE = int(5)  # Sample every n points
 
 # Set up orbits for each body
 orbit_sun = orbits.solar_pos(time_span)
@@ -39,8 +39,9 @@ for i in range(N):
 fig = plt.figure(figsize=(6, 6))
 ax = plt.axes(xlim=(-8, 8), ylim=(-8, 8))
 ax.set_aspect(aspect=1)  # So that circular orbits will not be distorted
+time_marker = ax.text(0.02, 0.05, "", transform=ax.transAxes)
 
-markers = [12, 8, 4, 4]
+markers = [12, 8, 2, 2]
 colours = ["yellow", "red", "blue", "green"]
 points = []
 for i in range(N):
@@ -69,16 +70,20 @@ def animate(t):
     """Takes in the frame point in time(t) as the parameter and creates a function dependant on t for each body """
     for i in range(N):
         points[i].set_data([orbit_data[t, 0, i], orbit_data[t, 1, i]])
-    return points
+    time_marker.set_text(str("{:.1f}".format(orbit_trojans.t[t] * SAMPLE)) + " years")
+    # return (points, time_marker)
+    return points + [time_marker]
 
 
 anim = animation.FuncAnimation(
     fig,
     animate,
     init_func=init,
-    frames=ORBIT_NUM * int(PRECISION / SAMPLE),
-    interval=1,
+    # frames=ORBIT_NUM * int(PRECISION / SAMPLE),
+    frames=5 * int(PRECISION / SAMPLE),
+    interval=0.75,
     blit=True,
+    repeat=False,
 )
 
 plt.title("Asteroid orbit in static frame")
@@ -90,6 +95,6 @@ plt.ylabel("Y distance/ AU")
 writer = animation.FFMpegWriter(
     fps=60, metadata=dict(artist="Kit Gallagher"), bitrate=1800
 )
-# anim.save("movie.mp4", writer=writer)
+anim.save("movie.mp4", writer=writer)
 
 plt.show()

### 2020-04-07 constants.py
Working video with lagrange point plotted Jupiter and asteroid distance plotted

@@ -8,7 +8,7 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 PRECISION = 1000  # evaluation points per orbit
-ORBIT_NUM = 5  # number of orbits
+ORBIT_NUM = 30  # number of orbits
 
 # DERIVED CONSTANTS
 solar_rad = R * M_P / (M_S + M_P)  # distance from origin to sun in CoM frame

### 2020-04-07 main_file.py

@@ -77,3 +77,19 @@ plt.plot(orbit_sol2.y[0, :], orbit_sol2.y[1, :], label="Greeks")
 plt.title("Stationary Frame")
 plt.legend()
 plt.show()
+
+
+plt.plot(
+    time_span,
+    np.linalg.norm(
+        [
+            (orbit_sol2.y[0, :] - orbits.planet_pos(time_span)[0]),
+            (orbit_sol2.y[1, :] - orbits.planet_pos(time_span)[1]),
+        ],
+        axis=0,
+    ),
+)
+plt.title("Separtion between Jupiter and Asteroid in rotating frame")
+plt.ylabel("Separation / AU")
+plt.xlabel("Time / years")
+plt.show()

### 2020-04-07 orbits.py

@@ -45,6 +45,19 @@ def planet_pos(t=0):
     return np.array([planet_rad * np.cos(omega * t), planet_rad * np.sin(omega * t), 0])
 
 
+print(planet_pos(time_span)[0])
+
+
+def lagrange_pos(t=0):
+    return np.array(
+        [
+            greek_rad * np.cos(omega * t + greek_theta),
+            greek_rad * np.sin(omega * t + greek_theta),
+            0,
+        ]
+    )
+
+
 # assuming massive bodies do not deviate from xy plane
 
 

### 2020-04-07 stat_frame_video.py

@@ -7,19 +7,20 @@ import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import lagrange, omega, time_span  # Derived constants
 
-SAMPLE = int(5)  # Sample every n points
+SAMPLE = int(1)  # Sample every n points
 
 # Set up orbits for each body
 orbit_sun = orbits.solar_pos(time_span)
 orbit_jupiter = orbits.planet_pos(time_span)
+orbit_lagrange = orbits.lagrange_pos(time_span)
 orbit_greeks = orbits.stationary_frame(
     (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
 )
 orbit_trojans = orbits.stationary_frame(
     (lagrange[0], -lagrange[1], 0, omega * lagrange[1], omega * lagrange[0], 0)
 )
-orbit_list = [orbit_sun, orbit_jupiter, orbit_greeks, orbit_trojans]
-bodies = ["Sun", "Jupiter", "Greeks", "Trojans"]
+orbit_list = [orbit_sun, orbit_jupiter, orbit_lagrange, orbit_greeks, orbit_trojans]
+bodies = ["Sun", "Jupiter", "Lagrange", "Greeks", "Trojans"]
 N = len(bodies)
 
 orbit_data = np.zeros(
@@ -41,15 +42,16 @@ ax = plt.axes(xlim=(-8, 8), ylim=(-8, 8))
 ax.set_aspect(aspect=1)  # So that circular orbits will not be distorted
 time_marker = ax.text(0.02, 0.05, "", transform=ax.transAxes)
 
-markers = [12, 8, 2, 2]
-colours = ["yellow", "red", "blue", "green"]
+markers = [12, 8, 5, 2, 2]
+colours = ["yellow", "red", "black", "blue", "green"]
+shapes = ["o", "o", "x", "o", "o"]
 points = []
 for i in range(N):
     point = ax.plot(
         [],
         [],
         label=bodies[i],
-        marker="o",
+        marker=shapes[i],
         linestyle="none",
         markersize=markers[i],
         color=colours[i],
@@ -81,7 +83,7 @@ anim = animation.FuncAnimation(
     init_func=init,
     # frames=ORBIT_NUM * int(PRECISION / SAMPLE),
     frames=5 * int(PRECISION / SAMPLE),
-    interval=0.75,
+    interval=0.5,
     blit=True,
     repeat=False,
 )

### 2020-04-12 Report/draft1.tex
update

@@
 \graphicspath{{"/Users/Kit/OneDrive/Documents/Computing/Trojan Asteroids/Report/Figures"}}
+\newcommand*{\subscript}[1]{\ensuremath{_\textrm{{\scriptsize #1}}}}
 
@@
 
-\section{History}
-This is a test with bib \ref{Ravindran2017}. Alternatively we consider \cite{Nakamura2008}.Nakamura2008
+%Use https://leancrew.com/all-this/2016/08/lagrange-points-redux/ to create your own contour plot?
+\section{Introduction}
+The Jupiter trojans, commonly known as the Trojan asteroids, are two large groups of asteroids that share the planet Jupiter's orbit around the Sun.
+These two groups are called the Greeks and the Trojans, named after opposing sides in the mythological Trojan war, and lead/trail Jupiter respectively in its orbit.
+They correspond to Jupiter's two stable Lagrange points: L\subscript{4}, lying 60° ahead of the planet in its orbit, and L\subscript{5}, 60° behind, with asteroids distributed in two elongated, curved regions around these Lagrangian points.
+CITE
 
+These Lagrange points are taken from Lagrange's initial analysis of the three-body problem in 1772 \cite{Lagrange1772}, were he demonstrated the existence of five equilibrium points for an object of negligible mass orbiting under the gravitational effect of two larger masses.
+Three of these equilibrium points, L\subscript{1} - L\subscript{3} lie on the line joining the two masses maxima in the potential function, making them unstable.
+Each of the remaining two points, L4 and L5, lies at the apex of an equilateral triangle with base equal to the separation of the two masses (see Fig.
+1); stable motion is possible around them despite them being maxima due to the Coriolis force.
+INSERT FIGURE - possibly from \cite{Marzari2002} which provides a solid analysis
+
+Include 1:1 orbit resonance details for the jupiter one
+
+\textit{
+The first Jupiter trojan discovered, 588 Achilles, was spotted in 1906 by German astronomer Max Wolf.[2] A total of 7,040 Jupiter trojans have been found as of October 2018.[3] By convention, they are each named from Greek mythology after a figure of the Trojan War, hence the name "Trojan".
+The total number of Jupiter trojans larger than 1 km in diameter is believed to be about 1 million, approximately equal to the number of asteroids larger than 1 km in the asteroid belt.[1] Like main-belt asteroids, Jupiter trojans form families.[4] }
+
+\subsection{Trojan Asteroids}
+previous lit
+Alternatively we consider \cite{Nakamura2008}
+
+\subsection{Orbit Geometry}
+Including mathematical assumptions in analysis
+
+\subsubsection{Assumptions}
+Circular orbit
+Constant Jupiter-Sun separation
+Planar orbit
+Negligible asteroid mass
+Newtonian gravity
+
+symmetries:
+trojan and greek symmetry (analysis focused on greeks)
+rotational symmetry - arbitrary initial point 
+direction of orbit
+"The combination of these symmetries allows the problem to be simplified. Such that
+only the Greeks, orbiting counter-clockwise, with perturbations at t = 0, need to be
+investigated.
+"
+
+\section{Methodology}
+\subsection{Integration Method}
+The default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) however this is non-stiff, giving a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
+This is larger than expected, suggesting the system of equations requires an unreasonable small step size for  numerically stability with respect to this numerical method, even regions where the solution curve is smooth \cite{Lambert1991}.
+This suggests the system is stiff, and solvers designed for this typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to the non-stiff solvers \cite{Byrne1987}.
+
+Instead the stiff "Radau" solver (an implicit Runge-Kutta method of the Radau IIA family of order 5 \cite{Hairer2010}) is used for increased stability \cite{Frank1985}, and achieves a deviation in asteroid position in the order of $ 10^{-13}$ AU instead.
+This also ensures stability in the rotating frame, with deviations of 0.76\% in asteroid separation from Jupiter over $ 10^{3} $ years, compared to 53\% for the best non-stiff solvers.
+
+%Can add more detail on A vs B stability (I believe B stability is relevant here but maybe check this)
+\section{Results}
+\subsection{Orbit Stability}
+\subsection{Wander Analysis}
+\subsubsection{Perturbations in z-direction}
+
+\section{Discussion}
+\section{Conclusion}
+
+%Talk about animations here too? or after solver method
 \printbibliography

### 2020-04-12 constants.py

@@ -7,8 +7,13 @@ M_S = 1  # solar mass
 M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
+<<<<<<< Updated upstream
 PRECISION = 5000  # evaluation points per orbit
 ORBIT_NUM = 10  # number of orbits
+=======
+PRECISION = 30  # evaluation points per orbit
+ORBIT_NUM = 100  # number of orbits
+>>>>>>> Stashed changes
 
 # DERIVED CONSTANTS
 solar_rad = R * M_P / (M_S + M_P)  # distance from origin to sun in CoM frame

### 2020-04-12 main_file.py

@@ -8,18 +8,24 @@ import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import solar_rad, planet_rad, period, time_span  # Derived constants
 
+#ROTATING FRAME
 orbit_sol = orbits.rotating_frame()
 # for i in range(len(orbit_sol.t)):
 #     print(orbit_sol.t[i], orbit_sol.y[0, i], orbit_sol.y[1, i])
 
+#DEVIATION FROM LAGRANGE POINT
 lagrange_x = planet_rad - R / 2
-plt.plot(time_span, (np.abs(lagrange_x - orbit_sol.y[0, :])))
-plt.yscale("log", basey=10)
-plt.ylabel("absolute error on x_pos")
-plt.xlabel("time")
+# plt.plot(time_span, (np.abs(lagrange_x - orbit_sol.y[0, :])))
+plt.plot(time_span, ((lagrange_x - orbit_sol.y[0, :])))
+# plt.yscale("log", basey=10)
+plt.ticklabel_format(axis="y", style="", scilimits=None)
+plt.ylabel("Magnitude of Deviation /AU")
+plt.xlabel("Time /years")
+plt.title("Asteroid Deviation from Lagrange point")
+plt.savefig("asteroid_deviation_linear.png")
 plt.show()
 
-
+#MOTION IN ROTATING FRAME
 plt.plot(
     orbits.solar_pos(0)[0],
     0,
@@ -48,12 +54,13 @@ plt.plot(orbit_sol.y[0, :], orbit_sol.y[1, :], label="Greeks")
 plt.title("Rotating Frame just Greeks")
 plt.show()
 
-
+#################################################################
+#STATIONARY FRAME
 orbit_sol2 = orbits.stationary_frame()
 # for i in range(len(orbit_sol2.t)):
 #     print(orbit_sol2.t[i], orbit_sol2.y[0, i], orbit_sol.y[1, i])
 
-
+#STATIONARY FRAME
 plt.plot(0, 0, label="CoM", color="black", marker="x", linestyle="None")
 plt.plot(
     orbits.solar_pos(time_span)[0],
@@ -77,3 +84,23 @@ plt.plot(orbit_sol2.y[0, :], orbit_sol2.y[1, :], label="Greeks")
 plt.title("Stationary Frame")
 plt.legend()
 plt.show()
+<<<<<<< Updated upstream
+=======
+
[…]

### 2020-04-12 orbits.py

@@ -45,6 +45,19 @@ def planet_pos(t=0):
     return np.array([planet_rad * np.cos(omega * t), planet_rad * np.sin(omega * t), 0])
 
 
+<<<<<<< Updated upstream
+=======
+def lagrange_pos(t=0):
+    return np.array(
+        [
+            greek_rad * np.cos(omega * t + greek_theta),
+            greek_rad * np.sin(omega * t + greek_theta),
+            0,
+        ]
+    )
+
+
+>>>>>>> Stashed changes
 # assuming massive bodies do not deviate from xy plane
 
 
@@ -91,7 +104,7 @@ def rotating_frame(y0_rot=(rcos, rsin, 0, 0, 0, 0)):
 
     return integrate.solve_ivp(
         fun=rot_derivatives,
-        method="LSODA",  ###on adams advice
+        method="Radau",  # Or LSODA for non-stiff alternative
         t_span=(0, ORBIT_NUM * period),
         y0=y0_rot,
         t_eval=time_span,  # selects points for storage
@@ -138,7 +151,7 @@ def stationary_frame(y0_stat=(rcos, rsin, 0, -vsin, vcos, 0)):
     """
     return integrate.solve_ivp(
         fun=stat_derivatives,
-        method="LSODA",  ##on adams advice
+        method="Radau",  # Or LSODA for non-stiff alternative
         t_span=(0, ORBIT_NUM * period),
         y0=y0_stat,
         t_eval=time_span,

### 2020-04-12 stat_frame_video.py

@@ -7,6 +7,11 @@ import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import lagrange, omega, time_span  # Derived constants
 
+<<<<<<< Updated upstream
+=======
+SAMPLE = int(100)  # Sample every n points
+
+>>>>>>> Stashed changes
 # Set up orbits for each body
 orbit_greeks = orbits.stationary_frame(
     (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)

### 2020-04-12 wander.py

@@ -1,0 +1,114 @@
+import math
+import matplotlib.pyplot as plt
+import numpy as np
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    omega,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+
+def perturb(data, max_pertubation_size=0.01):
+    """ Returns perturbed version of initial conditions array
+
+    data is the initial conditions, submitted as a numpy array
+    pertubation_size is the relative magnitude of the perturbation, measured as 
+    the percentage deivation from the lagrange point (for position)
+    Note that this will give a different value each time it is run
+    """
+    return data + data * (
+        max_pertubation_size * np.random.uniform(-1, 1, (np.shape(data)))
+    )  # random betweeon -1 and 1
+
+
+def max_wander(max_pertubation_size=0.01, samples=3):
+    """Returns maximum deviation over orbit in the rotating frame for given number of samples
+
+    This calculates the distange from the initial point (not the lagrange point)
+    pertubation_size determies the relative magnitude of the perturbation as in perturb()
+    Samples denotes the number of random pertubations sampled for gives perturbation size
+    """
+    wander = np.zeros((samples, 2))  # size of pertubation; size of wander
+    for n in range(samples):
+        init_cond = perturb(init_cond_rot, max_pertubation_size)
+        orbit = orbits.rotating_frame(init_cond)
+        sample_wander = np.zeros((len(orbit.t)))
+        for i in range(len(sample_wander)):
+            sample_wander[i] = np.linalg.norm(
+                orbit.y[0:3, i]
+                - init_cond[0:3]
+                # orbit.y[0:3, i]
+                # - lagrange[0:3]
+            )  # deviation in pos only
+        wander[n, 0] = np.linalg.norm(
+            init_cond[0:3] - lagrange[0:3]
+        )  # size of pertubation
+        wander[n, 1] = np.max(sample_wander)  # size of wander
+    return wander
+
+
+init_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+init_cond_stat = np.array(
[…]

### 2020-04-13 main_file.py
wander compared against pertubation magnitude

@@ -101,3 +101,20 @@ plt.ylabel("Separation / AU")
 plt.xlabel("Time / years")
 # plt.savefig("longseparation.png")
 plt.show()
+
+plt.plot(
+    time_span,
+    np.linalg.norm(
+        [
+            (orbit_sol2.y[0, :] - orbits.lagrange_pos(time_span)[0]),
+            (orbit_sol2.y[1, :] - orbits.lagrange_pos(time_span)[1]),
+        ],
+        axis=0,
+    ),
+)
+
+plt.title("Deviation of Greeks from Lagrnage Point")
+plt.ylabel("Separation / AU")
+plt.xlabel("Time / years")
+plt.savefig("asteroid_deviation_stat.png")
+plt.show()

### 2020-04-13 stat_frame_video.py

@@ -7,7 +7,7 @@ import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import lagrange, omega, time_span  # Derived constants
 
-SAMPLE = int(1)  # Sample every n points
+SAMPLE = int(100)  # Sample every n points
 
 # Set up orbits for each body
 orbit_sun = orbits.solar_pos(time_span)
@@ -97,6 +97,6 @@ plt.ylabel("Y distance/ AU")
 writer = animation.FFMpegWriter(
     fps=60, metadata=dict(artist="Kit Gallagher"), bitrate=1800
 )
-anim.save("movie.mp4", writer=writer)
+# anim.save("movie.mp4", writer=writer)
 
 plt.show()

### 2020-04-13 wander.py

@@ -110,5 +110,5 @@ plt.plot(
 
 plt.title("Stationary Frame")
 plt.legend()
-plt.savefig("drift5e-1.png")
+# plt.savefig("drift5e-1.png")
 plt.show()

### 2020-04-13 constants.py
non-vectorised wander

@@ -8,8 +8,8 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 
-PRECISION = 1000  # evaluation points per orbit
-ORBIT_NUM = 30  # number of orbits
+PRECISION = 30  # evaluation points per orbit
+ORBIT_NUM = 20  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-13 wander.py

@@ -5,12 +5,22 @@ import numpy as np
 import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import (
+    solar_rad,
     omega,
     time_span,
     lagrange,
 )  # Derived constants
 
+# Initial Conditions
+init_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+init_cond_stat = np.array(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
 
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+cos, sin = np.cos(greek_theta), np.sin(greek_theta)
+
+# Defined Functions
 def perturb(data, max_pertubation_size=0.01):
     """ Returns perturbed version of initial conditions array
 
@@ -28,10 +38,13 @@ def max_wander(max_pertubation_size=0.01, samples=3):
     """Returns maximum deviation over orbit in the rotating frame for given number of samples
 
     This calculates the distange from the initial point (not the lagrange point)
+    It returns a samples*3 array, giving the x and y components
+    of the pertubation, and then the maximum wander over this timespan
+
     pertubation_size determies the relative magnitude of the perturbation as in perturb()
     Samples denotes the number of random pertubations sampled for gives perturbation size
     """
-    wander = np.zeros((samples, 2))  # size of pertubation; size of wander
+    wander = np.zeros((samples, 3))  # size of pertubation; size of wander
     for n in range(samples):
         init_cond = perturb(init_cond_rot, max_pertubation_size)
         orbit = orbits.rotating_frame(init_cond)
@@ -40,29 +53,56 @@ def max_wander(max_pertubation_size=0.01, samples=3):
             sample_wander[i] = np.linalg.norm(
                 orbit.y[0:3, i]
                 - init_cond[0:3]
-                # orbit.y[0:3, i]
                 # - lagrange[0:3]
             )  # deviation in pos only
-        wander[n, 0] = np.linalg.norm(
-            init_cond[0:3] - lagrange[0:3]
-        )  # size of pertubation
[…]

### 2020-04-13 constants.py
Fixed mesh plot

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun
 
 
 PRECISION = 30  # evaluation points per orbit
-ORBIT_NUM = 20  # number of orbits
+ORBIT_NUM = 10  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-13 wander.py

@@ -69,15 +69,15 @@ def max_wander(max_pertubation_size=0.01, samples=3):
 
 
 def wander(position, samples=1):
-    """Returns maximum deviation over orbit in the rotating frame for given initial point
+    """Returns maximum wander over orbit in the rotating frame for given initial point
 
-    It returns the maximum wander from the initial point (not the lagrange point) over this timespan
-    pertubation_size determies the relative magnitude of the perturbation as in perturb()
+    Wander is the maximum deviation from the initial point (not the lagrange point) over this timespan
+    position gives the spatial pertubation from the lagrange point
     Samples denotes the number of random pertubations sampled within that point for gives perturbation size
     """
-    init_cond = init_cond_rot
-    init_cond[0] += position[0]
-    init_cond[1] += position[1]
+    init_cond = init_cond_rot + np.array(
+        (position[0], position[1], 0, 0, 0, 0)
+    )  # add pertubation
     orbit = orbits.rotating_frame(init_cond)
     wander_t = np.zeros((len(orbit.t)))
     for i in range(len(orbit.t)):
@@ -120,8 +120,8 @@ from matplotlib.colors import BoundaryNorm
 from matplotlib.ticker import MaxNLocator
 
 
-grid_size = 0.01
-sampling_points = 8
+grid_size = 0.001
+sampling_points = 6
 # make these smaller to increase the resolution
 dx, dy = grid_size / sampling_points, grid_size / sampling_points
 
@@ -130,14 +130,8 @@ y, x = np.mgrid[
     slice(-grid_size, grid_size + dy, dy), slice(-grid_size, grid_size + dx, dx)
 ]
 
-print(wander((x[0, 12], y[0, 12])))
-print(wander((x[0, 16], y[0, 16])))
-print(wander((x[3, 7], y[3, 7])))
-print(wander((x[3, 8], y[3, 8])))
-
 import time
 
-# z = np.sin(x) ** 10 + np.cos(10 + y * x) * np.cos(x)
 z = np.zeros_like(x)
 print(np.shape(z))
 for i, j in np.ndindex(z.shape):
@@ -150,20 +144,20 @@ for i, j in np.ndindex(z.shape):
 # x and y are bounds, so z should be the value *inside* those bounds.
 # Therefore, remove the last value from the z array.
 z = z[:-1, :-1]
[…]

### 2020-04-13 constants.py
Cleaned up mesh plots Working with printed version having 30 pixels, 30 precision and 50 orbits.

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun
 
 
 PRECISION = 30  # evaluation points per orbit
-ORBIT_NUM = 10  # number of orbits
+ORBIT_NUM = 5  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-13 orbits.py

@@ -101,7 +101,7 @@ def rotating_frame(y0_rot=(rcos, rsin, 0, 0, 0, 0)):
 
     return integrate.solve_ivp(
         fun=rot_derivatives,
-        method="Radau",  # Or LSODA for non-stiff alternative
+        method="LSODA",  # Or LSODA for non-stiff alternative
         t_span=(0, ORBIT_NUM * period),
         y0=y0_rot,
         t_eval=time_span,  # selects points for storage

### 2020-04-13 wander.py

@@ -115,13 +115,13 @@ def wander(position, samples=1):
 # need other way to define stability as other stable orbits away from lagrange point may form
 
 
-import matplotlib
-from matplotlib.colors import BoundaryNorm
-from matplotlib.ticker import MaxNLocator
+# import matplotlib
+# from matplotlib.colors import BoundaryNorm
+# from matplotlib.ticker import MaxNLocator
 
 
-grid_size = 0.001
-sampling_points = 6
+grid_size = 0.04
+sampling_points = 1
 # make these smaller to increase the resolution
 dx, dy = grid_size / sampling_points, grid_size / sampling_points
 
@@ -144,39 +144,57 @@ for i, j in np.ndindex(z.shape):
 # x and y are bounds, so z should be the value *inside* those bounds.
 # Therefore, remove the last value from the z array.
 z = z[:-1, :-1]
-# levels = MaxNLocator(nbins=15).tick_values(z.min(), z.max())
 
-# pick the desired colormap, sensible levels, and define a normalization
-# instance which takes data values and translates those into levels.#
-
+# pick the desired colormap
 # cmap = plt.get_cmap("PiYG")
-# norm = BoundaryNorm(levels, ncolors=cmap.N, clip=True)
-
-# fig, (ax0, ax1) = plt.subplots(nrows=2)
 
 fig = plt.figure()
 ax0 = fig.add_subplot()
 
-im = ax0.pcolormesh(x, y, z)  # cmap=cmap)  # , norm=norm)
-fig.colorbar(im, ax=ax0)
-ax0.set_title("Wander over spatial coordinates in rotating frame")
+im = ax0.pcolormesh(x, y, z, label="x")  # cmap=cmap)  # , norm=norm)
+cbar = fig.colorbar(im)
+cbar.ax.set_ylabel("Wander /AU")
 
+ax0.set_title("Wander from starting point in the rotating frame")
+ax0.set_xlabel("Deviation from Lagrange point in x direction /AU")
+ax0.set_ylabel("Deviation from Lagrange point in y direction /AU")
 
-# # contours are *point* based plots, so convert our bound into point
-# # centers
-# cf = ax1.contourf(
-#     x[:-1, :-1] + dx / 2.0, y[:-1, :-1] + dy / 2.0, z, levels=levels, cmap=cmap
-# )
-# fig.colorbar(cf, ax=ax1)
-# ax1.set_title("contourf with levels")
-
[…]

### 2020-04-14 constants.py
velocity phase space wander mesh plot setup

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun
 
 
 PRECISION = 30  # evaluation points per orbit
-ORBIT_NUM = 5  # number of orbits
+ORBIT_NUM = 50  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-14 wander.py

@@ -68,16 +68,22 @@ def max_wander(max_pertubation_size=0.01, samples=3):
     return wander
 
 
-def wander(position, samples=1):
+def wander(pertubation, samples=1, pertubation_type="position"):
     """Returns maximum wander over orbit in the rotating frame for given initial point
 
     Wander is the maximum deviation from the initial point (not the lagrange point) over this timespan
     position gives the spatial pertubation from the lagrange point
     Samples denotes the number of random pertubations sampled within that point for gives perturbation size
     """
-    init_cond = init_cond_rot + np.array(
-        (position[0], position[1], 0, 0, 0, 0)
-    )  # add pertubation
+    if pertubation_type == "position":
+        init_cond = init_cond_rot + np.array(
+            (pertubation[0], pertubation[1], 0, 0, 0, 0)
+        )  # add pertubation
+    else:
+        init_cond = init_cond_rot + np.array(
+            (0, 0, 0, pertubation[0], pertubation[1], 0)
+        )  # add pertubation
+
     orbit = orbits.rotating_frame(init_cond)
     wander_t = np.zeros((len(orbit.t)))
     for i in range(len(orbit.t)):
@@ -121,7 +127,7 @@ def wander(position, samples=1):
 
 
 grid_size = 0.04
-sampling_points = 1
+sampling_points = 30
 # make these smaller to increase the resolution
 dx, dy = grid_size / sampling_points, grid_size / sampling_points
 
@@ -133,10 +139,9 @@ y, x = np.mgrid[
 import time
 
 z = np.zeros_like(x)
-print(np.shape(z))
 for i, j in np.ndindex(z.shape):
     start = time.time()
-    z[i, j] = wander((x[i, j], y[i, j]))
+    z[i, j] = wander((x[i, j], y[i, j]), pertubation_type="velocity")
     end = time.time()
     print(str((i, j)) + " in time " + str(end - start) + " s")
 
@@ -156,10 +161,10 @@ cbar = fig.colorbar(im)
 cbar.ax.set_ylabel("Wander /AU")
 
 ax0.set_title("Wander from starting point in the rotating frame")
-ax0.set_xlabel("Deviation from Lagrange point in x direction /AU")
-ax0.set_ylabel("Deviation from Lagrange point in y direction /AU")
[…]

### 2020-04-14 wander.py
Sign implementation

@@ -18,10 +18,11 @@ init_cond_stat = np.array(
 )
 
 greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+print(greek_theta)
 cos, sin = np.cos(greek_theta), np.sin(greek_theta)
 
 # Defined Functions
-def perturb(data, max_pertubation_size=0.01):
+def perturb(data, max_pertubation_size):
     """ Returns perturbed version of initial conditions array
 
     data is the initial conditions, submitted as a numpy array
@@ -34,7 +35,7 @@ def perturb(data, max_pertubation_size=0.01):
     )  # random betweeon -1 and 1
 
 
-def max_wander(max_pertubation_size=0.01, samples=3):
+def max_wander(max_pertubation_size, samples):
     """Returns maximum deviation over orbit in the rotating frame for given number of samples
 
     This calculates the distange from the initial point (not the lagrange point)
@@ -44,28 +45,37 @@ def max_wander(max_pertubation_size=0.01, samples=3):
     pertubation_size determies the relative magnitude of the perturbation as in perturb()
     Samples denotes the number of random pertubations sampled for gives perturbation size
     """
-    wander = np.zeros((samples, 3))  # size of pertubation; size of wander
+    output = np.zeros((samples, 3))  # size of pertubation; size of wander
     for n in range(samples):
         init_cond = perturb(init_cond_rot, max_pertubation_size)
         orbit = orbits.rotating_frame(init_cond)
         sample_wander = np.zeros((len(orbit.t)))
+
+        perturb_theta = np.arctan(
+            (init_cond[1] - init_cond_rot[1]) / (init_cond[0] - init_cond_rot[0])
+        )
+        sign = np.sign(init_cond[1] - init_cond_rot[1]) * np.sign(
+            init_cond[0] - init_cond_rot[0]
+        )
         for i in range(len(sample_wander)):
             sample_wander[i] = np.linalg.norm(
                 orbit.y[0:3, i]
                 - init_cond[0:3]
                 # - lagrange[0:3]
             )  # deviation in pos only
-        # wander[n, 0] = np.linalg.norm(
-        #     init_cond[0:3] - lagrange[0:3]
[…]

### 2020-04-14 wander.py
radial and tangential wander plots

@@ -18,8 +18,7 @@ init_cond_stat = np.array(
 )
 
 greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
-print(greek_theta)
-cos, sin = np.cos(greek_theta), np.sin(greek_theta)
+
 
 # Defined Functions
 def perturb(data, max_pertubation_size):
@@ -39,7 +38,7 @@ def max_wander(max_pertubation_size, samples):
     """Returns maximum deviation over orbit in the rotating frame for given number of samples
 
     This calculates the distange from the initial point (not the lagrange point)
-    It returns a samples*3 array, giving the x and y components
+    It returns a samples*3 array, giving the radial and tangential components
     of the pertubation, and then the maximum wander over this timespan
 
     pertubation_size determies the relative magnitude of the perturbation as in perturb()
@@ -54,25 +53,18 @@ def max_wander(max_pertubation_size, samples):
         perturb_theta = np.arctan(
             (init_cond[1] - init_cond_rot[1]) / (init_cond[0] - init_cond_rot[0])
         )
-        sign = np.sign(init_cond[1] - init_cond_rot[1]) * np.sign(
-            init_cond[0] - init_cond_rot[0]
-        )
+
         for i in range(len(sample_wander)):
             sample_wander[i] = np.linalg.norm(
                 orbit.y[0:3, i]
                 - init_cond[0:3]
                 # - lagrange[0:3]
             )  # deviation in pos only
-        output[n, 0] = (
-            np.linalg.norm((init_cond[0:3] - lagrange[0:3]))
-            * np.abs(np.cos(greek_theta - perturb_theta))
-            # * sign
+        output[n, 0] = np.linalg.norm((init_cond[0:3] - lagrange[0:3])) * np.abs(
+            np.cos(greek_theta - perturb_theta)
         )  # radial component of pertubation
-        print(np.cos(greek_theta - perturb_theta))
-        output[n, 1] = (
-            np.linalg.norm((init_cond[0:3] - lagrange[0:3]))
-            * np.abs(np.sin(greek_theta - perturb_theta))
-            # * sign
[…]

### 2020-04-14 wander.py
random circling in circle not rectangle radial and tangential plots working

@@ -12,8 +12,8 @@ from constants import (
 )  # Derived constants
 
 # Initial Conditions
-init_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
-init_cond_stat = np.array(
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+initial_cond_stat = np.array(
     (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
 )
 
@@ -21,17 +21,29 @@ greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 
 
 # Defined Functions
-def perturb(data, max_pertubation_size):
+def perturb(initial_cond, max_pertubation_size=0.01):
     """ Returns perturbed version of initial conditions array
 
-    data is the initial conditions, submitted as a numpy array
+    init_cond is the initial conditions, submitted as a numpy array
     pertubation_size is the relative magnitude of the perturbation, measured as 
     the percentage deivation from the lagrange point (for position)
     Note that this will give a different value each time it is run
     """
-    return data + data * (
-        max_pertubation_size * np.random.uniform(-1, 1, (np.shape(data)))
-    )  # random betweeon -1 and 1
+    # return data + data * (
+    #     max_pertubation_size * np.random.uniform(-1, 1, (np.shape(data)))
+    # )  # random between -1 and 1, perturbs position only for initial_cond_rot
+    # theta = np.random.uniform(0, 2 * math.pi)
+    # return data + data * (
+    #     max_pertubation_size
+    #     * np.random.rand()
+    #     * np.array((math.cos(theta), math.sin(theta), 0, 0, 0, 0))
+    # )  # random betweeon -1 and 1, perturbs position only for initial_cond_rot
+    rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+    while np.linalg.norm(rand_array[0:2]) > 1:
+        rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+    return initial_cond + initial_cond * (
+        max_pertubation_size * rand_array
+    )  # random between -1 and 1, selects points in circle of radius 1
 
 
 def max_wander(max_pertubation_size, samples):
[…]

### 2020-04-18 Report/draft1.tex
Report update

--- 2020-04-06 Report/draft1.tex
@@
 \graphicspath{{"/Users/Kit/OneDrive/Documents/Computing/Trojan Asteroids/Report/Figures"}}
+\newcommand*{\subscript}[1]{\ensuremath{_\textrm{{\scriptsize #1}}}}
+
+\usepackage{bm}
+\usepackage{tikz}
+\usetikzlibrary{calc}
 
@@
 
-\section{History}
-This is a test with bib \ref{Ravindran2017}. Alternatively we consider \cite{Nakamura2008}.Nakamura2008
+%Use https://leancrew.com/all-this/2016/08/lagrange-points-redux/ to create your own contour plot?
+\section{Introduction}
+The Jupiter trojans, commonly known as the Trojan asteroids, are two large groups of asteroids that share the planet Jupiter's orbit around the Sun in a 1:1 orbit resonance.
+These two groups are called the Greeks and the Trojans, named after opposing sides in the mythological Trojan war, and lead/trail Jupiter respectively in its orbit.
+They correspond to Jupiter's two stable Lagrange points: L\subscript{4}, lying 60° ahead of the planet in its orbit, and L\subscript{5}, 60° behind, with asteroids distributed in two elongated, curved regions around these Lagrangian points.
 
+The first Jupiter trojan, 588 Achilles, was discovered in 1906 by the German astronomer Max Wolf \cite{Nicholson1961}, and a total of 7642 Jupiter trojans have been found as of February 2020 \cite{IAU2020}.
+
+Research into Jupiter's trojan asteroids continues, with the particular focus on their origins reliant on an understanding of their orbit stability \cite{DiSisto2019}, \cite{Nesvorn2018}.
+This informs studies into their composition \cite{Brown2016}, as travel to these asteroids is considered for their potential in mineral mining \cite{Okada2017} \cite{Levison2016}.
+
+The purpose of this report is to use numerical simulation techniques to investigate the stability of orbits about these Lagrange points, demonstrating the asteroid oscillate about these points under small perturbations and quantifying the absolute distance of the asteroids from the Lagrange point (the wander) during their orbits.
+The impact of variation in planetary/solar mass on asteroid orbit stability will also be considered.
+\textit{Signpost what is included in each section?}
+
+% \cite{Nakamura2008} gives spatial distribution
+
+\section{Theoretical Background?}
+
+\subsection{Lagrange Points}
+The asteroids exist at/near Lagrange points, defined in Lagrange's initial analysis of the three-body problem in 1772 \cite{Lagrange1772}, were he demonstrated the existence of five equilibrium points for an object of negligible mass orbiting under the gravitational effect of two larger masses.
+Three of these equilibrium points, L\subscript{1}-L\subscript{3} lie on the line joining the two masses, while each of the remaining two points, L\subscript{4} and L\subscript{5}, lie at the apex of an equilateral triangle with base equal to the separation of the two masses (see Figure \ref{fig:lagrangepoints}).
+Despite all these points being potential maxima, stable motion is possible around L\subscript{4} and L\subscript{5} due to the Coriolis force \cite{Lissauer2014}.
+%lissauer also says stability is provided that the most massive body has at least 25 times the mass of the secondary
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\linewidth]{Figures/lagrange_points}
+	\caption{The location of the five Lagrange equilibrium points in the circular-restricted three-body problem.
+The solar and planetary masses are denoted by the large and small filled circles, and the letters P, H, and T denote passing, horseshoe, and tadpole orbits respectively.
+Note that the two masses form an equilateral triangle with each of the L\subscript{4} and L\subscript{5} points.
+Reproduced from Marzari et al.
+\cite{Marzari2002}}
+	\label{fig:lagrangepoints}
+\end{figure}
+
+While orbits between Lagrange points are also possible, this report will focus on the tadpole orbits observed as asteroids deviate from L4 and L5 \cite{Murray1999}. \textit{More detail?}
+
+\subsection{Theoretical Model}
+The three body problem, where the dynamics of three interacting bodies are determined given their initial positions and velocities, has no analytical (closed-form) solution in the general case \cite{Barrow2008}.
+
+In this report, I will consider the circular, restricted, three-body problem, where two of the bodies move in circular, coplanar orbits about their common centre of mass, unaffected by the negligible mass of the third body.
+I will also assume that all interactions are via Newtonian gravity.
+
+Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in earth years and mass in multiples of the solar mass.
+
+The system of differential coordinates determine the position and velocity of the asteroids, with two equations per spatial coordinate.
+\begin{equation}
+\frac{dr_{i}}{dt} = v_{i}, \frac{dv_{i}}{dt} = g_{i}, \quad i = x,y 
+\end{equation}
+
+In this, $g_{i}$ is given by:
+\begin{equation}
+\textbf{g}= - \frac{G M_{s}}{\lvert \textbf{r} - \textbf{r}_{s} \rvert ^{3}} (\textbf{r} - \textbf{r}_{s})
+		 - \frac{G M_{p}}{\lvert \textbf{r} - \textbf{r}_{p} \rvert ^{3}} (\textbf{r} - \textbf{r}_{p})
+\end{equation}
+where the subscripts \textit{s} and \textit{p} refer to solar and planetary properties respectively.
+
+We may also consider a frame rotating at the same speed as the massive bodies.
+As there is 1:1 orbital resonance between Jupiter and the asteroids, all three bodies are stationary in this frame.
+This significantly increases the accuracy of numerical simulations, as \textbf{WHY}
+
+When transforming into this rotating non-inertial frame, $g_{i}$ gains an additional virtual force term with coupling between the spatial coordinates.
+This is given below as the sum of the centripetal and Coriolis forces:
+
+\begin{equation}
+\Delta g_{i} = \Omega^{2} r_{i} - 2[\bm{\Omega} \times \textbf{v}]_{i}
+\end{equation}
+where $\Omega$ is the angular speed of the rotating frame, and $\textbf{v}$ is the velocity of the asteroid within this frame.
+
+\subsubsection{Assumptions - FIX}
+\begin{itemize}
+	\item Circular orbit
+	\item Constant Jupiter-Sun separation
+	\item Coplanar orbits
+	\item Negligible asteroid mass
+	\item Newtonian gravity
+\end{itemize}
+
+I also applied the following symmetries:
+trojan and greek symmetry (analysis focused on greeks)
+rotational symmetry - arbitrary initial point 
+direction of orbit
+"The combination of these symmetries allows the problem to be simplified. Such that
+only the Greeks, orbiting counter-clockwise, with perturbations at t = 0, need to be
+investigated.
+"
+
+\subsection{Orbit Geometry}
+As the three bodies considered here form an equilateral triangle in the initial equilibrium state, as depicted in Figure \ref{fig:geometry}, we can derive the polar coordinates of each body with respect to the centre of mass about which the bodies orbit.
+
+\begin{figure}
+	\centering
+	\resizebox{\columnwidth}{!}{%
+		\begin{tikzpicture}
+		\draw (0,0) -- (6,0) -- (3, {6*sin(60)} ) node[above]{Asteroids} node[midway,right]{$a$} -- cycle node[midway,left]{$a$};
+		\draw  (1,0) arc (0:60:1cm)  ;
+		%\node [label={[shift = {(10mm,5mm)}]{$\frac{\pi}{3}$}}] {};
+		\node [label={[shift = {(13:11mm)}]{$\frac{\pi}{3}$}}] {};
+		\node [label={[shift = {(-7mm,-7mm)}]{Sun}}] {};
+		\node [label={[shift = {(67mm,-7mm)}]{Planet}}] {};
+
+		\fill[gray] ({6*(1/4)}, 0) circle (0.5mm);
+		\draw[gray] ({6*(1/4)}, 0) -- (3, {6*sin(60)} ) node[midway,left]{$r_{a}$};
+		\draw[dashed, gray] (3, 0) -- (3, {6*sin(60)} )node[midway,right]{$\frac{\sqrt{3}a}{2}$};
+		\draw[gray]  ({6*(1/4) + 1}, 0) arc (0:{atan(4*sin(60))}:1cm)  ;
+		\node [label={[gray, shift = {(25mm,3mm)}]$\theta$}] {};
+		\draw[gray] (2.75,0) -- (2.75, 0.25) -- (3, 0.25);
+
+		\draw[arrows=<-](0,-0.5)--(0.53,-0.5);
+		\node at (0.75, -0.5) {$R_s$};
+		\draw[arrows=->](1.0,-0.5)--(1.48,-0.5);
+
+		\draw[arrows=<-](1.52,-0.5)--(3.53,-0.5);
+		\node at (3.75, -0.5) {$R_p$};
+		\draw[arrows=->](4.00,-0.5)--(6,-0.5);
+
+		%\draw[arrows=<-](0,-1)--(2.88,-1);
+		%\node at (3, -1) {$a$};
+		%\draw[arrows=->](3.1,-1)--(6, -1);
+
+		\fill[black] (0,0) circle (3.5mm);% node[anchor=north west] {Sun};
+		\fill[black] (6,0) circle (1mm);
+		\fill[black] (3, {6*sin(60)} ) circle (0.5mm);
+
+		\end{tikzpicture}
+	}	
+	\caption{A geometric depiction of the three-body system, in the case where the planet has a mass equal to a third of the sun, and asteroids are considered at the L\subscript{4} point.
+$ R_{s} $ and $ R_{p}$ denote the (fixed) radii from the centre of mass (the grey point) to the Sun and the planet respectively, while $ r_{a} $ denotes the radius of the asteroids.}
+	\label{fig:geometry}
+\end{figure}
+
+Using standard trigonometric relations, it is simple to show that the values $ r_{a} $ and $ \theta $ are given by:
+
+\begin{equation}
+r_{a} = \sqrt{a^{2} + R_{s} R_{p}} , \enspace \theta =  \tan^{-1} \left( \frac{a \sin(\frac{\pi}{3})}{R_{p} - \frac{a}{2}} \right)
+\end{equation}
+
+Furthermore, the Lagrange point in Cartesian coordinates based about the centre of mass is easily found to be:
+
+\begin{equation}
+(x,y) = \left( R_{p} - \frac{a}{2}, \frac{ \sqrt{3} a}{2} \right)
+\end{equation}
+
+Finally, equating the gravitational and centripetal forces on the planet allows the derivation of its (and all other bodies') orbital velocity:
+
+\begin{equation}
+\Omega = \sqrt{\frac{G (M_{s} + M_{p})}{a^{3}}}
+\end{equation}
+
+\section{Methodology}
+\subsection{Integration Method}
+%Put this with theory as "numerical methods"?
+%Then all theory could be in methodology?
+The default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) however this is non-stiff, giving a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
+This is larger than expected, suggesting the system of equations requires an unreasonable small step size for  numerically stability with respect to this numerical method, even regions where the solution curve is smooth \cite{Lambert1991}.
+This suggests the system is stiff, and solvers designed for this typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to the non-stiff solvers \cite{Byrne1987}.
+
+Instead the stiff "Radau" solver (an implicit Runge-Kutta method of the Radau IIA family of order 5 \cite{Hairer2010}) is used for increased stability \cite{Frank1985}, and achieves a deviation in asteroid position in the order of $ 10^{-13}$ AU instead.
+This also ensures stability in the rotating frame, with deviations of 0.76\% in asteroid separation from Jupiter over $ 10^{3} $ years, compared to 53\% for the best non-stiff solvers.
+
+%Can add more detail on A vs B stability (I believe B stability is relevant here but maybe check this)
+\section{Results}
+\subsection{Orbit Stability}
+\subsection{Wander Analysis}
+\subsubsection{Perturbations in z-direction}
+
+\section{Discussion}
+\section{Conclusion}
+
+%Talk about animations here too? or after solver method
 \printbibliography

### 2020-04-18 wander.py

@@ -110,7 +110,7 @@ def wander(pertubation, samples=1, pertubation_type="position"):
     return np.max(wander_t)
 
 
-wander_data = max_wander(max_pertubation_size=0.001, samples=200)
+wander_data = max_wander(max_pertubation_size=0.001, samples=20)
 
 fig = plt.figure()
 ax = fig.add_subplot()

### 2020-04-18 constants.py
Plot tadpole orbits Occurs for radial perturbations within wander_orbits

@@ -8,8 +8,8 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 
-PRECISION = 30  # evaluation points per orbit
-ORBIT_NUM = 50  # number of orbits
+PRECISION = 100  # evaluation points per orbit
+ORBIT_NUM = 100  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-18 orbits.py

@@ -101,7 +101,7 @@ def rotating_frame(y0_rot=(rcos, rsin, 0, 0, 0, 0)):
 
     return integrate.solve_ivp(
         fun=rot_derivatives,
-        method="LSODA",  # Or LSODA for non-stiff alternative
+        method="Radau",  # Or LSODA for non-stiff alternative
         t_span=(0, ORBIT_NUM * period),
         y0=y0_rot,
         t_eval=time_span,  # selects points for storage

### 2020-04-18 planet_mass_variation.py

@@ -1,0 +1,52 @@
+import math
+import matplotlib.pyplot as plt
+import numpy as np
+
+import orbits
+
+# from wander import wander
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    solar_rad,
+    planet_rad,
+    omega,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+# Initial Conditions
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+initial_cond_stat = np.array(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+cos = math.cos(greek_theta)
+sin = math.sin(greek_theta)
+
+
+def wander(pertubation, samples=1, pertubation_type="position"):
+    """Returns maximum wander over orbit in the rotating frame for given initial point
+
+    Wander is the maximum deviation from the initial point (not the lagrange point) over this timespan
+    position gives the spatial pertubation from the lagrange point
+    Samples denotes the number of random pertubations sampled within that point for gives perturbation size
+    """
+    if pertubation_type == "position":
+        initial_cond = initial_cond_rot + np.array(
+            (pertubation[0], pertubation[1], 0, 0, 0, 0)
+        )  # add pertubation
+    else:
+        initial_cond = initial_cond_rot + np.array(
+            (0, 0, 0, pertubation[0], pertubation[1], 0)
+        )  # add pertubation
+
+    orbit = orbits.rotating_frame(initial_cond)
+    wander_t = np.zeros((len(orbit.t)))
+    for i in range(len(orbit.t)):
+        wander_t[i] = np.linalg.norm(
+            orbit.y[0:3, i]
+            - initial_cond[0:3]
+            # - lagrange[0:3]
+        )  # deviation in pos only
+    return np.max(wander_t)

### 2020-04-18 wander.py

@@ -115,7 +115,7 @@ wander_data = max_wander(max_pertubation_size=0.001, samples=20)
 fig = plt.figure()
 ax = fig.add_subplot()
 
-coeff = np.polyfit(wander_data[:, 0], wander_data[:, 2], 1)
+coeff = np.polyfit(wander_data[:, 0], wander_data[:, 2], 1)  # for line of best fit
 plt.plot(
     [wander_data[:, 0].min(), wander_data[:, 0].max()],
     [

### 2020-04-18 wander_orbits.py

@@ -1,0 +1,98 @@
+import math
+import matplotlib.pyplot as plt
+import numpy as np
+
+import orbits
+
+# from wander import wander
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    solar_rad,
+    planet_rad,
+    omega,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+# Initial Conditions
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+initial_cond_stat = np.array(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+cos = math.cos(greek_theta)
+sin = math.sin(greek_theta)
+
+
+def wander(pertubation, samples=1, pertubation_type="position"):
+    """Returns maximum wander over orbit in the rotating frame for given initial point
+
+    Wander is the maximum deviation from the initial point (not the lagrange point) over this timespan
+    position gives the spatial pertubation from the lagrange point
+    Samples denotes the number of random pertubations sampled within that point for gives perturbation size
+    """
+    if pertubation_type == "position":
+        initial_cond = initial_cond_rot + np.array(
+            (pertubation[0], pertubation[1], 0, 0, 0, 0)
+        )  # add pertubation
+    else:
+        initial_cond = initial_cond_rot + np.array(
+            (0, 0, 0, pertubation[0], pertubation[1], 0)
+        )  # add pertubation
+
+    orbit = orbits.rotating_frame(initial_cond)
+    wander_t = np.zeros((len(orbit.t)))
+    for i in range(len(orbit.t)):
+        wander_t[i] = np.linalg.norm(
+            orbit.y[0:3, i]
+            - initial_cond[0:3]
+            # - lagrange[0:3]
+        )  # deviation in pos only
+    return np.max(wander_t)
+
+
+# ROTATING FRAME
+# orbit_sol = orbits.rotating_frame(
+#     (initial_cond_rot + (0, 0, 0, 0, 0, 0))
+# )  # no perturbation
[…]

### 2020-04-20 Report/draft1.tex
Further report writing (And changes to planet mass variation)

@@
 \graphicspath{{"/Users/Kit/OneDrive/Documents/Computing/Trojan Asteroids/Report/Figures"}}
-\newcommand*{\subscript}[1]{\ensuremath{_\textrm{{\scriptsize #1}}}}
-
-\usepackage{bm}
-\usepackage{tikz}
-\usetikzlibrary{calc}
 
@@
 
-Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in earth years and mass in multiples of the solar mass.
-
 The system of differential coordinates determine the position and velocity of the asteroids, with two equations per spatial coordinate.
@@
 "
+(This reduces computing time not hvaing to solve for trojans as well)
 
@@
 \section{Methodology}
+This system of coupled first-order differential equations was solved using the scipy solve\_ivp function.
+The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years.
+Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in earth years and mass in multiples of the solar mass, and to prevent floating point overflows due to the magnitude of the quantities considered in SI units.
+
+The wander was defined as the maximum distance of the asteroid from the initial point during the orbit (for small perturbations this is also approximately the separation from the Lagrange point).
+Initial conditions are defined by the Lagrange point in each frame, with the initial velocity in the stationary frame defined by the period of Jupiter's orbit, and split into Cartesian components.
+
 \subsection{Integration Method}
@@
 %Then all theory could be in methodology?
-The default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) however this is non-stiff, giving a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
+Within solve\_ivp, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) however this is non-stiff, giving a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
 This is larger than expected, suggesting the system of equations requires an unreasonable small step size for  numerically stability with respect to this numerical method, even regions where the solution curve is smooth \cite{Lambert1991}.
@@
 %Can add more detail on A vs B stability (I believe B stability is relevant here but maybe check this)
+
+\subsection{Programme Structure}
+Global constants such as solar mass, and sun\textendash planet separation, along with derived values from these such as orbital period and solar radius from centre of mass, as given in an importable python module "\textit{constants}".
+
+Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit (under different sampling routines) are implemented in "\textit{wander}".
+Further files then import these modules and produce the plots given in this report.
+
+To consider a varying planet mass, it was considered preferable to avoid reconstructing all functions to take this as an argument as this requires re-evaluating all initial derived constants.
+Therefore, I iterate over alternative masses, re-defining constant values in this instance, and then directly import the required functions to compute the orbit.
+\textbf{REPHRASE?}
+
+Complete code listings are given in appendix \ref{Code}.
+
 \section{Results}
@@
 \printbibliography
+
+\onecolumn
+\begin{appendices}
+
+\section{Code 1} \label{Code}
+\begin{lstlisting}[language=Python]
+import math
+import numpy as np
+from scipy import integrate
+import matplotlib.pyplot as plt
+
+from constants import G, M_S, M_P, R, ORBIT_NUM, PRECISION  # User defined constants
+from constants import (
+	solar_rad,
+	planet_rad,
+	period,
+	omega,
+	time_span,
+)  # Derived constants
+
+\end{lstlisting}
+%Can add my own style guide to this, see suggestions on https://www.overleaf.com/learn/latex/code_listing
+%\section{Code 2}
+
+\end{appendices}
 

### 2020-04-20 Report/structure.tex
Further report writing (And changes to planet mass variation)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Wenneker Article
% Structure Specification File
% Version 1.0 (28/2/17)
%
% This file originates from:
% http://www.LaTeXTemplates.com
%
% Authors:
% Frits Wenneker
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\usepackage[english]{babel} % English language hyphenation

\usepackage{microtype} % Better typography

\usepackage{verbatim} % Allows mulitline commenting

\usepackage{amsmath,amsfonts,amsthm} % Math packages for equations

\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'

\usepackage[hang, small, labelfont=bf, up, textfont=it]{caption} % Custom captions under/above tables and figures

\usepackage{subcaption}

\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")

\usepackage{graphicx} % Required for adding images

\usepackage{enumitem} % Required for customising lists
\setlist{noitemsep} % Remove spacing between bullet/numbered list elements

\usepackage{sectsty} % Enables custom section titles
\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands (Helvetica)

\usepackage{bm} % Used for bold font in math environments

\usepackage{tikz} % Used for tikz diagrams
\usetikzlibrary{calc} % Enable math in tikz code

\newcommand*{\subscript}[1]{\ensuremath{_\textrm{{\scriptsize #1}}}}  %Non italic subscripts

%----------------------------------------------------------------------------------------
%	MARGINS AND SPACING
%----------------------------------------------------------------------------------------

\usepackage{geometry} % Required for adjusting page dimensions

\geometry{
	top=1cm, % Top margin
	bottom=1.5cm, % Bottom margin
	left=2cm, % Left margin
	right=2cm, % Right margin
	includehead, % Include space for a header
	includefoot, % Include space for a footer
	%showframe, % Uncomment to show how the type block is set on the page
}

\setlength{\columnsep}{7mm} % Column separation width

%----------------------------------------------------------------------------------------
%	FONTS
%----------------------------------------------------------------------------------------

\usepackage[T1]{fontenc} % Output font encoding for international characters
\usepackage[utf8]{inputenc} % Required for inputting international characters

\usepackage{XCharter} % Use the XCharter font

%----------------------------------------------------------------------------------------
%	HEADERS AND FOOTERS
%----------------------------------------------------------------------------------------

\usepackage{fancyhdr} % Needed to define custom headers/footers
\pagestyle{fancy} % Enables the custom headers/footers

\renewcommand{\headrulewidth}{0.0pt} % No header rule
\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule

\renewcommand{\sectionmark}[1]{\markboth{#1}{}} % Removes the section number from the header when \leftmark is used

%\nouppercase\leftmark % Add this to one of the lines below if you want a section title in the header/footer

% Headers
\lhead{} % Left header
\chead{\textit{\thetitle}} % Center header - currently printing the article title
\rhead{} % Right header

% Footers
\lfoot{} % Left footer
\cfoot{} % Center footer
\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % Right footer, "Page 1 of 2"

\fancypagestyle{firstpage}{ % Page style for the first page with the title
	\fancyhf{}
	\renewcommand{\footrulewidth}{0pt} % Suppress footer rule
}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\authorstyle}[1]{{\large\usefont{OT1}{phv}{b}{n}\color{NavyBlue}#1}} % Authors style (Helvetica)

\newcommand{\institution}[1]{{\footnotesize\usefont{OT1}{phv}{m}{sl}\color{Black}#1}} % Institutions style (Helvetica)

\usepackage{titling} % Allows custom title configuration

\newcommand{\HorRule}{\color{SteelBlue}\rule{\linewidth}{1pt}} % Defines the gold horizontal rule around the title

\pretitle{
	\vspace{-30pt} % Move the entire title section up
	\HorRule\vspace{10pt} % Horizontal rule before the title
	\fontsize{32}{36}\usefont{OT1}{phv}{b}{n}\selectfont % Helvetica
	\color{Navy} % Text colour for the title and author(s)
}

\posttitle{\par\vskip 15pt} % Whitespace under the title

\preauthor{} % Anything that will appear before \author is printed

\postauthor{ % Anything that will appear after \author is printed
	\vspace{10pt} % Space before the rule
	\par\HorRule % Horizontal rule after the title
	\vspace{20pt} % Space after the title section
}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\usepackage{lettrine} % Package to accentuate the first letter of the text (lettrine)
\usepackage{fix-cm}	% Fixes the height of the lettrine

\newcommand{\initial}[1]{ % Defines the command and style for the lettrine
	\lettrine[lines=3,findent=4pt,nindent=0pt]{% Lettrine takes up 3 lines, the text to the right of it is indented 4pt and further indenting of lines 2+ is stopped
		\color{NavyBlue}% Lettrine colour gold DarkGoldenRod
		{#1}% The letter
	}{}%
}

\usepackage{xstring} % Required for string manipulation

\newcommand{\lettrineabstract}[1]{
	\StrLeft{#1}{1}[\firstletter] % Capture the first letter of the abstract for the lettrine
	\initial{\firstletter}\textbf{\StrGobbleLeft{#1}{1}} % Print the abstract with the first letter as a lettrine and the rest in bold
}

%	BIBLIOGRAPHY

\usepackage[backend=biber,style=phys,natbib=true,doi=false]{biblatex} 
%Can equally use numeric citation style without extra phys packaging (but doesn't change capitalisation), or authoryear for alphabetical listing without the codes (resembles APA)

%\addbibresource{references.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography

%   CODE LISTING
\usepackage[title]{appendix} %in appendix

\usepackage{listings}
\usepackage{pythontex}

\usepackage{setspace}
\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,0} %was 1 at the end
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{0.98,0.98,0.98}
%\lstdefinelanguage{Python}{  %replaces two lines below
\lstset{
	language=Python,
	showspaces=false,
	showtabs=false,
	showstringspaces=false,
	%frame=l,   % gives black line by text
	tabsize=4,
	% Basic
	basicstyle=\ttfamily\small\setstretch{1},  % determines font and other style
	%backgroundcolor=\color{Background},      % gives background colour for all code
	% Comments
	commentstyle=\color{Comments}\slshape,
	% Strings
	stringstyle=\color{Strings},
	morecomment=[s][\color{Strings}]{"""}{"""},
	morecomment=[s][\color{Strings}]{'''}{'''},
	% keywords
	morekeywords={import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert},
	keywordstyle={\color{Keywords}\bfseries},
	% additional keywords
	morekeywords={[2]@invariant,pylab,numpy,np,scipy},
	keywordstyle={[2]\color{Decorators}\slshape},
	emph={self},
	emphstyle={\color{self}\slshape},
	%
}

### 2020-04-20 constants.py

@@ -8,8 +8,8 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 
-PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 100  # number of orbits
+PRECISION = 50  # evaluation points per orbit
+ORBIT_NUM = 50  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-20 planet_mass_variation.py

@@ -14,39 +14,31 @@ from constants import (
     lagrange,
 )  # Derived constants
 
-# Initial Conditions
-initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
-initial_cond_stat = np.array(
-    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
-)
+mass = np.linspace(0.001, 0.02, 100)  # range of planetary masses
+max_wander = np.zeros_like(mass)
+for n in range(len(mass)):
+    import constants
 
-greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
-cos = math.cos(greek_theta)
-sin = math.sin(greek_theta)
+    constants.M_P = mass[n]
+    constants.solar_rad = R * mass[n] / (M_S + mass[n])
+    constants.planet_rad = R * M_S / (M_S + mass[n])
+    constants.period = math.sqrt(R ** 3 / (M_S + mass[n]))
+    constants.omega = 2 * np.pi / constants.period  # angular velocity of frame
+    constants.lagrange = (constants.planet_rad - R / 2, R * math.sqrt(3) / 2, 0)
 
+    import orbits
 
-def wander(pertubation, samples=1, pertubation_type="position"):
-    """Returns maximum wander over orbit in the rotating frame for given initial point
-
-    Wander is the maximum deviation from the initial point (not the lagrange point) over this timespan
-    position gives the spatial pertubation from the lagrange point
-    Samples denotes the number of random pertubations sampled within that point for gives perturbation size
-    """
-    if pertubation_type == "position":
-        initial_cond = initial_cond_rot + np.array(
-            (pertubation[0], pertubation[1], 0, 0, 0, 0)
-        )  # add pertubation
-    else:
-        initial_cond = initial_cond_rot + np.array(
-            (0, 0, 0, pertubation[0], pertubation[1], 0)
-        )  # add pertubation
-
+    initial_cond = np.array((constants.lagrange[0], constants.lagrange[1], 0, 0, 0, 0))
     orbit = orbits.rotating_frame(initial_cond)
     wander_t = np.zeros((len(orbit.t)))
     for i in range(len(orbit.t)):
         wander_t[i] = np.linalg.norm(
-            orbit.y[0:3, i]
-            - initial_cond[0:3]
[…]

### 2020-04-20 Report/draft1.tex
largrange_stability created finds wander for un-perturbed case Also times ODE solver for performance

@@
 \section{Methodology}
-This system of coupled first-order differential equations was solved using the scipy solve\_ivp function.
+This system of coupled first-order ordinary differential equations (ODEs) was solved using the scipy solve\_ivp function.
 The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years.
@@
 Instead the stiff "Radau" solver (an implicit Runge-Kutta method of the Radau IIA family of order 5 \cite{Hairer2010}) is used for increased stability \cite{Frank1985}, and achieves a deviation in asteroid position in the order of $ 10^{-13}$ AU instead.
-This also ensures stability in the rotating frame, with deviations of 0.76\% in asteroid separation from Jupiter over $ 10^{3} $ years, compared to 53\% for the best non-stiff solvers.
+This also ensures stability in the stationary frame, with deviations of 0.76\% in asteroid separation from Jupiter over $ 10^{3} $ years, compared to 53\% for the best non-stiff solvers.
 
@@
 \subsection{Programme Structure}
-Global constants such as solar mass, and sun\textendash planet separation, along with derived values from these such as orbital period and solar radius from centre of mass, as given in an importable python module "\textit{constants}".
+Global constants such as solar mass, and sun\textendash planet separation, along with derived values from these such as orbital period and solar radius from centre of mass, are given in an importable python module "\textit{constants}".
 
 Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit (under different sampling routines) are implemented in "\textit{wander}".
-Further files then import these modules and produce the plots given in this report.
+Further files then import these modules and produce the plots given in this report, fully detailed in appendix \ref{structure}.
 
@@
 
+\subsection{Performance}
+Sampling 100 points per orbit for 100 orbits takes a mean time of 0.01723 seconds \textbf{(add error?)}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within solve\_ivp.
+
 \section{Results}
-\subsection{Orbit Stability}
+\subsection{Stability of Lagrange Points}
+Considering the Greeks' orbit without perturbations applied, it has a maximum deviation of $4.68 \times 10^{-13}$ AU over 100 orbits (1185 years).
+This value is unchanged if 1000 orbits are considered instead, confirming the stability of this Lagrange point.
+
 \subsection{Wander Analysis}
@@
 \begin{appendices}
+\section{Program Structure} \label{structure}
+Global constants, along with derived values from these are given in an importable python module "\textit{constants}".
 
-\section{Code 1} \label{Code}
+Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit are implemented in "\textit{wander}".
+
+Wander from unperturbed initial conditions is calculated within \textit{lagrange\_stability}, with ODE solver performance also evaluated here.
+
+\section{Code Listings} \label{Code}
 \begin{lstlisting}[language=Python]

### 2020-04-20 constants.py

@@ -8,8 +8,8 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 
-PRECISION = 50  # evaluation points per orbit
-ORBIT_NUM = 50  # number of orbits
+PRECISION = 100  # evaluation points per orbit
+ORBIT_NUM = 100  # number of orbits
 
 
 # DERIVED CONSTANTS
@@ -19,3 +19,4 @@ period = math.sqrt(R ** 3 / (M_S + M_P))
 omega = 2 * np.pi / period  # angular velocity of frame
 time_span = np.linspace(0, ORBIT_NUM * period, int(ORBIT_NUM * PRECISION))
 lagrange = (planet_rad - R / 2, R * math.sqrt(3) / 2, 0)
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))

### 2020-04-20 lagrange_stability.py

@@ -1,0 +1,40 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import timeit
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    solar_rad,
+    planet_rad,
+    period,
+    greek_theta,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+time_repetitions = 100
+time = timeit.timeit(
+    "orbits.rotating_frame()", globals=globals(), number=time_repetitions
+)
+print(
+    "Evaluated ODE system in "
+    + str("{0:.4g}".format(time / time_repetitions))
+    + " seconds"
+)
+
+orbit_values = orbits.rotating_frame()
+wander = np.zeros((len(orbit_values.t)))
+
+for i in range(len(orbit_values.t)):
+    wander[i] = np.linalg.norm(
+        (orbit_values.y[0:3, i] - lagrange[0:3])
+    )  # magnitude of distance from lagrange point
+
+print(
+    "The maximum wander from the Lagrange point over "
+    + str("{0:.0f}".format(ORBIT_NUM * period))
+    + " years is: "
+    + str("{0:.3g}".format(np.max(wander)))
+    + " AU"
+)

### 2020-04-20 main_file.py

@@ -22,7 +22,7 @@ plt.ticklabel_format(axis="y", style="", scilimits=None)
 plt.ylabel("Magnitude of Deviation /AU")
 plt.xlabel("Time /years")
 plt.title("Asteroid Deviation from Lagrange point")
-plt.savefig("asteroid_deviation_linear.png")
+# plt.savefig("asteroid_deviation_linear.png")
 plt.show()
 
 # MOTION IN ROTATING FRAME
@@ -113,7 +113,7 @@ plt.plot(
     ),
 )
 
-plt.title("Deviation of Greeks from Lagrnage Point")
+plt.title("Deviation of Greeks from Lagrange Point")
 plt.ylabel("Separation / AU")
 plt.xlabel("Time / years")
 plt.savefig("asteroid_deviation_stat.png")

### 2020-04-20 stability_plots.py

@@ -1,0 +1,165 @@
+import math
+
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy import integrate
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    solar_rad,
+    planet_rad,
+    period,
+    greek_theta,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+sin = math.sin(greek_theta)
+cos = math.cos(greek_theta)
+
+# ROTATING FRAME
+orbit_sol = orbits.rotating_frame()
+
+
+# DEVIATION FROM LAGRANGE POINT
+plt.plot(time_span, ((lagrange[0] - orbit_sol.y[0, :])))
+# plt.yscale("log", basey=10)
+plt.ticklabel_format(axis="y", style="sci", scilimits=None)
+plt.ylabel("Magnitude of Deviation in x direction /AU")
+plt.xlabel("Time /years")
+plt.title("Asteroid Deviation from Lagrange point")
+# plt.savefig("asteroid_deviation_linear.png")
+plt.show()
+
+
+orbit_values = orbits.rotating_frame()
+
+plt.plot(orbit_values.y[0, :] - lagrange[0], orbit_values.y[1, :] - lagrange[1])
+plt.ticklabel_format(axis="both", style="sci", scilimits=None, useOffset=False)
+plt.title("Rotating Frame just Greeks")
+plt.show()
+
+
+wander = np.zeros((len(orbit_values.t), 2))
+
+for i in range(len(wander)):
+    theta = np.arctan(
+        (orbit_values.y[1, i] - lagrange[1]) / (orbit_values.y[0, i] - lagrange[0])
+    )
+    wander[i, 0] = np.linalg.norm((orbit_values.y[0:3, i] - lagrange[0:3])) * np.abs(
+        np.cos(greek_theta - theta)
+    )  # radial component of wander
+    wander[i, 1] = np.linalg.norm((orbit_values.y[0:3, i] - lagrange[0:3])) * np.abs(
+        np.sin(greek_theta - theta)
+    )  # tangential component of wander
+
+plt.plot(time_span, wander[:, 0])
+# plt.yscale("log", basey=10)
+plt.ticklabel_format(axis="y", style="", scilimits=None)
+plt.ylabel("Radial Deviation /AU")
+plt.xlabel("Time /years")
[…]

### 2020-04-20 wander.py

@@ -9,6 +9,7 @@ from constants import (
     omega,
     time_span,
     lagrange,
+    greek_theta,
 )  # Derived constants
 
 # Initial Conditions
@@ -16,8 +17,6 @@ initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotati
 initial_cond_stat = np.array(
     (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
 )
-
-greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 
 
 # Defined Functions

### 2020-04-20 Report/draft1.tex
lagrange_stability completed

@@
 \subsection{Performance}
-Sampling 100 points per orbit for 100 orbits takes a mean time of 0.01723 seconds \textbf{(add error?)}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within solve\_ivp.
+Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second} \textbf{(add error?)}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within solve\_ivp.
 
@@
 This value is unchanged if 1000 orbits are considered instead, confirming the stability of this Lagrange point.
+
+In the stationary frame, this wander from the (now moving) Lagrange point is depicted in Figure \ref{fig:greeksdeviationstationaryframe}.
+The deviation oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, and a period of $148.2 \pm 0.8$ years, where the error was determined by the width of the Fourier peak.
+This is modulated with a faster oscillation component of $11.85 \pm 0.27$ years, equivalent to the orbital period of the asteroids and clearly demonstrating an periodic error in the integration routine, becoming more significant over time.
+
+These much more significant errors are due to \textbf{YYY}.
+Energy can also be evaluated, and conserved, in this inertial frame; asteroid specific energy varies within only 0.113\% of the initial value, and with a similar periodicity to wander.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{Figures/greeks_deviation_stationary_frame}
+	\caption[]{The wander of Greek asteroids from the Lagrange point in the stationary frame.
+Note the two oscillation components and constant maximum oscillation amplitude over time, demonstrating this point is indeed stable.
+}
+	\label{fig:greeksdeviationstationaryframe}
+\end{figure}
+
+Animations show orbit....
 

### 2020-04-20 lagrange_stability.py

@@ -1,9 +1,10 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import timeit
+import scipy.signal
 
 import orbits
-from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import G, M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import (
     solar_rad,
     planet_rad,
@@ -12,6 +13,8 @@ from constants import (
     time_span,
     lagrange,
 )  # Derived constants
+
+# ROTATING FRAME
 
 time_repetitions = 100
 time = timeit.timeit(
@@ -32,9 +35,77 @@ for i in range(len(orbit_values.t)):
     )  # magnitude of distance from lagrange point
 
 print(
-    "The maximum wander from the Lagrange point over "
+    "Maximum wander from the Lagrange point in the rotating frame over "
     + str("{0:.0f}".format(ORBIT_NUM * period))
-    + " years is: "
+    + " years: "
     + str("{0:.3g}".format(np.max(wander)))
     + " AU"
+)
+
+# # STATIONARY FRAME
+
+orbit_values = orbits.stationary_frame()
+wander = np.zeros((len(orbit_values.t)))
+
+for i in range(len(orbit_values.t)):
+    wander[i] = np.linalg.norm(
+        (orbit_values.y[0:3, i] - orbits.lagrange_pos(orbit_values.t[i])[0:3])
+    )  # magnitude of distance from lagrange point
+
+# wander = np.linalg.norm((orbit_values.y[0:3] - orbits.lagrange_pos(time_span)[0:3]))
+
+print(
+    "Maximum wander from the Lagrange point in the stationary frame over "
+    + str("{0:.0f}".format(ORBIT_NUM * period))
+    + " years: "
+    + str("{0:.3g}".format(np.max(wander)))
+    + " AU"
+)
+
+plt.plot(time_span, wander, linewidth=0.8)
+plt.title("Wander of Greeks from Unperturbed State")
+plt.ylabel("Deviation of Greeks from Lagrange Point /AU")
+plt.xlabel("Time /years")
+plt.savefig("greeks_deviation_stationary_frame.png")
+plt.show()
+
+sampling_rate = PRECISION / period
+freqs = np.fft.fftfreq(len(wander)) * sampling_rate
[…]

### 2020-04-20 orbits.py

@@ -46,6 +46,10 @@ def planet_pos(t=0):
 
 
 def lagrange_pos(t=0):
+    """Position of Greeks' lagrange point in 3-dimensions at time t
+
+    Note that z coordinate is fixed to 0, 
+    so movement is constrained to xy plane in this exact solution"""
     return np.array(
         [
             greek_rad * np.cos(omega * t + greek_theta),

### 2020-04-21 Report/draft1.tex
corrected lagrange stability added energy function into orbits

@@
 
-\subsection{Theoretical Model}
+\subsection{Theoretical Model} \label{theory}
 The three body problem, where the dynamics of three interacting bodies are determined given their initial positions and velocities, has no analytical (closed-form) solution in the general case \cite{Barrow2008}.
@@
 As there is 1:1 orbital resonance between Jupiter and the asteroids, all three bodies are stationary in this frame.
-This significantly increases the accuracy of numerical simulations, as \textbf{WHY}
+This significantly increases the accuracy of numerical simulations, as the exact solution is stationary with no explicit time dependence, rather then requiring an infinite power series \cite{Guglielmi2001}, \cite{LeVeque2007}.
 
@@
 %Then all theory could be in methodology?
-Within solve\_ivp, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) however this is non-stiff, giving a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
-This is larger than expected, suggesting the system of equations requires an unreasonable small step size for  numerically stability with respect to this numerical method, even regions where the solution curve is smooth \cite{Lambert1991}.
-This suggests the system is stiff, and solvers designed for this typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to the non-stiff solvers \cite{Byrne1987}.
+Within solve\_ivp, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) and gives a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
+This is larger than expected, indicating the system of equations requires an unreasonably small step size for numerical stability with respect to this method, even in regions where the solution curve is smooth \cite{Lambert1991}.
+This suggests the system is stiff, and solvers designed for this typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to the non-stiff solvers such as RK45 \cite{Byrne1987}.
 
@@
 \subsection{Performance}
-Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second} \textbf{(add error?)}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within solve\_ivp.
+Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within solve\_ivp.
 
@@
 
-These much more significant errors are due to \textbf{YYY}.
-Energy can also be evaluated, and conserved, in this inertial frame; asteroid specific energy varies within only 0.113\% of the initial value, and with a similar periodicity to wander.
-
 \begin{figure}
@@
 	\caption[]{The wander of Greek asteroids from the Lagrange point in the stationary frame.
-Note the two oscillation components and constant maximum oscillation amplitude over time, demonstrating this point is indeed stable.
-}
+Note the two oscillation components and constant maximum oscillation amplitude over time, demonstrating this point is indeed stable.}
 	\label{fig:greeksdeviationstationaryframe}
 \end{figure}
+
+These much more significant errors are due to time dependence in the exact solution, as detailed in Section \ref{theory}.
+Energy can also be evaluated, and conserved, in this inertial frame; asteroid specific energy varies within only 0.113\% of the initial value, and with a similar periodicity to wander.
 

### 2020-04-21 lagrange_stability.py

@@ -9,6 +9,7 @@ from constants import (
     solar_rad,
     planet_rad,
     period,
+    omega,
     greek_theta,
     time_span,
     lagrange,
@@ -52,8 +53,6 @@ for i in range(len(orbit_values.t)):
         (orbit_values.y[0:3, i] - orbits.lagrange_pos(orbit_values.t[i])[0:3])
     )  # magnitude of distance from lagrange point
 
-# wander = np.linalg.norm((orbit_values.y[0:3] - orbits.lagrange_pos(time_span)[0:3]))
-
 print(
     "Maximum wander from the Lagrange point in the stationary frame over "
     + str("{0:.0f}".format(ORBIT_NUM * period))
@@ -92,12 +91,11 @@ print(np.abs(peak_freq).round(2)[0:5])
 
 # ENERGY CONSERVATION
 
-reduced_mass = (M_S * M_P) + (M_S + M_P)
-radius = np.linalg.norm(orbit_values.y[0:3], axis=0)
-velocity = np.linalg.norm(orbit_values.y[3:], axis=0)
-energy = -G * reduced_mass / radius + 1 / 2 * velocity ** 2
+energy = np.zeros(len(time_span))
+for i in range(len(time_span)):
+    energy[i] = orbits.specific_energy(time_span[i], orbit_values.y[:, i])
 
-plt.plot(time_span, energy - energy[0], linewidth=0.5)
+plt.plot(time_span, energy[:] - energy[0], linewidth=0.5)
 plt.title("Energy Conservation in the Stationary Frame")
 plt.xlabel("Time /years")
 plt.ylabel("Specific Energy offset from initial value(J/kg)")

### 2020-04-21 orbits.py

@@ -144,7 +144,7 @@ def stat_derivatives(t, y):
 
 
 def stationary_frame(y0_stat=(rcos, rsin, 0, -vsin, vcos, 0)):
-    """Gives position and vlocity of asteroids in the stationary frame
+    """Gives position and velocity of asteroids in the stationary frame
 
     Uses scipy solve_ivp method with LSODA, taking an input in the form
     of (x_pos, y_pos, z_pos, x_vel, y_vel, z_vel) for the initial state.
@@ -157,3 +157,17 @@ def stationary_frame(y0_stat=(rcos, rsin, 0, -vsin, vcos, 0)):
         y0=y0_stat,
         t_eval=time_span,
     )
+
+
+def specific_energy(t, y):
+    """ Gives energy per unit mass of asteroids
+
+    t is the time after the start of the orbit
+    y has 6 components; first three for position, second three for velocity
+    These are in the form of the output of a scipy intergrate function
+    """
+
+    radius_p = np.linalg.norm(y[0:3] - planet_pos(t), axis=0)
+    radius_s = np.linalg.norm(y[0:3] - solar_pos(t), axis=0)
+    velocity = np.linalg.norm(y[3:], axis=0)
+    return -G * M_P / radius_p - G * M_S / radius_s + 1 / 2 * velocity ** 2

### 2020-04-22 Report/draft1.tex
tadpole orbits with associated fourier analysis implemented in perturbations_2D

@@
 	%\authorstyle{Christopher Gallagher}}
-\author{\authorstyle{Christopher Gallagher} 
+\author{\authorstyle{Christopher Gallagher} %Candidate 6952R
 	\institution{University of Cambridge}}
@@
 
-\subsection{Lagrange Points}
+\subsection{Lagrange Points} \label{lagrange}
 The asteroids exist at/near Lagrange points, defined in Lagrange's initial analysis of the three-body problem in 1772 \cite{Lagrange1772}, were he demonstrated the existence of five equilibrium points for an object of negligible mass orbiting under the gravitational effect of two larger masses.
@@
 
-While orbits between Lagrange points are also possible, this report will focus on the tadpole orbits observed as asteroids deviate from L4 and L5 \cite{Murray1999}. \textit{More detail?}
+Figure \ref{fig:lagrangepoints} depicts orbits about L\subscript{4} and L\subscript{5} (known as tadpole orbits), as well as orbits between Lagrange points are also possible (such as horseshoe orbits), described by Murray et al.
+\cite{Murray1999}.
+
+Horseshoe taylor \cite{Taylor1981}
+
+This report will focus on tadpole orbits, a well documented feature of Trojan orbits \cite{Garfinkel1983}, \cite{Dermott1981}, and the restricted three-body problem in general.
+Their distinctive shape is result of a long-period motion about the equilibrium point combining with a short-period oscillation due to Keplerian motion of the asteroid.
+
+Szebehely et al. \cite{Szebehely1969} predict this short period tends to the planetary period in the small planetary mass limit, while the long period is given by
+
+\begin{equation}
+T_{long} = T_{P} \sqrt{\frac{4}{27 \mu_{2}}},  
+\end{equation}
+
+where $\mu_{2} = m_{2} / (m_{1} + m_{2})$ and $ T_{P} $ is the period of planetary orbit.
 
@@
 
-In this report, I will consider the circular, restricted, three-body problem, where two of the bodies move in circular, coplanar orbits about their common centre of mass, unaffected by the negligible mass of the third body.
+In this report, I will consider the circular, restricted, three-body problem, where two of the bodies move in circular, coplanar orbits about their common centre of mass (CoM), unaffected by the negligible mass of the third body.
 I will also assume that all interactions are via Newtonian gravity.
@@
 
-\subsubsection{Assumptions - FIX}
-\begin{itemize}
-	\item Circular orbit
-	\item Constant Jupiter-Sun separation
-	\item Coplanar orbits
-	\item Negligible asteroid mass
-	\item Newtonian gravity
-\end{itemize}
-
-I also applied the following symmetries:
-trojan and greek symmetry (analysis focused on greeks)
-rotational symmetry - arbitrary initial point 
-direction of orbit
-"The combination of these symmetries allows the problem to be simplified. Such that
-only the Greeks, orbiting counter-clockwise, with perturbations at t = 0, need to be
-investigated.
-"
-(This reduces computing time not hvaing to solve for trojans as well)
+\subsection{Symmetry}
+This problem contains a number of symmetries, which could be employed to simplify the problem and reduce the computational load.
+The Trojan and Greek asteroids are in equivalent positions, so experience the same forces, and the system is rotationally and inversion symmetric, so the choice of initial point and orbit direction is arbitrary.
+Therefore only the Greeks, orbiting counter-clockwise with perturbations applied at t = 0, need to be investigated.
 
@@
 
-Furthermore, the Lagrange point in Cartesian coordinates based about the centre of mass is easily found to be:
+Furthermore, the Lagrange point in Cartesian coordinates based about the CoM is easily found to be:
 
@@
 \subsection{Programme Structure}
-Global constants such as solar mass, and sun\textendash planet separation, along with derived values from these such as orbital period and solar radius from centre of mass, are given in an importable python module "\textit{constants}".
+Global constants such as solar mass, and sun\textendash planet separation, along with derived values from these such as orbital period and solar radius from the CoM, are given in an importable python module "\textit{constants}".
 
@@
 In the stationary frame, this wander from the (now moving) Lagrange point is depicted in Figure \ref{fig:greeksdeviationstationaryframe}.
-The deviation oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, and a period of $148.2 \pm 0.8$ years, where the error was determined by the width of the Fourier peak.
-This is modulated with a faster oscillation component of $11.85 \pm 0.27$ years, equivalent to the orbital period of the asteroids and clearly demonstrating an periodic error in the integration routine, becoming more significant over time.
+The deviation oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, and a period of $148 \pm 4$ years, where the error was determined by the width of the Fourier peak.
+This is modulated with a faster oscillation component of $11.85 \pm 0.27$ years, equivalent to the orbital period of the asteroids.
 
@@
 	\includegraphics[width=\linewidth]{Figures/greeks_deviation_stationary_frame}
-	\caption[]{The wander of Greek asteroids from the Lagrange point in the stationary frame.
+	\caption{The wander of Greek asteroids from the Lagrange point in the stationary frame.
 Note the two oscillation components and constant maximum oscillation amplitude over time, demonstrating this point is indeed stable.}
@@
 Energy can also be evaluated, and conserved, in this inertial frame; asteroid specific energy varies within only 0.113\% of the initial value, and with a similar periodicity to wander.
+Energy is also negative, confirming the asteroids are located within a bound orbit.
 
-Animations show orbit....
+Animations produced to demonstrate the orbit in the stationary frame are included in Supplementary Material I-II.
+Animation I depicts the orbit as evaluated by the radau solver, while II depicts it with LSODA, demonstrating the drift present over time with non-stiff solvers.
 
 \subsection{Wander Analysis}
-\subsubsection{Perturbations in z-direction}
+The wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin, and considered  separately perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively).
+By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).
+
+\begin{figure}[ht]
+	\centering
+	\begin{subfigure}{.45\textwidth}
+		\centering
+		% include first image
+		\includegraphics[width=\linewidth]{Figures/wanderagainstradialpertubation2}  
+		\caption{Radial}
+		\label{fig:wander_rad}
+	\end{subfigure}
+	\hfill %% useful if width of each figure is less the .5\textwidth
+	\begin{subfigure}{.45\textwidth}
+		\centering
+		% include second image
+		\includegraphics[width=\linewidth]{Figures/wanderagainsttangentialpertubation2}  
+		\caption{Tangential}
+		\label{fig:wander_tan}
+	\end{subfigure}
+	\caption{Wander from an initial perturbation with maximum relative magnitude of 0.1\% of the displacement from the origin, given in radial and tangential directions.}
+	\label{fig:wanderplots}
+\end{figure}
+
+Figure \ref{fig:wander_rad} demonstrates a polynomial dependence on perturbation size in the radial direction, with a negligible constant term.
+While a quadratic has been fitted here, it was not possible to eliminate the possibility of higher order terms, as increasing perturbation size beyond 0.06 AU can lead to unstable orbits.
+
+We may consider the wander resulting from perturbations in position and velocity space.
+Figure \ref{fig:position_mesh} clearly shows the dependence on radial position perturbations only, while Figure \ref{fig:velocity_mesh} shows a similar the dependence on tangential velocity perturbations.
+
+\begin{figure}[ht]
+	\centering
+	\begin{subfigure}{.45\textwidth}
+		\centering
+		% include first image
+		\includegraphics[width=\linewidth]{Figures/testcolourmesh6}  
+		\caption{Position}
+		\label{fig:position_mesh}
+	\end{subfigure}
+	\hfill %% useful if width of each figure is less the .5\textwidth
+	\begin{subfigure}{.45\textwidth}
+		\centering
+		% include second image
+		\includegraphics[width=\linewidth]{Figures/testvelocitymesh2}  
+		\caption{Velocity}
+		\label{fig:velocity_mesh}
+	\end{subfigure}
+	\caption{Wander from an initial point in position and velocity space, only calculated over 50 orbits with 30 points per orbit to reduce computational load.
+Note the dependence on radial position and tangential velocity perturbation components.
+}
+	\label{fig:mesh_plots}
+\end{figure}
+
+\subsection{Orbit Types}
+Figure \ref{fig:orbitplots} displays orbits resulting from small perturbations in the radial and tangential directions.
+
+\begin{figure}[ht]
+	\centering
+	\begin{subfigure}{.23\textwidth}
+		\centering
+		% include first image
+		\includegraphics[width=\linewidth]{Figures/radialp_orbits}  
+		\caption{Radial}
+		\label{fig:orbit_rad}
+	\end{subfigure}
+	\hfill %% useful if width of each figure is less the .5\textwidth
+	\begin{subfigure}{.23\textwidth}
+		\centering
+		% include second image
+		\includegraphics[width=\linewidth]{Figures/tangentialp_orbits}  
+		\caption{Tangential}
+		\label{fig:orbit_tan}
+	\end{subfigure}
+	\caption{Orbits from an initial perturbation of magnitude 0.01 AU from the origin, in radial and tangential directions, over 12 orbital periods.
+Note the 'tadpole' features resulting from the radial perturbation, and the significantly larger wander than from the tangential perturbation.}
+	\label{fig:orbitplots}
+\end{figure}
+
+The "tadpole" orbit in Figure \ref{fig:orbit_rad} consists of two oscillating components, as described in Section \ref{lagrange}.
+The short period, measured at $11.85 \pm 0.61$ years, is in agreement with  planetary orbital period as expected; meanwhile the long period is measured to be $148 \pm 5$ years, in agreement with the analytical value of $144$ years.
+
+For larger perturbations, horseshoe orbits may also be observed...
+
+\subsection{Perturbations in z-direction}
 
@@
 
-%Talk about animations here too? or after solver method
+\clearpage
 \printbibliography
@@
 \section{Program Structure} \label{structure}
-Global constants, along with derived values from these are given in an importable python module "\textit{constants}".
+Global constants, along with derived values from these are given in an python module "\textit{constants}", from which relevant variable are imported in all other scripts.
 
 Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit are implemented in "\textit{wander}".
+These modules are both imported into the scripts below for plotting and analysis, and given in the listings in Appendix \ref{Code}.
 
-Wander from unperturbed initial conditions is calculated within \textit{lagrange\_stability}, with ODE solver performance also evaluated here.
+Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
+Orbits in the stationary frame are animated and saved within "\textit{stat\_frame\_video}".
+
+2D perturbations are characterised in \textit{perturbation\_2D}.
 

### 2020-04-22 Reference Scripts/wander_full.py

@@ -1,0 +1,288 @@
+import math
+import matplotlib.pyplot as plt
+import numpy as np
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    solar_rad,
+    omega,
+    time_span,
+    lagrange,
+    greek_theta,
+)  # Derived constants
+
+# Initial Conditions
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+initial_cond_stat = np.array(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+
+
+# Defined Functions
+def perturb(initial_cond, max_pertubation_size=0.01):
+    """ Returns perturbed version of initial conditions array
+
+    init_cond is the initial conditions, submitted as a numpy array
+    pertubation_size is the relative magnitude of the perturbation, measured as 
+    the percentage deivation from the lagrange point (for position)
+    Note that this will give a different value each time it is run
+    """
+    # return data + data * (
+    #     max_pertubation_size * np.random.uniform(-1, 1, (np.shape(data)))
+    # )  # random between -1 and 1, perturbs position only for initial_cond_rot
+    # theta = np.random.uniform(0, 2 * math.pi)
+    # return data + data * (
+    #     max_pertubation_size
+    #     * np.random.rand()
+    #     * np.array((math.cos(theta), math.sin(theta), 0, 0, 0, 0))
+    # )  # random betweeon -1 and 1, perturbs position only for initial_cond_rot
+    rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+    while np.linalg.norm(rand_array[0:2]) > 1:
+        rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+    return initial_cond + initial_cond * (
+        max_pertubation_size * rand_array
+    )  # random between -1 and 1, selects points in circle of radius 1
+
+
+def max_wander(max_pertubation_size, samples):
+    """Returns maximum deviation over orbit in the rotating frame for given number of samples
+
+    This calculates the distange from the initial point (not the lagrange point)
[…]

### 2020-04-22 constants.py

@@ -8,8 +8,8 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 
-PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 100  # number of orbits
+PRECISION = 10  # evaluation points per orbit
+ORBIT_NUM = 10  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-22 lagrange_stability.py

@@ -75,7 +75,7 @@ fourier = np.abs(np.fft.fft(wander))
 plt.plot(freqs, fourier)
 plt.title("Fourier Transform of Wander")
 plt.ylabel("'Frequency Domain (Spectrum) Magnitude'")
-plt.xlabel("Frequency /Hz")
+plt.xlabel("Frequency (1/year)")
 plt.xlim(0, sampling_rate / 200)  # select low freq region of data
 plt.show()
 
@@ -86,8 +86,8 @@ peak_freq = np.zeros(len(peaks))
 for i in range(len(peaks)):
     peak_freq[i] = 1 / freqs[peaks[i]]
 print("Period (in years) of primary frequency components in Fourier spectrum: ")
-print(np.abs(peak_freq).round(2)[0:5])
-
+print(np.unique(np.abs(peak_freq)).round(2))
+# Evaluated for 100 points per orbit, 100 orbits
 
 # ENERGY CONSERVATION
 

### 2020-04-22 perturbations_2D.py

@@ -1,0 +1,260 @@
+import math
+import numpy as np
+import scipy.signal
+import matplotlib.pyplot as plt
+
+import orbits
+import wander
+from constants import R, PRECISION  # User defined constants
+from constants import (
+    solar_rad,
+    omega,
+    period,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+# INITIAL CONDITIONS
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+initial_cond_stat = np.array(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+cos = math.cos(greek_theta)
+sin = math.sin(greek_theta)
+
+# RANDOM SAMPLING OF WANDER
+
+wander_data = wander.rand_sample(max_pertubation_size=0.01, samples=20)
+
+fig = plt.figure()
+ax = fig.add_subplot()
+
+coeff = np.polyfit(
+    wander_data[:, 0], wander_data[:, 2], 2
+)  # for quadratic line of best fit
+best_fit = np.poly1d(coeff)
+order = np.argsort(wander_data[:, 0])
+plt.plot(
+    wander_data[order, 0], best_fit(wander_data[order, 0]), linewidth=0.75,
+)
+plt.plot(
+    wander_data[:, 0], wander_data[:, 2], linestyle="None", marker="+", color="navy"
+)
+
+plt.title("Wander under Radial Pertubations")
+plt.ylabel("Wander /AU")
+plt.xlabel("Radial Pertubation Size /AU")
+plt.text(
+    0.02,
+    0.95,
+    "Line of Best fit: "
+    + str("{0:.2f}".format(coeff[0]))
+    + "x\u00b2 "
+    + str("{0:+.2f}".format(coeff[1]))
+    + "x "
+    + str("{0:+.2f}".format(coeff[2])),
+    transform=ax.transAxes,
+    fontsize=9.5,
+)
+plt.savefig("wanderagainstradialpertubation2.png")
+plt.show()
+
+
+plt.plot(
+    wander_data[:, 1], wander_data[:, 2], linestyle="None", marker="x", color="navy"
+)
+plt.title("Wander under Tangential Pertubations")
[…]

### 2020-04-22 stat_frame_video.py

@@ -7,34 +7,31 @@ import orbits
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import lagrange, omega, time_span  # Derived constants
 
-SAMPLE = int(100)  # Sample every n points
-
 # Set up orbits for each body
 orbit_sun = orbits.solar_pos(time_span)
 orbit_jupiter = orbits.planet_pos(time_span)
-orbit_lagrange = orbits.lagrange_pos(time_span)
 orbit_greeks = orbits.stationary_frame(
     (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
 )
 orbit_trojans = orbits.stationary_frame(
     (lagrange[0], -lagrange[1], 0, omega * lagrange[1], omega * lagrange[0], 0)
 )
-orbit_list = [orbit_sun, orbit_jupiter, orbit_lagrange, orbit_greeks, orbit_trojans]
-bodies = ["Sun", "Jupiter", "Lagrange", "Greeks", "Trojans"]
+orbit_list = [orbit_sun, orbit_jupiter, orbit_greeks, orbit_trojans]
+bodies = ["Sun", "Jupiter", "Greeks", "Trojans"]
 N = len(bodies)
 
 orbit_data = np.zeros(
-    (len(orbit_trojans.t[0::SAMPLE]), 2, int(N))
+    (len(orbit_trojans.t), 2, int(N))
 )  # select time coordinate, spatial coordinate (x/y), and index of body in list 'bodies'
 
 
 for i in range(N):
     if type(orbit_list[i]) is np.ndarray:  # Different format of orbit results
-        orbit_data[:, 0, i] = orbit_list[i][0][0::SAMPLE]
-        orbit_data[:, 1, i] = orbit_list[i][1][0::SAMPLE]
+        orbit_data[:, 0, i] = orbit_list[i][0]
+        orbit_data[:, 1, i] = orbit_list[i][1]
     else:
-        orbit_data[:, 0, i] = orbit_list[i].y[0, 0::SAMPLE]
-        orbit_data[:, 1, i] = orbit_list[i].y[1, 0::SAMPLE]
+        orbit_data[:, 0, i] = orbit_list[i].y[0, :]
+        orbit_data[:, 1, i] = orbit_list[i].y[1, :]
 
 
 fig = plt.figure(figsize=(6, 6))
@@ -42,16 +39,15 @@ ax = plt.axes(xlim=(-8, 8), ylim=(-8, 8))
[…]

### 2020-04-22 wander.py

@@ -1,16 +1,7 @@
-import math
-import matplotlib.pyplot as plt
 import numpy as np
 
 import orbits
-from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
-from constants import (
-    solar_rad,
-    omega,
-    time_span,
-    lagrange,
-    greek_theta,
-)  # Derived constants
+from constants import omega, lagrange, greek_theta  # Derived constants
 
 # Initial Conditions
 initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
@@ -27,16 +18,10 @@ def perturb(initial_cond, max_pertubation_size=0.01):
     pertubation_size is the relative magnitude of the perturbation, measured as 
     the percentage deivation from the lagrange point (for position)
     Note that this will give a different value each time it is run
+
+    Perturbations are chosen randomly within a circle of radius one, 
+    to give a more uniform radial distribution than from a square.
     """
-    # return data + data * (
-    #     max_pertubation_size * np.random.uniform(-1, 1, (np.shape(data)))
-    # )  # random between -1 and 1, perturbs position only for initial_cond_rot
-    # theta = np.random.uniform(0, 2 * math.pi)
-    # return data + data * (
-    #     max_pertubation_size
-    #     * np.random.rand()
-    #     * np.array((math.cos(theta), math.sin(theta), 0, 0, 0, 0))
-    # )  # random betweeon -1 and 1, perturbs position only for initial_cond_rot
     rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
     while np.linalg.norm(rand_array[0:2]) > 1:
         rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
@@ -45,15 +30,15 @@ def perturb(initial_cond, max_pertubation_size=0.01):
     )  # random between -1 and 1, selects points in circle of radius 1
 
 
-def max_wander(max_pertubation_size, samples):
-    """Returns maximum deviation over orbit in the rotating frame for given number of samples
+def rand_sample(max_pertubation_size, samples):
+    """Returns maximum deviation over orbit in the rotating frame for given number of random samples
 
[…]

### 2020-04-22 wander_orbits.py

@@ -1,6 +1,7 @@
 import math
 import matplotlib.pyplot as plt
 import numpy as np
+import scipy.signal
 
 import orbits
 
@@ -10,6 +11,7 @@ from constants import (
     solar_rad,
     planet_rad,
     omega,
+    period,
     time_span,
     lagrange,
 )  # Derived constants
@@ -53,46 +55,137 @@ def wander(pertubation, samples=1, pertubation_type="position"):
 
 
 # ROTATING FRAME
-# orbit_sol = orbits.rotating_frame(
-#     (initial_cond_rot + (0, 0, 0, 0, 0, 0))
-# )  # no perturbation
-# orbit_sol = orbits.rotating_frame(
-#     (initial_cond_rot + 0.01 * np.array((cos, sin, 0, 0, 0, 0)))
-# )  # radial perturbation GIVES TADPOLES
 orbit_sol = orbits.rotating_frame(
+    (initial_cond_rot + (0, 0, 0, 0, 0, 0))
+)  # no perturbation
+orbit_sol_rad = orbits.rotating_frame(
+    (initial_cond_rot + 0.01 * np.array((cos, sin, 0, 0, 0, 0)))
+)  # radial perturbation GIVES TADPOLES
+orbit_sol_tan = orbits.rotating_frame(
     (initial_cond_rot + 0.01 * np.array((sin, -cos, 0, 0, 0, 0)))
 )  # tangential perturbation   GIVES ORBITS
 
+orbit_sol_rad2 = orbits.rotating_frame(
+    (initial_cond_rot + 0.07 * np.array((cos, sin, 0, 0, 0, 0)))
+)  # radial perturbation GIVES TADPOLES
+
+
+orbit_sol_rad2 = orbits.rotating_frame(
+    np.array((-1.02745, 0, 0, 0.04032, 0, 0))
+)  # radial perturbation GIVES TADPOLES
+
+
+# FOURIER ANALYSIS OF POLAR ANGLE
+# angle = np.zeros_like(time_span)
+# for i in range(len(time_span)):
+#     angle[i] = np.arctan(orbit_sol_rad.y[1, i] / orbit_sol_rad.y[0, i])
+# plt.plot(time_span, angle)
+# plt.ticklabel_format(axis="y", style="", scilimits=None)
+# plt.ylabel("Polar Angle from CoM /rad")
+# plt.xlabel("Time /years")
+# plt.title("Asteroid Deviation from Lagrange point")
+# plt.show()
+
[…]

### 2020-04-23 Report/draft1.tex
final 2d_perturbations

@@
 \cite{Murray1999}.
-
-Horseshoe taylor \cite{Taylor1981}
 
@@
 The "tadpole" orbit in Figure \ref{fig:orbit_rad} consists of two oscillating components, as described in Section \ref{lagrange}.
-The short period, measured at $11.85 \pm 0.61$ years, is in agreement with  planetary orbital period as expected; meanwhile the long period is measured to be $148 \pm 5$ years, in agreement with the analytical value of $144$ years.
+The short period, measured at $11.85 \pm 0.61$ years, is in excellent agreement with  planetary orbital period as expected; meanwhile the long period is measured to be $148 \pm 5$ years, consistent with the analytical value of $144$ years.
 
-For larger perturbations, horseshoe orbits may also be observed...
+For a narrow range of larger perturbations, stable horseshoe orbits encompassing both Lagrange points may also be observed as depicted in Figure \ref{fig:horseshoe}.
+This has a full period of $353$ years, in agreement with Taylor et al.'s numerical result of $358 \pm 7$ years \cite{Taylor1981}.
+
+%smooth families of horseshoe-shaped longperiodic orbits for alpha2 > 1.
+Smooth, because they are restricted to shallow resonance, where the evolution of the epicyclic cusps and loops cannot be seen.
+- according to taylor, but havn't found smooth orbits here
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/horseshoe}
+	\caption{Horseshoe and tadpole orbits in the rotating frame, from radial perturbations of 0.07 and 0.01 AU respectively, over 30 orbital periods.}
+	\label{fig:horseshoe}
+\end{figure}
 
 \subsection{Perturbations in z-direction}
+
+\subsection{Variation of Planetary Mass}
 

### 2020-04-23 constants.py

@@ -8,8 +8,8 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 
-PRECISION = 10  # evaluation points per orbit
-ORBIT_NUM = 10  # number of orbits
+PRECISION = 1000  # evaluation points per orbit
+ORBIT_NUM = 40  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-23 perturbations_2D.py

@@ -162,12 +162,60 @@ orbit_sol = orbits.rotating_frame(
 )  # no perturbation
 orbit_sol_rad = orbits.rotating_frame(
     (initial_cond_rot + 0.01 * np.array((cos, sin, 0, 0, 0, 0)))
-)  # radial perturbation GIVES TADPOLES
+)  # radial perturbation
+orbit_sol_large_rad = orbits.rotating_frame(
+    (initial_cond_rot + 0.07 * np.array((cos, sin, 0, 0, 0, 0)))
+)  # large radial perturbation
 orbit_sol_tan = orbits.rotating_frame(
     (initial_cond_rot + 0.01 * np.array((sin, -cos, 0, 0, 0, 0)))
-)  # tangential perturbation   GIVES ORBITS
+)  # tangential perturbation
+
+# PLOT OF TADPOLE AND HORSESHOE ORBITS
+plt.plot(orbit_sol_large_rad.y[0, :], orbit_sol_large_rad.y[1, :], label="Horseshoe")
+plt.plot(orbit_sol_rad.y[0, :], orbit_sol_rad.y[1, :], label="Tadpole")
+
+plt.plot(
+    lagrange[0],
+    lagrange[1],
+    marker="+",
+    markersize=8,
+    color="black",
+    linestyle="",
+    label="Lagrange",
+)
+
+plt.plot(
+    lagrange[0], -lagrange[1], marker="+", markersize=8, color="black", linestyle="",
+)
+
+plt.plot(
+    orbits.solar_pos(0)[0],
+    0,
+    label="Sun",
+    color="yellow",
+    markersize=18,
+    marker="o",
+    linestyle="None",
+)
+plt.plot(
+    orbits.planet_pos(0)[0],
+    0,
+    label="Jupiter",
+    color="red",
+    markersize=9,
+    marker="o",
+    linestyle="None",
+)
+plt.title("Tadpole and Horseshoe Orbits in the Rotating Frame")
+plt.xlabel("X Position /AU")
+plt.ylabel("Y Position /AU")
+plt.legend(bbox_to_anchor=(0, 0, 1, 0.2), loc="lower left", mode="expand", ncol=5)
+plt.savefig("horseshoe.png")
+plt.show()
+
 
 # FOURIER ANALYSIS OF POLAR ANGLE
+# Consider polar angle from the centre of mass
 
 angle = np.zeros_like(time_span)
 for i in range(len(time_span)):
[…]

### 2020-04-23 wander_orbits.py

@@ -70,11 +70,6 @@ orbit_sol_rad2 = orbits.rotating_frame(
 )  # radial perturbation GIVES TADPOLES
 
 
-orbit_sol_rad2 = orbits.rotating_frame(
-    np.array((-1.02745, 0, 0, 0.04032, 0, 0))
-)  # radial perturbation GIVES TADPOLES
-
-
 # FOURIER ANALYSIS OF POLAR ANGLE
 # angle = np.zeros_like(time_span)
 # for i in range(len(time_span)):
@@ -140,8 +135,17 @@ orbit_sol_rad2 = orbits.rotating_frame(
 # plt.savefig("tangentialp_orbits.png")
 # plt.show()
 
-plt.plot(orbit_sol_rad.y[0, :], orbit_sol_rad.y[1, :], label="Tadpole")
 plt.plot(orbit_sol_rad2.y[0, :], orbit_sol_rad2.y[1, :], label="Horseshoe")
+plt.plot(orbit_sol_rad.y[0, :], orbit_sol_rad.y[1, :], label="Tadpole")
+plt.plot(
+    orbit_sol_rad2.y[0, -1],
+    orbit_sol_rad2.y[1, -1],
+    marker="o",
+    markersize=10,
+    color="pink",
+    linestyle="",
+)
+
 # plt.plot(
 #     initial_cond_rot[0] + 0.01 * cos,
 #     initial_cond_rot[1] + 0.01 * sin,
@@ -158,19 +162,20 @@ plt.plot(
     markersize=8,
     color="black",
     linestyle="",
-    label="Lagrange Points",
+    label="Lagrange",
 )
 
 plt.plot(
     lagrange[0], -lagrange[1], marker="+", markersize=8, color="black", linestyle="",
 )
 
+
 plt.plot(
     orbits.solar_pos(0)[0],
     0,
     label="Sun",
     color="yellow",
-    markersize=12,
+    markersize=18,
     marker="o",
     linestyle="None",
 )
@@ -179,13 +184,14 @@ plt.plot(
     0,
     label="Jupiter",
     color="red",
-    markersize=8,
+    markersize=9,
     marker="o",
     linestyle="None",
 )
 plt.title("Tadpole and Horseshoe Orbits in the Rotating Frame")
 plt.xlabel("X Position /AU")
 plt.ylabel("Y Position /AU")
-plt.legend()
+# plt.legend()
[…]

### 2020-04-24 Report/draft1.tex
Basic 3D perturbations before implementing into wander func

@@
 
-\subsection{Performance}
+\subsection{Performance} \label{Performance}
 Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within solve\_ivp.
@@
 \section{Results}
-\subsection{Stability of Lagrange Points}
+\subsection{Unperturbed Stability of Lagrange Points} \label{unperturbed}
 Considering the Greeks' orbit without perturbations applied, it has a maximum deviation of $4.68 \times 10^{-13}$ AU over 100 orbits (1185 years).
@@
 
-\subsection{Wander Analysis}
+\subsection{Wander Analysis} \label{perturbed}
 The wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin, and considered  separately perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively).
@@
 
-\subsection{Orbit Types}
+\subsection{Orbit Types} \label{orbits}
 Figure \ref{fig:orbitplots} displays orbits resulting from small perturbations in the radial and tangential directions.
@@
 
-\subsection{Perturbations in z-direction}
+\subsection{Perturbations in z-direction} \label{3D}
 
-\subsection{Variation of Planetary Mass}
+\subsection{Variation of Planetary Mass} \label{planet}
+The variation of wander with planetary mass is given in Figure \ref{fig:planetmass}.
 
-\section{Discussion}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p5b}
+	\caption{Variation of maximum wander with planetary mass. Note the well defined quadratic trend up to $ M_{p} = 0.026 M_{s}$, and instability beyond this.}
+	\label{fig:planetmass}
+\end{figure}
+
+A quadratic trend (with a large quadratic component) is initially observed, but wander deviates from this at approximately $ M_{p} = 0.026 M_{s}$, with a sharp increase indicating orbit instability.
+This occurs significantly before the theoretical prediction of $ M_{p} = 0.04 M_{s}$ \cite{Darwin1897}, however it is suggested this is due to instabilities in the integration solver.
+
 \section{Conclusion}
@@
 
-Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
+\textit{Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
 Orbits in the stationary frame are animated and saved within "\textit{stat\_frame\_video}".
 
-2D perturbations are characterised in \textit{perturbation\_2D}.
+2D perturbations are characterised in \textit{perturbation\_2D}, while perturbations in the z direction are considered in \textit{perturbation\_3D}.
+Variation in planetary mass, for results in section \ref{planet}, are given in \textit{planet\_mass\_variation}.}
+
+Other code files are listed by Section below in Table \ref{Codefiles}:
+
+\begin{table}[ht]
+	\caption{Program files used to generate results in each section, with a description of their role.}
+	\centering
+	\begin{tabular}{|c|c|c|}
+		\hline
+		\textbf{Section} & \textbf{Description} & \textbf{Code File} \\
+		\hline \hline
+		\ref{Performance} & ODE solver performance & \textit{lagrange\_stability} \\
+		\hline
+		\ref{unperturbed} & Wander from unperturbed initial conditions & \textit{lagrange\_stability} \\
+		\hline
+		\ref{unperturbed} & Animations of unperturbed orbits & \textit{stat\_frame\_video} \\
+		\hline
+		\ref{perturbed}& Perturbations restricted to the xy plane & \textit{perturbation\_2D} \\
+		\hline
+		\ref{orbits}& Plotting of Orbit types in xy plane & \textit{perturbation\_2D} \\
+		\hline
+		\ref{3D}& Perturbations in the z-direction & \textit{perturbation\_3D} \\
+		\hline
+		\ref{planet}& Variation of planetary mass & \textit{planet\_mass\_variation} \\
+		\hline
+	\end{tabular}
+	\label{Codefiles}
+\end{table}
 

### 2020-04-24 Reference Scripts/backupfile.py

@@ -201,3 +201,144 @@ plt.legend()
 plt.show()
 
 # ani.save("asteroid_orbit.gif", writer="imagemagick")
+
+
+# VERSION OF PERTURBATIONS_3D.PY MADE AT 10:47 ON 24/04/2020
+
+import math
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    solar_rad,
+    planet_rad,
+    period,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+cos = math.cos(greek_theta)
+sin = math.sin(greek_theta)
+
+# ROTATING FRAME
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))
+perturb = 0.01 * np.array((cos, sin, 0, 0, 0, 0))
+perturb_z = 0.01 * np.array((cos, sin, 10, 0, 0, 0))
+orbit_sol = orbits.rotating_frame(initial_cond_rot + perturb)
+orbit_solz = orbits.rotating_frame(initial_cond_rot + perturb_z)
+
+
+# wander
+
+z_perturbations = np.linspace(0, 0.1, 30)
+wander_data = np.zeros((len(z_perturbations), 2))
+for n in range(len(z_perturbations)):
+    orbit_solz = orbits.rotating_frame(
+        # initial_cond_rot + np.array((0, 0, z_perturbations[n], 0, 0, 0))
+        initial_cond_rot
+        + np.array((0.02 * cos, 0.02 * sin, z_perturbations[n], 0, 0, 0))
+    )
+    wander_t = np.zeros((len(orbit_solz.t)))
+    wander_z = np.zeros((len(orbit_solz.t)))
+    for i in range(len(orbit_solz.t)):
+        wander_t[i] = np.linalg.norm(
+            orbit_solz.y[0:3, i] - initial_cond_rot[0:3] - perturb_z[0:3]
+        )  # deviation in pos magnitude
+        wander_z[i] = orbit_solz.y[2, i] - perturb_z[2]  # deviation in z pos only
+    wander_data[n, 0] = np.max(wander_t)
+    wander_data[n, 1] = np.max(wander_z)
+    print(n)
+
[…]

### 2020-04-24 constants.py

@@ -8,8 +8,8 @@ M_P = 0.001  # planetary mass
 R = 5.2  # average planetary radius from sun
 
 
-PRECISION = 1000  # evaluation points per orbit
-ORBIT_NUM = 40  # number of orbits
+PRECISION = 100  # evaluation points per orbit
+ORBIT_NUM = 30  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-24 perturbations_2D.py

@@ -89,7 +89,7 @@ import time
 z = np.zeros_like(x)
 for i, j in np.ndindex(z.shape):
     start = time.time()
-    z[i, j] = wander.initial_point((x[i, j], y[i, j]), pertubation_type="velocity")
+    z[i, j] = wander.initial_point((x[i, j], y[i, j], 0), pertubation_type="velocity")
     end = time.time()
     print(str((i, j)) + " in time " + str(end - start) + " s")
 

### 2020-04-24 perturbations_3D.py

@@ -1,0 +1,137 @@
+import math
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+import orbits
+from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import (
+    solar_rad,
+    planet_rad,
+    period,
+    time_span,
+    lagrange,
+)  # Derived constants
+
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+cos = math.cos(greek_theta)
+sin = math.sin(greek_theta)
+
+# ROTATING FRAME
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))
+perturb = 0.01 * np.array((cos, sin, 0, 0, 0, 0))
+perturb_z = 0.01 * np.array((cos, sin, 10, 0, 0, 0))
+orbit_sol = orbits.rotating_frame(initial_cond_rot + perturb)
+orbit_solz = orbits.rotating_frame(initial_cond_rot + perturb_z)
+
+
+# wander
+
+z_perturbations = np.linspace(0, 0.1, 30)
+wander_data = np.zeros((len(z_perturbations), 2))
+for n in range(len(z_perturbations)):
+    orbit_solz = orbits.rotating_frame(
+        # initial_cond_rot + np.array((0, 0, z_perturbations[n], 0, 0, 0))
+        initial_cond_rot
+        + np.array((0.02 * cos, 0.02 * sin, z_perturbations[n], 0, 0, 0))
+    )
+    wander_t = np.zeros((len(orbit_solz.t)))
+    wander_z = np.zeros((len(orbit_solz.t)))
+    for i in range(len(orbit_solz.t)):
+        wander_t[i] = np.linalg.norm(
+            orbit_solz.y[0:3, i] - initial_cond_rot[0:3] - perturb_z[0:3]
+        )  # deviation in pos magnitude
+        wander_z[i] = orbit_solz.y[2, i] - perturb_z[2]  # deviation in z pos only
+    wander_data[n, 0] = np.max(wander_t)
+    wander_data[n, 1] = np.max(wander_z)
+    print(n)
+
+fig = plt.figure()
+ax = fig.add_subplot()
+
+coeff = np.polyfit(
+    z_perturbations, wander_data[:, 0], 2
+)  # for quadratic line of best fit
+best_fit = np.poly1d(coeff)
[…]

### 2020-04-24 planet_mass_variation.py

@@ -6,15 +6,15 @@ import orbits
 
 # from wander import wander
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
-from constants import (
-    solar_rad,
-    planet_rad,
-    omega,
-    time_span,
-    lagrange,
-)  # Derived constants
+from constants import solar_rad  # Derived constants
 
-mass = np.linspace(0.001, 0.02, 100)  # range of planetary masses
+# import wander
+
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+cos = math.cos(greek_theta)
+sin = math.sin(greek_theta)
+
+mass = np.linspace(0.002, 0.05, 100)  # range of planetary masses
 max_wander = np.zeros_like(mass)
 for n in range(len(mass)):
     import constants
@@ -29,16 +29,58 @@ for n in range(len(mass)):
     import orbits
 
     initial_cond = np.array((constants.lagrange[0], constants.lagrange[1], 0, 0, 0, 0))
-    orbit = orbits.rotating_frame(initial_cond)
+    # orbit = orbits.rotating_frame(initial_cond)
+    orbit = orbits.rotating_frame(
+        (initial_cond + 0.001 * np.array((cos, sin, 0, sin, -cos, 0)))
+    )  # radial perturbation
     wander_t = np.zeros((len(orbit.t)))
     for i in range(len(orbit.t)):
         wander_t[i] = np.linalg.norm(
             orbit.y[0:3, i] - initial_cond[0:3]
         )  # deviation in pos only
     max_wander[n] = np.max(wander_t)
-plt.plot(mass, max_wander)
-plt.title("Change in wander with different planet mass")
+    print(n)
+
+fig = plt.figure()
+ax = fig.add_subplot()
+plt.plot(mass, max_wander, marker="x", linestyle="")
+
+coeff_quad = np.polyfit(
+    mass[0:50], max_wander[0:50], 2
+)  # for quadratic line of best fit
+best_fit_quad = np.poly1d(coeff_quad)
+best_fit_lin = np.poly1d(coeff_quad[1:])
+# without quadratic term for comparison
+order = np.argsort(mass)
+
+quad_label = str(
+    "Quadratic Best Fit: "
+    + str("{0:.2f}".format(coeff_quad[0]))
+    + "x\u00b2 "
[…]

### 2020-04-24 wander.py

@@ -78,11 +78,11 @@ def initial_point(pertubation, samples=1, pertubation_type="position"):
     """
     if pertubation_type == "position":
         initial_cond = initial_cond_rot + np.array(
-            (pertubation[0], pertubation[1], 0, 0, 0, 0)
+            (pertubation[0], pertubation[1], pertubation[2], 0, 0, 0)
         )  # add position pertubation
     elif pertubation_type == "velocity":
         initial_cond = initial_cond_rot + np.array(
-            (0, 0, 0, pertubation[0], pertubation[1], 0)
+            (0, 0, 0, pertubation[0], pertubation[1], pertubation[2])
         )  # add velocity pertubation
     else:
         raise ValueError(

### 2020-04-24 Report/draft1.tex
periodicity of z perturbations

@@
 \section{Methodology}
-This system of coupled first-order ordinary differential equations (ODEs) was solved using the scipy solve\_ivp function.
+This system of coupled first-order ordinary differential equations (ODEs) was solved using the scipy \textit{solve\_ivp} function.
 The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years.
@@
 %Then all theory could be in methodology?
-Within solve\_ivp, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) and gives a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
+Within \textit{solve\_ivp}, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) and gives a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
 This is larger than expected, indicating the system of equations requires an unreasonably small step size for numerical stability with respect to this method, even in regions where the solution curve is smooth \cite{Lambert1991}.
@@
 \subsection{Performance} \label{Performance}
-Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within solve\_ivp.
+Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within \textit{solve\_ivp}.
+
+\subsection{Fourier Analysis} \label{Fourier}
+For periodic oscillations/orbits, the Fast Fourier Transform (FFT) implemented in \textit{scipy} is used to obtain values for period, with peak finding in \textit{scipy.signal} identifying the exact values.
+The errors estimated using the associated peak widths of each frequency component.
 
@@
 In the stationary frame, this wander from the (now moving) Lagrange point is depicted in Figure \ref{fig:greeksdeviationstationaryframe}.
-The deviation oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, and a period of $148 \pm 4$ years, where the error was determined by the width of the Fourier peak.
+The deviation oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, and a period of $148 \pm 4$ years.
 This is modulated with a faster oscillation component of $11.85 \pm 0.27$ years, equivalent to the orbital period of the asteroids.
@@
 
-\subsection{Perturbations in z-direction} \label{3D}
+\subsection{Perturbations in Z-direction} \label{3D}
+Perturbations in the Z-direction are aligned with angular velocity vector for rotating frame, so experience no virtual forces in this direction, and oscillate under the influence of gravity alone.
+Considering small perturbations in the z direction (so that the distance from the CoM can be considered constant), the period of such oscillations tends to the orbital period of the rotating frame, under Newtonian gravity.
+
+These are measured to be $11.85 \pm 0.48$ years, in strong agreement with the theoretical predictions.
 
@@
 	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p5b}
-	\caption{Variation of maximum wander with planetary mass. Note the well defined quadratic trend up to $ M_{p} = 0.026 M_{s}$, and instability beyond this.}
+	\caption{Variation of maximum wander with planetary mass, under a 0.01 AU radial perturbation and 0.01 AU/year tangential boost.
+Note the well defined quadratic trend up to $ M_{p} = 0.026 M_{s}$, and instability beyond this.}
 	\label{fig:planetmass}
@@
 		\ref{Performance} & ODE solver performance & \textit{lagrange\_stability} \\
+		\hline
+		\ref{Fourier} & Fourier Analysis & \textit{lagrange\_stability} \\
 		\hline

### 2020-04-24 constants.py

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun
 
 
 PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 30  # number of orbits
+ORBIT_NUM = 100  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-24 perturbations_2D.py

@@ -245,7 +245,7 @@ peak_freq = np.zeros(len(peaks))
 for i in range(len(peaks)):
     peak_freq[i] = 1 / freqs[peaks[i]]
 print("Period (in years) of primary frequency components in Fourier spectrum: ")
-print(np.abs(np.unique(peak_freq)).round(2))
+print(np.unique(np.abs(peak_freq)).round(2))
 # Evaluated for 1000 points per orbit, 100 orbits
 
 # DEVIATION FROM LAGRANGE POINT

### 2020-04-24 perturbations_3D.py

@@ -2,8 +2,10 @@ import math
 
 import matplotlib.pyplot as plt
 import numpy as np
+import scipy.signal
 
 import orbits
+import wander
 from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
 from constants import (
     solar_rad,
@@ -25,25 +27,40 @@ orbit_sol = orbits.rotating_frame(initial_cond_rot + perturb)
 orbit_solz = orbits.rotating_frame(initial_cond_rot + perturb_z)
 
 
+# DEVIATION FROM LAGRANGE POINT
+plt.plot(time_span, orbit_solz.y[2, :])
+# plt.yscale("log", basey=10)
+plt.ticklabel_format(axis="y", style="", scilimits=None)
+plt.ylabel("Magnitude of Deviation in z-direction /AU")
+plt.xlabel("Time /years")
+plt.title("Asteroid Deviation from Lagrange point")
+# plt.savefig("asteroid_zdeviation_linear.png")
+plt.show()
+
+
+sampling_rate = PRECISION / period
+freqs = np.fft.fftfreq(len(orbit_solz.y[2, :])) * sampling_rate
+fourier = np.abs(np.fft.fft(orbit_solz.y[2, :]))
+
+peaks, properties = scipy.signal.find_peaks(fourier, prominence=(100), width=(0,))
+
+peak_freq = np.zeros(len(peaks))
+for i in range(len(peaks)):
+    peak_freq[i] = 1 / freqs[peaks[i]]
+print("Period (in years) of primary frequency components in Fourier spectrum: ")
+print(np.unique(np.abs(peak_freq)).round(2))
+# Evaluated for 100 points per orbit, 100 orbits
+
+
 # wander
 
-z_perturbations = np.linspace(0, 0.1, 30)
+z_perturbations = np.linspace(0, 0.2, 1)
 wander_data = np.zeros((len(z_perturbations), 2))
 for n in range(len(z_perturbations)):
-    orbit_solz = orbits.rotating_frame(
-        # initial_cond_rot + np.array((0, 0, z_perturbations[n], 0, 0, 0))
-        initial_cond_rot
-        + np.array((0.02 * cos, 0.02 * sin, z_perturbations[n], 0, 0, 0))
-    )
-    wander_t = np.zeros((len(orbit_solz.t)))
-    wander_z = np.zeros((len(orbit_solz.t)))
[…]

### 2020-04-24 planet_mass_variation.py

@@ -28,6 +28,15 @@ for n in range(len(mass)):
 
     import orbits
 
+    orbits.M_P = mass[n]
+    orbits.solar_rad = R * mass[n] / (M_S + mass[n])
+    orbits.planet_rad = R * M_S / (M_S + mass[n])
+    orbits.period = math.sqrt(R ** 3 / (M_S + mass[n]))
+    orbits.time_span = np.linspace(
+        0, ORBIT_NUM * orbits.period, int(ORBIT_NUM * PRECISION)
+    )
+    orbits.omega = 2 * np.pi / constants.period  # angular velocity of frame
+
     initial_cond = np.array((constants.lagrange[0], constants.lagrange[1], 0, 0, 0, 0))
     # orbit = orbits.rotating_frame(initial_cond)
     orbit = orbits.rotating_frame(
@@ -39,7 +48,7 @@ for n in range(len(mass)):
             orbit.y[0:3, i] - initial_cond[0:3]
         )  # deviation in pos only
     max_wander[n] = np.max(wander_t)
-    print(n)
+    print(str(n) + ": " + str(orbits.M_P))
 
 fig = plt.figure()
 ax = fig.add_subplot()

### 2020-04-24 wander.py

@@ -67,13 +67,12 @@ def rand_sample(max_pertubation_size, samples):
     return output
 
 
-def initial_point(pertubation, samples=1, pertubation_type="position"):
+def initial_point(pertubation, pertubation_type="position"):
     """Returns maximum wander over orbit in the rotating frame for given initial point 
 
     
     Wander is the maximum deviation from the initial point (not the lagrange point) over this timespan
     perturbation is the initial point in 2D position/velocity space
-    samples denotes the number of random pertubations sampled within that point for gives perturbation size
     perturbation_type can be "position" or "velocity" 
     """
     if pertubation_type == "position":

### 2020-04-24 Report/draft1.tex
perturbation_3D written (not tidy) pre-clean up planetary_mass variation

@@
 
-To consider a varying planet mass, it was considered preferable to avoid reconstructing all functions to take this as an argument as this requires re-evaluating all initial derived constants.
-Therefore, I iterate over alternative masses, re-defining constant values in this instance, and then directly import the required functions to compute the orbit.
+To consider a varying planet mass, it was considered preferable to avoid reconstructing all functions to take this as an argument, as this requires re-evaluating all initial derived constants.
+Therefore, I iterate over alternative planetary masses, re-defining constant values in this instance, and then directly importing the required functions to compute the orbit.
 \textbf{REPHRASE?}
@@
 Considering small perturbations in the z direction (so that the distance from the CoM can be considered constant), the period of such oscillations tends to the orbital period of the rotating frame, under Newtonian gravity.
+These are measured to be $11.85 \pm 0.48$ years, in strong agreement with the theoretical predictions.
 
-These are measured to be $11.85 \pm 0.48$ years, in strong agreement with the theoretical predictions.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/z_perturbations5c}
+	\caption{Variation of wander with increasing z perturbations.
+A strong linear component is observed initially in the un-perturbed case, however tends towards the perturbed quadratic case, while the linear offsets result from the initial perturbation in the xy plane.}
+	\label{fig:zwander}
+\end{figure}
+
+The overall wander, however, is observed to follow Fig \ref{fig:zwander}.
+It is suggested that, while the maximum deviation from the Lagrange point in the z direction is simply twice the initial perturbation, wander in the xy plane increases quadratically.
+This can be initiated from a radial displacement due to the reduced gravitational force  under the z perturbation, demonstrated in Fig \ref{fig:zorbit}, or more significantly from an initial perturbation in the xy plane.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/z_perturb_orbit}
+	\caption{Orbit of Greek asteroids following a 0.2 AU perturbation in the positive z direction, over 12 orbital periods.
+Note the initial outwards radial displacement despite the lack of perturbation in the xy plane, and subsequent tadpole-like orbit.}
+	\label{fig:zorbit}
+\end{figure}
 
@@
 
-\textit{Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
+Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
 Orbits in the stationary frame are animated and saved within "\textit{stat\_frame\_video}".
@@
 2D perturbations are characterised in \textit{perturbation\_2D}, while perturbations in the z direction are considered in \textit{perturbation\_3D}.
-Variation in planetary mass, for results in section \ref{planet}, are given in \textit{planet\_mass\_variation}.}
+Variation in planetary mass, for results in section \ref{planet}, are given in \textit{planet\_mass\_variation}.
 

### 2020-04-24 constants.py

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun
 
 
 PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 100  # number of orbits
+ORBIT_NUM = 20  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-24 perturbations_3D.py

@@ -19,46 +19,41 @@ greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 cos = math.cos(greek_theta)
 sin = math.sin(greek_theta)
 
-# ROTATING FRAME
 initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))
-perturb = 0.01 * np.array((cos, sin, 0, 0, 0, 0))
-perturb_z = 0.01 * np.array((cos, sin, 10, 0, 0, 0))
-orbit_sol = orbits.rotating_frame(initial_cond_rot + perturb)
-orbit_solz = orbits.rotating_frame(initial_cond_rot + perturb_z)
 
 
-# DEVIATION FROM LAGRANGE POINT
-plt.plot(time_span, orbit_solz.y[2, :])
-# plt.yscale("log", basey=10)
-plt.ticklabel_format(axis="y", style="", scilimits=None)
-plt.ylabel("Magnitude of Deviation in z-direction /AU")
-plt.xlabel("Time /years")
-plt.title("Asteroid Deviation from Lagrange point")
-# plt.savefig("asteroid_zdeviation_linear.png")
-plt.show()
+# # DEVIATION FROM LAGRANGE POINT
+# plt.plot(time_span, orbit_solz.y[2, :])
+# # plt.yscale("log", basey=10)
+# plt.ticklabel_format(axis="y", style="", scilimits=None)
+# plt.ylabel("Magnitude of Deviation in z-direction /AU")
+# plt.xlabel("Time /years")
+# plt.title("Asteroid Deviation from Lagrange point")
+# # plt.savefig("asteroid_zdeviation_linear.png")
+# plt.show()
 
 
-sampling_rate = PRECISION / period
-freqs = np.fft.fftfreq(len(orbit_solz.y[2, :])) * sampling_rate
-fourier = np.abs(np.fft.fft(orbit_solz.y[2, :]))
+# sampling_rate = PRECISION / period
+# freqs = np.fft.fftfreq(len(orbit_solz.y[2, :])) * sampling_rate
+# fourier = np.abs(np.fft.fft(orbit_solz.y[2, :]))
 
-peaks, properties = scipy.signal.find_peaks(fourier, prominence=(100), width=(0,))
+# peaks, properties = scipy.signal.find_peaks(fourier, prominence=(100), width=(0,))
 
-peak_freq = np.zeros(len(peaks))
-for i in range(len(peaks)):
[…]

### 2020-04-24 planet_mass_variation.py

@@ -14,7 +14,7 @@ greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 cos = math.cos(greek_theta)
 sin = math.sin(greek_theta)
 
-mass = np.linspace(0.002, 0.05, 100)  # range of planetary masses
+mass = np.linspace(0.001, 0.04, 50)  # range of planetary masses
 max_wander = np.zeros_like(mass)
 for n in range(len(mass)):
     import constants
@@ -50,34 +50,34 @@ for n in range(len(mass)):
     max_wander[n] = np.max(wander_t)
     print(str(n) + ": " + str(orbits.M_P))
 
-fig = plt.figure()
-ax = fig.add_subplot()
+# fig = plt.figure()
+# ax = fig.add_subplot()
 plt.plot(mass, max_wander, marker="x", linestyle="")
 
-coeff_quad = np.polyfit(
-    mass[0:50], max_wander[0:50], 2
-)  # for quadratic line of best fit
-best_fit_quad = np.poly1d(coeff_quad)
-best_fit_lin = np.poly1d(coeff_quad[1:])
-# without quadratic term for comparison
-order = np.argsort(mass)
+# coeff_quad = np.polyfit(
+#     mass[0:50], max_wander[0:50], 2
+# )  # for quadratic line of best fit
+# best_fit_quad = np.poly1d(coeff_quad)
+# best_fit_lin = np.poly1d(coeff_quad[1:])
+# # without quadratic term for comparison
+# order = np.argsort(mass)
 
-quad_label = str(
-    "Quadratic Best Fit: "
-    + str("{0:.2f}".format(coeff_quad[0]))
-    + "x\u00b2 "
-    + str("{0:+.2f}".format(coeff_quad[1]))
-    + "x "
-    + str("{0:+.2f}".format(coeff_quad[2]))
-)
+# quad_label = str(
+#     "Quadratic Best Fit: "
+#     + str("{0:.2f}".format(coeff_quad[0]))
+#     + "x\u00b2 "
+#     + str("{0:+.2f}".format(coeff_quad[1]))
+#     + "x "
+#     + str("{0:+.2f}".format(coeff_quad[2]))
+# )
 
-plt.plot(
-    mass[0:50],
-    best_fit_quad(mass)[0:50],
-    linewidth=0.75,
-    color="navy",
-    label=quad_label,
-)
+# plt.plot(
[…]

### 2020-04-24 Report/draft1.tex
Complete planet_mass_variation

@@
 I'm going to make that paragraph a little longer to fill out the space in case this is forming some kind of error.
-i really hope this is long enough.}
+i really hope this is long enough.
+Word Count: XXXX}
 
@@
 This can be initiated from a radial displacement due to the reduced gravitational force  under the z perturbation, demonstrated in Fig \ref{fig:zorbit}, or more significantly from an initial perturbation in the xy plane.
+However, the oscillation of gravitational forces due to perturbations in the z direction make this problem significantly more complicated and likely result in the observed deviations from the quadratic relations.
 
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p5b}
+	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p6}
 	\caption{Variation of maximum wander with planetary mass, under a 0.01 AU radial perturbation and 0.01 AU/year tangential boost.
-Note the well defined quadratic trend up to $ M_{p} = 0.026 M_{s}$, and instability beyond this.}
+Note the orbit stability up to $ M_{p} = 0.04 M_{s}$, and instability beyond this.}
 	\label{fig:planetmass}
@@
 
-A quadratic trend (with a large quadratic component) is initially observed, but wander deviates from this at approximately $ M_{p} = 0.026 M_{s}$, with a sharp increase indicating orbit instability.
-This occurs significantly before the theoretical prediction of $ M_{p} = 0.04 M_{s}$ \cite{Darwin1897}, however it is suggested this is due to instabilities in the integration solver.
+An approximate quadratic trend is initially observed in the small mass region, with minimum wander occurring at $ M_{p} = 0.026 M_{s}$.
+However the variation of wander below $ M_{p} = 0.04 M_{s}$ in minimal, while orbits are unstable beyond this point, in agreement with theoretical predictions by Darwin \cite{Darwin1897}.
 

### 2020-04-24 constants.py

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun
 
 
 PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 20  # number of orbits
+ORBIT_NUM = 100  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-24 planet_mass_variation.py

@@ -3,27 +3,22 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 import orbits
-
-# from wander import wander
-from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import M_S, ORBIT_NUM, PRECISION, R  # User defined constants
 from constants import solar_rad  # Derived constants
 
-# import wander
 
 greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 cos = math.cos(greek_theta)
 sin = math.sin(greek_theta)
 
-mass = np.linspace(0.001, 0.04, 50)  # range of planetary masses
+mass = np.linspace(0.0005, 0.0425, 100)  # range of planetary masses
 max_wander = np.zeros_like(mass)
 for n in range(len(mass)):
     import constants
 
-    constants.M_P = mass[n]
     constants.solar_rad = R * mass[n] / (M_S + mass[n])
     constants.planet_rad = R * M_S / (M_S + mass[n])
     constants.period = math.sqrt(R ** 3 / (M_S + mass[n]))
-    constants.omega = 2 * np.pi / constants.period  # angular velocity of frame
     constants.lagrange = (constants.planet_rad - R / 2, R * math.sqrt(3) / 2, 0)
 
     import orbits
@@ -38,7 +33,6 @@ for n in range(len(mass)):
     orbits.omega = 2 * np.pi / constants.period  # angular velocity of frame
 
     initial_cond = np.array((constants.lagrange[0], constants.lagrange[1], 0, 0, 0, 0))
-    # orbit = orbits.rotating_frame(initial_cond)
     orbit = orbits.rotating_frame(
         (initial_cond + 0.001 * np.array((cos, sin, 0, sin, -cos, 0)))
     )  # radial perturbation
@@ -50,45 +44,10 @@ for n in range(len(mass)):
     max_wander[n] = np.max(wander_t)
     print(str(n) + ": " + str(orbits.M_P))
 
-# fig = plt.figure()
-# ax = fig.add_subplot()
 plt.plot(mass, max_wander, marker="x", linestyle="")
 
-# coeff_quad = np.polyfit(
-#     mass[0:50], max_wander[0:50], 2
-# )  # for quadratic line of best fit
-# best_fit_quad = np.poly1d(coeff_quad)
[…]

### 2020-04-24 constants.py
cleaned perturbations_3D ready to submit

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun
 
 
 PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 100  # number of orbits
+ORBIT_NUM = 30  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-24 perturbations_3D.py

@@ -6,10 +6,9 @@ import scipy.signal
 
 import orbits
 import wander
-from constants import M_P, M_S, ORBIT_NUM, PRECISION, G, R  # User defined constants
+from constants import ORBIT_NUM, PRECISION, R  # User defined constants
 from constants import (
     solar_rad,
-    planet_rad,
     period,
     time_span,
     lagrange,
@@ -23,40 +22,45 @@ initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))
 
 
 # # DEVIATION FROM LAGRANGE POINT
-# plt.plot(time_span, orbit_solz.y[2, :])
-# # plt.yscale("log", basey=10)
-# plt.ticklabel_format(axis="y", style="", scilimits=None)
-# plt.ylabel("Magnitude of Deviation in z-direction /AU")
-# plt.xlabel("Time /years")
-# plt.title("Asteroid Deviation from Lagrange point")
-# # plt.savefig("asteroid_zdeviation_linear.png")
-# plt.show()
+perturb_z = np.array((0, 0, 0.2, 0, 0, 0))
+orbit_sol_z = orbits.rotating_frame(initial_cond_rot + perturb_z)
+
+plt.plot(time_span, orbit_sol_z.y[2, :])
+plt.ylabel("Magnitude of Deviation in z-direction /AU")
+plt.xlabel("Time /years")
+plt.title("Asteroid Deviation from Lagrange point")
+# plt.savefig("asteroid_zdeviation_linear.png")
+plt.show()
 
 
-# sampling_rate = PRECISION / period
-# freqs = np.fft.fftfreq(len(orbit_solz.y[2, :])) * sampling_rate
-# fourier = np.abs(np.fft.fft(orbit_solz.y[2, :]))
+sampling_rate = PRECISION / period
+freqs = np.fft.fftfreq(len(orbit_sol_z.y[2, :])) * sampling_rate
+fourier = np.abs(np.fft.fft(orbit_sol_z.y[2, :]))
 
-# peaks, properties = scipy.signal.find_peaks(fourier, prominence=(100), width=(0,))
+peaks, properties = scipy.signal.find_peaks(fourier, prominence=(100), width=(0,))
 
-# peak_freq = np.zeros(len(peaks))
-# for i in range(len(peaks)):
-#     peak_freq[i] = 1 / freqs[peaks[i]]
-# print("Period (in years) of primary frequency components in Fourier spectrum: ")
[…]

### 2020-04-26 Report/draft1.tex
cleaned perturbations_3D ready to submit
updated report clean main function files

--- 2020-04-24 Report/draft1.tex
@@
 	%\authorstyle{Christopher Gallagher}}
-\author{\authorstyle{Christopher Gallagher} %Candidate 6952R
+\author{\authorstyle{Christopher Gallagher} %Candidate 6952R}
 	\institution{University of Cambridge}}
@@
 %\AtBeginBibliography{\small}
+
+\usepackage{amsmath}
+\usepackage{bm}
+\usepackage{fixltx2e}% for \textsubscript
+
 %----------------------------------------------------------------------------------------
@@
 
-\lettrineabstract{Trojan asteroids are v important.
-I'm going to make that paragraph a little longer to fill out the space in case this is forming some kind of error.
-i really hope this is long enough.}
+\lettrineabstract{Trojan asteroids orbit the Sun in resonance with Jupiter at Lagrange points,
+	and have their own stable orbits about these points.
+This can be treated as a restricted three-body problem, which is numerically evaluated in this report using a stiff Radau method to solve the system of coupled differential equations.
+Without perturbations, a maximal deviation of 4.68\pmb{$\times$}10\pmb{$^{-13}$} AU  from the Lagrange points was observed over 1000 orbits (11,852 years).
+The application of perturbations to these asteroids demonstrated the predicted existence of stable tadpole and horseshoe orbits, and suggested that wander from the Lagrange point depends quadratically on perturbation size.
+Such orbits were observed to be stable up to a planetary mass of 0.04 \pmb{$M_{\odot}$}, also in agreement with theoretical predictions.
+Word Count: XXXX}
 
@@
 \section{Introduction}
-The Jupiter trojans, commonly known as the Trojan asteroids, are two large groups of asteroids that share the planet Jupiter's orbit around the Sun in a 1:1 orbit resonance.
+The Jupiter Trojans, commonly known as the Trojan asteroids, are two large groups of asteroids that share the planet Jupiter's orbit around the Sun in a 1:1 orbit resonance.
 These two groups are called the Greeks and the Trojans, named after opposing sides in the mythological Trojan war, and lead/trail Jupiter respectively in its orbit.
-They correspond to Jupiter's two stable Lagrange points: L\subscript{4}, lying 60° ahead of the planet in its orbit, and L\subscript{5}, 60° behind, with asteroids distributed in two elongated, curved regions around these Lagrangian points.
+They correspond to Jupiter's two stable Lagrange points: L\subscript{4}, lying 60° ahead of the planet in its orbit, and L\subscript{5}, 60° behind, with asteroids distributed in two elongated, curved regions around these Lagrange points.
 
-The first Jupiter trojan, 588 Achilles, was discovered in 1906 by the German astronomer Max Wolf \cite{Nicholson1961}, and a total of 7642 Jupiter trojans have been found as of February 2020 \cite{IAU2020}.
+The first Jupiter Trojan, 588 Achilles, was discovered in 1906 by the German astronomer Max Wolf \cite{Nicholson1961}, and a total of 7642 Jupiter Trojans have been found as of February 2020 \cite{IAU2020}.
 
-Research into Jupiter's trojan asteroids continues, with the particular focus on their origins reliant on an understanding of their orbit stability \cite{DiSisto2019}, \cite{Nesvorn2018}.
+Research into Jupiter's Trojan asteroids continues, with the particular focus on their origins reliant on an understanding of their orbit stability \cite{DiSisto2019} \cite{Nesvorn2018}.
 This informs studies into their composition \cite{Brown2016}, as travel to these asteroids is considered for their potential in mineral mining \cite{Okada2017} \cite{Levison2016}.
 
-The purpose of this report is to use numerical simulation techniques to investigate the stability of orbits about these Lagrange points, demonstrating the asteroid oscillate about these points under small perturbations and quantifying the absolute distance of the asteroids from the Lagrange point (the wander) during their orbits.
+The purpose of this report is to use numerical simulation techniques to investigate the stability of orbits about the Lagrange points, demonstrating that asteroids oscillate about these points with small perturbations and quantifying the absolute distance of the asteroids from the Lagrange point (the wander) during their orbits.
 The impact of variation in planetary/solar mass on asteroid orbit stability will also be considered.
-\textit{Signpost what is included in each section?}
 
@@
 
-\section{Theoretical Background?}
+\section{Theoretical Background}
 
 \subsection{Lagrange Points} \label{lagrange}
-The asteroids exist at/near Lagrange points, defined in Lagrange's initial analysis of the three-body problem in 1772 \cite{Lagrange1772}, were he demonstrated the existence of five equilibrium points for an object of negligible mass orbiting under the gravitational effect of two larger masses.
-Three of these equilibrium points, L\subscript{1}-L\subscript{3} lie on the line joining the two masses, while each of the remaining two points, L\subscript{4} and L\subscript{5}, lie at the apex of an equilateral triangle with base equal to the separation of the two masses (see Figure \ref{fig:lagrangepoints}).
-Despite all these points being potential maxima, stable motion is possible around L\subscript{4} and L\subscript{5} due to the Coriolis force \cite{Lissauer2014}.
+The asteroids exist at or near Lagrange points, defined in Lagrange's initial analysis of the three-body problem in 1772 \cite{Lagrange1772}, in which he demonstrated the existence of five equilibrium points for an object of negligible mass orbiting under the gravitational effect of two larger masses.
+Three of these equilibrium points, L\subscript{1}\textendash L\subscript{3} lie on the line joining the two masses, while each of the remaining two points, L\subscript{4} and L\subscript{5}, lie at the apex of an equilateral triangle with base equal to the separation of the two masses (see Figure \ref{fig:lagrangepoints}).
+Although these points are all potential maxima, stable motion is possible around L\subscript{4} and L\subscript{5} due to the Coriolis force \cite{Lissauer2014}.
 %lissauer also says stability is provided that the most massive body has at least 25 times the mass of the secondary
@@
 The solar and planetary masses are denoted by the large and small filled circles, and the letters P, H, and T denote passing, horseshoe, and tadpole orbits respectively.
-Note that the two masses form an equilateral triangle with each of the L\subscript{4} and L\subscript{5} points.
+Note that the two massive bodies form an equilateral triangle with each of the L\subscript{4} and L\subscript{5} points.
 Reproduced from Marzari et al.
@@
 
-Figure \ref{fig:lagrangepoints} depicts orbits about L\subscript{4} and L\subscript{5} (known as tadpole orbits), as well as orbits between Lagrange points are also possible (such as horseshoe orbits), described by Murray et al.
+Figure \ref{fig:lagrangepoints} depicts orbits about L\subscript{4} and L\subscript{5} (known as tadpole orbits), as well as horseshoe orbits between Lagrange points, described by Murray et al.
 \cite{Murray1999}.
 
-This report will focus on tadpole orbits, a well documented feature of Trojan orbits \cite{Garfinkel1983}, \cite{Dermott1981}, and the restricted three-body problem in general.
-Their distinctive shape is result of a long-period motion about the equilibrium point combining with a short-period oscillation due to Keplerian motion of the asteroid.
+This report will focus on tadpole orbits, a well-documented feature of Trojan orbits \cite{Garfinkel1983} \cite{Dermott1981} and of the restricted three-body problem in general.
+Their distinctive shape results from a long-period motion about the equilibrium point combined with a short-period oscillation due to the Keplerian motion of the asteroid.
 
-Szebehely et al. \cite{Szebehely1969} predict this short period tends to the planetary period in the small planetary mass limit, while the long period is given by
+Szebehely et al. \cite{Szebehely1969} predict that this short period tends to the planetary period in the small planetary mass limit, while the long period is given by
 
 \begin{equation}
-T_{long} = T_{P} \sqrt{\frac{4}{27 \mu_{2}}},  
+T_{long} = T_{P} \sqrt{\frac{4}{27 \mu_{2}}}  
 \end{equation}
@@
 \subsection{Theoretical Model} \label{theory}
-The three body problem, where the dynamics of three interacting bodies are determined given their initial positions and velocities, has no analytical (closed-form) solution in the general case \cite{Barrow2008}.
+The three-body problem, where the dynamics of three interacting bodies are determined from their initial positions and velocities, has no analytical (closed-form) solution in the general case \cite{Barrow2008}.
 
@@
 
-The system of differential coordinates determine the position and velocity of the asteroids, with two equations per spatial coordinate.
+The system of differential equations determines the position and velocity of the asteroids, with two equations per spatial coordinate.
 \begin{equation}
@@
 
-We may also consider a frame rotating at the same speed as the massive bodies.
+We may also consider a frame rotating at the same velocity as the massive bodies.
 As there is 1:1 orbital resonance between Jupiter and the asteroids, all three bodies are stationary in this frame.
-This significantly increases the accuracy of numerical simulations, as the exact solution is stationary with no explicit time dependence, rather then requiring an infinite power series \cite{Guglielmi2001}, \cite{LeVeque2007}.
+This significantly increases the accuracy of numerical simulations, as the exact solution is stationary with no explicit time dependence, rather than requiring an infinite power series \cite{Guglielmi2001}, \cite{LeVeque2007}.
 
@@
 
-\subsection{Symmetry}
-This problem contains a number of symmetries, which could be employed to simplify the problem and reduce the computational load.
-The Trojan and Greek asteroids are in equivalent positions, so experience the same forces, and the system is rotationally and inversion symmetric, so the choice of initial point and orbit direction is arbitrary.
-Therefore only the Greeks, orbiting counter-clockwise with perturbations applied at t = 0, need to be investigated.
+\subsection{Symmetry} \label{symmetry}
+This problem contains a number of symmetries, which were employed to simplify the problem and reduce the computational load.
+The Trojan and Greek asteroids are in equivalent positions, so experience the same forces, and the system has rotational and inversion symmetry, so the choice of initial point and orbit direction is arbitrary.
+Therefore, only the Greeks, orbiting counter-clockwise with perturbations applied at t = 0, need be considered.
 
 \subsection{Orbit Geometry}
-As the three bodies considered here form an equilateral triangle in the initial equilibrium state, as depicted in Figure \ref{fig:geometry}, we can derive the polar coordinates of each body with respect to the centre of mass about which the bodies orbit.
+As the three bodies considered here form an equilateral triangle in the initial equilibrium state, as depicted in Figure \ref{fig:geometry}, we can easily derive the polar coordinates of each body with respect to the centre of mass about which the bodies orbit.
 
-\begin{figure}
+\begin{figure} %place in text?
 	\centering
@@
 	}	
-	\caption{A geometric depiction of the three-body system, in the case where the planet has a mass equal to a third of the sun, and asteroids are considered at the L\subscript{4} point.
+	\caption{A geometric depiction of the three-body system, in the case where the planet has a mass equal to one-third of the sun, and asteroids are considered at the L\subscript{4} point.
 $ R_{s} $ and $ R_{p}$ denote the (fixed) radii from the centre of mass (the grey point) to the Sun and the planet respectively, while $ r_{a} $ denotes the radius of the asteroids.}
@@
 
-\begin{equation}
+\begin{equation} \label{frequency}
 \Omega = \sqrt{\frac{G (M_{s} + M_{p})}{a^{3}}}
@@
 The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years.
-Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in earth years and mass in multiples of the solar mass, and to prevent floating point overflows due to the magnitude of the quantities considered in SI units.
+Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in Earth years and mass in multiples of the solar mass, to prevent floating point overflows due to the magnitude of the quantities considered in SI units.
 
@@
 \subsection{Integration Method}
-%Put this with theory as "numerical methods"?
-%Then all theory could be in methodology?
-Within \textit{solve\_ivp}, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) and gives a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 years.
-This is larger than expected, indicating the system of equations requires an unreasonably small step size for numerical stability with respect to this method, even in regions where the solution curve is smooth \cite{Lambert1991}.
-This suggests the system is stiff, and solvers designed for this typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to the non-stiff solvers such as RK45 \cite{Byrne1987}.
+Within \textit{solve\_ivp}, the default solver is RK45 (an explicit Runge-Kutta method of order 5(4) \cite{Dormand1980}) which gives a deviation in asteroid position (from the Lagrange point) in the order of $ 10^{-4}$ AU in the rotating frame over 50 orbits.
+This is larger than expected, suggesting that the system of equations may require an unreasonably small step-size for numerical stability in this method, even in regions where the solution curve is smooth \cite{Lambert1991}.
+Such systems are known as stiff, and solvers designed for these systems typically do more work per step, allowing them to take much larger steps, and have improved numerical stability compared to non-stiff solvers, such as RK45 \cite{Byrne1987}.
 
-Instead the stiff "Radau" solver (an implicit Runge-Kutta method of the Radau IIA family of order 5 \cite{Hairer2010}) is used for increased stability \cite{Frank1985}, and achieves a deviation in asteroid position in the order of $ 10^{-13}$ AU instead.
+Therefore the stiff Radau solver (an implicit Runge-Kutta method from the Radau IIA family of order 5 \cite{Hairer2010}) is used for increased stability \cite{Frank1985}, and achieves a deviation in asteroid position in the order of $ 10^{-13}$ AU instead.
 This also ensures stability in the stationary frame, with deviations of 0.76\% in asteroid separation from Jupiter over $ 10^{3} $ years, compared to 53\% for the best non-stiff solvers.
@@
 \subsection{Programme Structure}
-Global constants such as solar mass, and sun\textendash planet separation, along with derived values from these such as orbital period and solar radius from the CoM, are given in an importable python module "\textit{constants}".
+Global constants, such as solar mass, and sun\textendash planet separation, along with derived values from these, such as orbital period and solar radius from the CoM, are given in an importable python module, "\textit{constants}".
 
-Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit (under different sampling routines) are implemented in "\textit{wander}".
-Further files then import these modules and produce the plots given in this report, fully detailed in appendix \ref{structure}.
+Functions to evaluate these coupled differential equation systems are defined in the module "\textit{orbits}", while additional functions to evaluate the wander during the orbit (under different sampling routines) are implemented in "\textit{wander}".
+Complete code listings for these files are given in Appendix \ref{Code}.
+Further files then import these modules and produce the plots given in this report, fully detailed in Appendix \ref{structure}.
 
-To consider a varying planet mass, it was considered preferable to avoid reconstructing all functions to take this as an argument as this requires re-evaluating all initial derived constants.
-Therefore, I iterate over alternative masses, re-defining constant values in this instance, and then directly import the required functions to compute the orbit.
-\textbf{REPHRASE?}
-
-Complete code listings are given in appendix \ref{Code}.
+When varying planetary mass, I considered it preferable to avoid reconstructing all functions to take planetary mass as an argument, as this requires re-evaluating all initial derived constants in the "\textit{orbits}" module.
+Therefore, I iterate over alternative planetary masses, re-defining constant values in each instance, and directly import the required functions to compute the orbit.
 
 \subsection{Performance} \label{Performance}
-Sampling 100 points per orbit for 100 orbits takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and orbit number up to the array memory limit, achieved through the optimised integration routines within \textit{solve\_ivp}.
+Sampling 100 points per orbit for 100 orbital periods takes a mean time of $16.97 \pm 34$ \si{\milli\second}, with sub-linear scaling for sampling rate and approximate linear scaling with orbit number up to the array memory limit, achieved through the optimised integration routines within \textit{solve\_ivp}.
+Computing the wander over position/velocity space with larger perturbations was, however, more time consuming, averaging $2.04 \pm 0.16$ \si{\second} per point with the same parameters.
+
+The main computational load was within the \textit{solve\_ivp} function, which is already optimised well beyond the capabilities of the author.
+However, assumptions made in Section \ref{symmetry}, such as considering only one group of asteroids, reduces this time somewhat, and vectorisation of other aspects of the \textit{orbit} functions reduced the running time of the whole module.
+These approaches were not taken universally however, due to the size of the arrays required, and the dominant effect of the ODE solver on the running time; instead, the orbit number was reduced when only a comparison of the maximum wander was required.
 
 \subsection{Fourier Analysis} \label{Fourier}
-For periodic oscillations/orbits, the Fast Fourier Transform (FFT) implemented in \textit{scipy} is used to obtain values for period, with peak finding in \textit{scipy.signal} identifying the exact values.
-The errors estimated using the associated peak widths of each frequency component.
+For periodic oscillations and orbits, the Fast Fourier Transform (FFT) implemented in \textit{scipy} is used to obtain the period, with \textit{find\_peaks} in \textit{scipy.signal} identifying the exact values.
+The errors were estimated using the associated peak widths of each frequency component.
 
@@
 \subsection{Unperturbed Stability of Lagrange Points} \label{unperturbed}
-Considering the Greeks' orbit without perturbations applied, it has a maximum deviation of $4.68 \times 10^{-13}$ AU over 100 orbits (1185 years).
+Without perturbations applied, the Greeks' orbit has a maximum deviation from L\subscript{4} of $4.68 \times 10^{-13}$ AU over 100 orbits (1185 years).
 This value is unchanged if 1000 orbits are considered instead, confirming the stability of this Lagrange point.
@@
 In the stationary frame, this wander from the (now moving) Lagrange point is depicted in Figure \ref{fig:greeksdeviationstationaryframe}.
-The deviation oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, and a period of $148 \pm 4$ years.
+The wander oscillates with a magnitude of $9.10 \times 10^{-2}$ AU, over a period of $148 \pm 4$ years.
 This is modulated with a faster oscillation component of $11.85 \pm 0.27$ years, equivalent to the orbital period of the asteroids.
@@
 
-These much more significant errors are due to time dependence in the exact solution, as detailed in Section \ref{theory}.
+These much more significant deviations are due to time-dependence in the exact solution, as detailed in Section \ref{theory}.
 Energy can also be evaluated, and conserved, in this inertial frame; asteroid specific energy varies within only 0.113\% of the initial value, and with a similar periodicity to wander.
@@
 Animations produced to demonstrate the orbit in the stationary frame are included in Supplementary Material I-II.
-Animation I depicts the orbit as evaluated by the radau solver, while II depicts it with LSODA, demonstrating the drift present over time with non-stiff solvers.
+Animation I depicts the orbit as evaluated by the Radau solver, while II depicts it with LSODA, demonstrating the drift present over time with non-stiff solvers.
 
 \subsection{Wander Analysis} \label{perturbed}
-The wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin, and considered  separately perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively).
+Wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin.
+The perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively) were considered separately.
 By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).
@@
 Figure \ref{fig:wander_rad} demonstrates a polynomial dependence on perturbation size in the radial direction, with a negligible constant term.
-While a quadratic has been fitted here, it was not possible to eliminate the possibility of higher order terms, as increasing perturbation size beyond 0.06 AU can lead to unstable orbits.
+While a quadratic has been fitted here, the possibility of higher order terms could not be eliminated, as increasing perturbation size beyond 0.06 AU can lead to unstable orbits.
 
 We may consider the wander resulting from perturbations in position and velocity space.
-Figure \ref{fig:position_mesh} clearly shows the dependence on radial position perturbations only, while Figure \ref{fig:velocity_mesh} shows a similar the dependence on tangential velocity perturbations.
+Figure \ref{fig:position_mesh} clearly shows dependence on radial position perturbations only, while Figure \ref{fig:velocity_mesh} shows a similar dependence on tangential velocity perturbations.
 
@@
 	\caption{Orbits from an initial perturbation of magnitude 0.01 AU from the origin, in radial and tangential directions, over 12 orbital periods.
-Note the 'tadpole' features resulting from the radial perturbation, and the significantly larger wander than from the tangential perturbation.}
+Note the tadpole features resulting from the radial perturbation, and the significantly larger wander than from the tangential perturbation.}
 	\label{fig:orbitplots}
@@
 
-The "tadpole" orbit in Figure \ref{fig:orbit_rad} consists of two oscillating components, as described in Section \ref{lagrange}.
+The tadpole orbit in Figure \ref{fig:orbit_rad} consists of two oscillating components, as described in Section \ref{lagrange}.
 The short period, measured at $11.85 \pm 0.61$ years, is in excellent agreement with  planetary orbital period as expected; meanwhile the long period is measured to be $148 \pm 5$ years, consistent with the analytical value of $144$ years.
 
-For a narrow range of larger perturbations, stable horseshoe orbits encompassing both Lagrange points may also be observed as depicted in Figure \ref{fig:horseshoe}.
+For a narrow range of larger perturbations, stable horseshoe orbits encompassing both Lagrange points may also be observed, as depicted in Figure \ref{fig:horseshoe}.
 This has a full period of $353$ years, in agreement with Taylor et al.'s numerical result of $358 \pm 7$ years \cite{Taylor1981}.
@@
 
-\subsection{Perturbations in Z-direction} \label{3D}
-Perturbations in the Z-direction are aligned with angular velocity vector for rotating frame, so experience no virtual forces in this direction, and oscillate under the influence of gravity alone.
-Considering small perturbations in the z direction (so that the distance from the CoM can be considered constant), the period of such oscillations tends to the orbital period of the rotating frame, under Newtonian gravity.
+\subsection{Perturbations in Z direction} \label{3D}
+Perturbations in the Z direction are aligned with the angular velocity vector for the rotating frame, so experience no virtual forces in this direction, and oscillate under the influence of gravity alone.
+Considering small perturbations in the Z direction (so that the distance from the CoM can be considered constant), the period of such oscillations tend to the orbital period of the rotating frame under Newtonian gravity, as detailed in Appendix \ref{zperiod}.
+These oscillations are observed to be approximately sinusoidal, with a measured period of $11.85 \pm 0.48$ years, in strong agreement with the theoretical predictions.
 
-These are measured to be $11.85 \pm 0.48$ years, in strong agreement with the theoretical predictions.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/z_perturbations5c}
+	\caption{Variation of wander with increasing Z perturbations.
+A strong linear component is observed initially in the un-perturbed case, but then tends towards the perturbed quadratic case, while the linear offsets result from the initial perturbation in the XY plane.}
+	\label{fig:zwander}
+\end{figure}
+
+The overall wander, however, is observed to follow Figure \ref{fig:zwander}.
+It is suggested that, while the maximum deviation from the Lagrange point in the Z direction is simply twice the initial perturbation, wander in the XY plane increases quadratically.
+This can be initiated from a radial displacement due to the reduced gravitational force  under the Z perturbation, demonstrated in Figure \ref{fig:zorbit} or, more significantly, from an initial perturbation in the XY plane.
+However, the oscillation of gravitational forces due to perturbations in the Z direction make this problem significantly more complicated and are a possible cause of the observed deviations from the quadratic relations.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/z_perturb_orbit}
+	\caption{Orbit of Greek asteroids following a 0.2 AU perturbation in the positive Z direction, over 12 orbital periods.
+Note the initial outwards radial displacement despite the lack of perturbation in the XY plane, and subsequent tadpole-like orbit.}
+	\label{fig:zorbit}
+\end{figure}
 
@@
 
-A quadratic trend (with a large quadratic component) is initially observed, but wander deviates from this at approximately $ M_{p} = 0.026 M_{s}$, with a sharp increase indicating orbit instability.
-This occurs significantly before the theoretical prediction of $ M_{p} = 0.04 M_{s}$ \cite{Darwin1897}, however it is suggested this is due to instabilities in the integration solver.
+An approximate quadratic trend is initially observed in the small mass region, with minimum wander occurring at $ M_{p} = 0.026 M_{s}$.
+However, the variation of wander below $ M_{p} = 0.04 M_{s}$ is minimal, while orbits are unstable beyond this point, in agreement with theoretical predictions by Darwin \cite{Darwin1897}.
 
 \section{Conclusion}
+This report demonstrates the stability of Lagrange points L\subscript{4} and L\subscript{5}, with a maximum deviation from the Lagrange points of $4.68 \times 10^{-13}$ AU over 1000 orbits (11,852 years), in a frame rotating with the asteroids and without perturbations applied.
+Stability under perturbations in both the XY plane and in the Z direction is also demonstrated, with a finite wander from these points independent of orbit duration.
+It was possible to further quantify this wander, suggesting a strong quadratic dependence on radial perturbation magnitude and independence of tangential perturbations, which did not grow beyond the initial perturbation size.
+It was also possible to replicate both tadpole and horseshoe orbits from radial perturbations, and frequency components were obtained for these orbits in agreement with previous literature.
+It was also found that orbits become unstable when the planetary\textendash solar mass ratio is greater than 0.04, in agreement with theoretical predictions, and there is little mass dependence below this point.
+
+Further work is required to evaluate coefficient values for polynomial relationships observed, and to demonstrate a theoretical basis for such dependencies.
+Further research would also be beneficial on the categorisation of different orbit types, and examination of initial conditions corresponding to formation mechanisms of these orbits.
 
@@
 \begin{appendices}
+\section{Z Perturbation Period} \label{zperiod}
+Figure \ref{fig:geometryz} depicts a Z perturbation out of the XY plane, from the L\subscript{4} point.
+\begin{figure} [h]
+	\centering
+	\resizebox{0.5\columnwidth}{!}{%
+		\begin{tikzpicture}
+		\draw (0,0) -- (6,0)  node[midway, below]{$a$} node[right]{L\subscript{4}}; 
+		\draw [arrows=->] (6,0) --  (6, 1.95 ) node[above]{Asteroids} node[midway,right]{$z_{0}$};
+		\draw [dashed] (6,2) -- (0,0) node[midway, above]{$R$};
+		\node [label={[shift = {(-7mm,-7mm)}]{Sun}}] {};
+		\fill[black] (0,0) circle (3.5mm);% node[anchor=north west] {Sun};
+		\fill[black] (6,2) circle (0.5mm);
+		\fill[gray] (6,0) circle (0.5mm);
+
+		\draw  (1,0) arc (0:{atan(1/3)}:1cm)  ;
+		\node [label={[shift = {(14mm,-2mm)}]{$\phi$}}] {};
+		\draw[gray] (5.75,0) -- (5.75, 0.25) -- (6, 0.25);
+		\end{tikzpicture}
+	}	
+	\caption{A geometric description of asteroids at the L\subscript{4} point undergoing a perturbation of magnitude $z_{0}$ out of the XY plane.
+The fixed equilibrium distance between the asteroids and the Sun is denoted by $a$, while $R$ represents the separation between the asteroids and the Sun as a result of the perturbation}
+	\label{fig:geometryz}
+\end{figure}
+
+As the asteroid is equidistant from the Sun and the planet, the magnitudes of these forces depend on these bodies' mass only, and gravitational force components from both massive bodies are similarly orientated in the Z direction.
+The total gravitational force acting on the asteroid is therefore given by:
+
+\begin{equation}
+F = - \frac{G (M_{s} + M_{p}) M_{a}}{R^{2}}
+\end{equation}
+
+where $M_{s}$, $M_{p}$ and $M_{a}$ represent the solar, planetary and asteroid masses respectively. Resolving this force in Cartesian components, the Z component is given by:
+
+\begin{equation}
+F_{z} = F cos(\phi) = F \frac{z}{R} = - \frac{G (M_{s} + M_{p}) M_{a} z}{R^{3}}
+\end{equation}
+
+In conjunction with Newton's Second Law, we therefore obtain a second order differential equation for $z$:
+
+\begin{equation}
+\ddot{z} + \frac{G (M_{s} + M_{p})}{R^{3}} z = 0
+\end{equation}
+
+which can be treated as simple harmonic motion in the small perturbation limit where $R \approx a$. This has a angular frequency equal to (\ref{frequency}):
+
+\begin{equation}
+\Omega = \sqrt{\frac{G (M_{s} + M_{p})}{a^{3}}}
+\end{equation}
+
 \section{Program Structure} \label{structure}
-Global constants, along with derived values from these are given in an python module "\textit{constants}", from which relevant variable are imported in all other scripts.
+Global constants, along with derived values from these, are given in a python module "\textit{constants}", from which relevant variables are imported in all other scripts.
 
-Functions to evaluate these coupled differential equation systems are defined in module "\textit{orbits}", while additional functions to evaluate the wander during the orbit are implemented in "\textit{wander}".
-These modules are both imported into the scripts below for plotting and analysis, and given in the listings in Appendix \ref{Code}.
+Functions to evaluate the coupled differential equation systems are defined in the module "\textit{orbits}", while additional functions to evaluate the wander during the orbit are implemented in "\textit{wander}".
+These modules are both imported into the scripts detailed in Table \ref{Codefiles} for plotting and analysis, and given in the listings in Appendix \ref{Code}.
 
-\textit{Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
-Orbits in the stationary frame are animated and saved within "\textit{stat\_frame\_video}".
+%Wander from unperturbed initial conditions is calculated within "\textit{lagrange\_stability}", with ODE solver performance also evaluated here.
+%Orbits in the stationary frame are animated and saved within "\textit{stat\_frame\_video}".
 
-2D perturbations are characterised in \textit{perturbation\_2D}, while perturbations in the z direction are considered in \textit{perturbation\_3D}.
-Variation in planetary mass, for results in section \ref{planet}, are given in \textit{planet\_mass\_variation}.}
+%2D perturbations are characterised in \textit{perturbation\_2D}, while perturbations in the Z direction are considered in \textit{perturbation\_3D}.
+Variation in planetary mass, for results in section \ref{planet}, are given in \textit{planet\_mass\_variation}.
 
-Other code files are listed by Section below in Table \ref{Codefiles}:
+% Other code files are listed by Section below in Table :
 
-\begin{table}[ht]
-	\caption{Program files used to generate results in each section, with a description of their role.}
+\newgeometry{
+	top=1cm, % Top margin
+	bottom=1.5cm, % Bottom margin
+	left=1cm, % Left margin
+	right=1cm, % Right margin
+	includehead, % Include space for a header
+	includefoot, % Include space for a footer
+	%showframe, % Uncomment to show how the type block is set on the page
+}
+\onecolumn
+\clearpage
+
+\begin{table}[ht!]
+	\caption{Program files used to analyse and plot data for each section, with a description of their role.}
 	\centering
@@
 		\hline
-		\ref{Fourier} & Fourier Analysis & \textit{lagrange\_stability} \\
+		\ref{Fourier} & Fourier analysis & \textit{lagrange\_stability} \\
 		\hline
@@
 		\hline
-		\ref{perturbed}& Perturbations restricted to the xy plane & \textit{perturbation\_2D} \\
+		\ref{perturbed}& Perturbations restricted to the XY plane & \textit{perturbation\_2D} \\
 		\hline
-		\ref{orbits}& Plotting of Orbit types in xy plane & \textit{perturbation\_2D} \\
+		\ref{orbits}& Plotting of orbit types in XY plane & \textit{perturbation\_2D} \\
 		\hline
-		\ref{3D}& Perturbations in the z-direction & \textit{perturbation\_3D} \\
+		\ref{3D}& Perturbations in the Z direction & \textit{perturbation\_3D} \\
 		\hline
@@
 \section{Code Listings} \label{Code}
+\subsection{\textit{constants.py}}
+\begin{lstlisting}[language=Python]
+import math
+import numpy as np
+
+# USER DEFINED CONSTANTS
+G = 4 * np.pi ** 2  # gravitational constant in astronomical units
+M_S = 1  # solar mass
+M_P = 0.001  # planetary mass as a fraction of solar mass
+R = 5.2  # average planetary radius from sun in AU
+
+PRECISION = 100  # evaluation points per orbit
+ORBIT_NUM = 100  # number of orbits
+
+# DERIVED CONSTANTS
+solar_rad = R * M_P / (M_S + M_P)  # distance from origin to sun in CoM frame
+planet_rad = R * M_S / (M_S + M_P)
+period = math.sqrt(R ** 3 / (M_S + M_P))
+omega = 2 * np.pi / period  # angular velocity of frame
+time_span = np.linspace(0, ORBIT_NUM * period, int(ORBIT_NUM * PRECISION))
+lagrange = (planet_rad - R / 2, R * math.sqrt(3) / 2, 0)
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+\end{lstlisting}
+
+\bigskip
+\subsection{\textit{orbits.py}}
 \begin{lstlisting}[language=Python]
@@
 
+# DERIVED QUANTITIES
+# refers to greeks in this case, symmetric to trojans
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+greek_rad = math.sqrt(R ** 2 - solar_rad * planet_rad)
+greek_v = omega * (greek_rad)
+
+# Initial conditions for Greeks, given as equillibrium state 
+cos, sin = np.cos(greek_theta), np.sin(greek_theta)
+rcos, rsin = greek_rad * cos, greek_rad * sin
+vcos, vsin = greek_v * cos, greek_v * sin
+
+# EXACT SOLUTIONS FOR POSITION
+def solar_pos(t=0):
+	"""Position of sun in 3-dimensions at time t
+
+	Note that z coordinate is fixed to 0, 
+	so movement is constrained to xy plane in this exact solution
+	"""
+	return np.array([-solar_rad * np.cos(omega * t), -solar_rad * np.sin(omega * t), 0])
+
+def planet_pos(t=0):
+	"""Position of planet in 3-dimensions at time t
+
+	Note that z coordinate is fixed to 0, 
+	so movement is constrained to xy plane in this exact solution
+	"""
+	return np.array([planet_rad * np.cos(omega * t), planet_rad * np.sin(omega * t), 0])
+
+def lagrange_pos(t=0):
+	"""Position of Greeks' lagrange point in 3-dimensions at time t
+
+	Note that z coordinate is fixed to 0, 
+	so movement is constrained to xy plane in this exact solution"""
+	return np.array(
+		[
+			greek_rad * np.cos(omega * t + greek_theta),
+			greek_rad * np.sin(omega * t + greek_theta),
+			0,
+		]
+	)
+
+# ODE SOLVER IN THE ROTATING FRAME
+def rot_derivatives(t, y):
+	"""Gives derivative of each term of y at time t in the rotating frame
+
+	y should be of the form (x_pos, y_pos, z_pos, x_vel, y_vel, z_vel)
+	"""
+
+	position, velocity = np.array(y[0:3]), np.array(y[3:6])
+
+	# Factors defined for simplicity in acceleration term:
+	solar_dr3 = np.linalg.norm(solar_pos(0) - position) ** 3
+	planet_dr3 = np.linalg.norm(planet_pos(0) - position) ** 3
+
+	virtual_force = (
+		position[0] * omega ** 2 + 2 * omega * velocity[1],
+		position[1] * omega ** 2 - 2 * omega * velocity[0],
+		0,
+	)
+
+	acceleration = (
+		-G
+		* (
+			M_S * (position[0] - solar_pos(0)[0]) / solar_dr3
+			+ M_P * (position[0] - planet_pos(0)[0]) / planet_dr3
+		)
+		+ virtual_force[0], # x component 
+		-G * (M_S * position[1] / solar_dr3 + M_P * position[1] / planet_dr3)
+		+ virtual_force[1], # y component 
+		-G * (M_S * position[2] / solar_dr3 + M_P * position[2] / planet_dr3)
+		+ virtual_force[2], # z component 
+	)
+
+	return np.concatenate((velocity, acceleration))
+
+def rotating_frame(y0_rot=(rcos, rsin, 0, 0, 0, 0)):
+	"""Gives position and velocity of asteroids in the rotating frame
+
+	Uses scipy solve_ivp method with Radau, taking an input in the form
+	of (x_pos, y_pos, z_pos, x_vel, y_vel, z_vel) for the initial state.
+	This takes a default value of the equillibrium position for y0_rot 
+	if this is not given.
+	"""
+
+	return integrate.solve_ivp(
+		fun=rot_derivatives,
+		method="Radau",  # Or LSODA for non-stiff alternative
+		t_span=(0, ORBIT_NUM * period),
+		y0=y0_rot,
+		t_eval=time_span,  # selects points for storage
+	)
+
+# ODE SOLVER IN THE STATIONARY FRAME
+def stat_acceleration(position, t):
+	"""Gives acceleration in stationary frame at a given position and time"""
+	# factors defined for convenience
+	solar_dr3 = np.linalg.norm(solar_pos(t) - position) ** 3
+	planet_dr3 = np.linalg.norm(planet_pos(t) - position) ** 3
+
+	return (
+		-G
+		* (
+			M_S * (position[0] - solar_pos(t)[0]) / solar_dr3
+			+ M_P * (position[0] - planet_pos(t)[0]) / planet_dr3
+		),
+		-G
+		* (
+			M_S * (position[1] - solar_pos(t)[1]) / solar_dr3
+			+ M_P * (position[1] - planet_pos(t)[1]) / planet_dr3
+		),
+		-G * (M_S * position[2] / solar_dr3 + M_P * position[2] / planet_dr3),
+	)
+
+def stat_derivatives(t, y):
+	"""Gives derivative of each term of y at time t in the stationary frame
+
+	y should be of the form (x_pos, y_pos, z_pos, x_vel, y_vel, z_vel)
+	"""
+	position, velocity = np.array(y[0:3]), np.array(y[3:6])
+	return np.concatenate((velocity, stat_acceleration(position, t)))
+
+def stationary_frame(y0_stat=(rcos, rsin, 0, -vsin, vcos, 0)):
+	"""Gives position and velocity of asteroids in the stationary frame
+
+	Uses scipy solve_ivp method with LSODA, taking an input in the form
+	of (x_pos, y_pos, z_pos, x_vel, y_vel, z_vel) for the initial state.
+	This has a default value of the equillibrium state if not given.
+	"""
+	return integrate.solve_ivp(
+		fun=stat_derivatives,
+		method="Radau",  # Or LSODA for non-stiff alternative
+		t_span=(0, ORBIT_NUM * period),
+		y0=y0_stat,
+		t_eval=time_span,
+	)
+
+def specific_energy(t, y):
+	""" Gives energy per unit mass of asteroids
+
+	t is the time after the start of the orbit
+	y has six components; first three for position, second three for velocity
+	These are in the form of the output of a scipy integrate function
+	"""
+
+	radius_p = np.linalg.norm(y[0:3] - planet_pos(t), axis=0)
+	radius_s = np.linalg.norm(y[0:3] - solar_pos(t), axis=0)
+	velocity = np.linalg.norm(y[3:], axis=0)
+	return -G * M_P / radius_p - G * M_S / radius_s + 1 / 2 * velocity ** 2
+
 \end{lstlisting}
-%Can add my own style guide to this, see suggestions on https://www.overleaf.com/learn/latex/code_listing
-%\section{Code 2}
+\bigskip
+
+\subsection{\textit{wander.py}}
+\begin{lstlisting}[language=Python]
+import numpy as np
+
+import orbits
+from constants import omega, lagrange, greek_theta  # Derived constants
+
+# INITIAL CONDITIONS
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+initial_cond_stat = np.array(
+	(lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+
+# DEFINED FUNCTIONS
+def perturb(initial_cond, max_pertubation_size=0.01):
+	""" Returns perturbed version of initial conditions array
+
+	init_cond is the initial conditions, submitted as a numpy array
+	pertubation_size is the relative magnitude of the perturbation, measured as 
+	the percentage deivation from the lagrange point (for position)
+
+	Perturbations are only chosen randomly from within a circle of radius one, 
+	to give a more uniform radial distribution than from a square.
+	Note that this function will give a different value each time it is run
+	"""
+	rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+	while np.linalg.norm(rand_array[0:2]) > 1: # reject points outside circle
+		rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+	return initial_cond + initial_cond * (
+		max_pertubation_size * rand_array	
+	)  # random between -1 and 1, selects points in circle of radius 1
+
+def rand_sample(max_pertubation_size, samples):
+	"""Returns maximum deviation over orbit in the 
+	rotating frame for given number of random samples
+
+	This calculates the distange from the initial point (not the lagrange point)
+	It returns a samples*3 array, giving the radial and tangential components
+	of the pertubation, and then the maximum wander over this timespan
+
+	pertubation_size determies the relative magnitude of the perturbation
+	Samples denotes the number of random pertubations sampled at the given size
+	"""
+	output = np.zeros((samples, 3))  # size of pertubation; size of wander
+	for n in range(samples):
+		initial_cond = perturb(initial_cond_rot, max_pertubation_size)
+		orbit = orbits.rotating_frame(initial_cond)
+		sample_wander = np.zeros((len(orbit.t)))
+
+		perturb_theta = np.arctan(
+			(initial_cond[1] - initial_cond_rot[1])
+			/ (initial_cond[0] - initial_cond_rot[0])
+		)
+
+		for i in range(len(sample_wander)):
+			sample_wander[i] = np.linalg.norm(
+				orbit.y[0:3, i] - initial_cond[0:3]
+			)  # deviation in pos only
+		output[n, 0] = np.linalg.norm((initial_cond[0:3] - lagrange[0:3])) * np.abs(
+			np.cos(greek_theta - perturb_theta)
+		)  # radial component of pertubation
+		output[n, 1] = np.linalg.norm((initial_cond[0:3] - lagrange[0:3])) * np.abs(
+			np.sin(greek_theta - perturb_theta)
+		)  # tangential component of pertubation
+		output[n, 2] = np.max(sample_wander)  # overall magnitude of wander
+	return output
+
+def initial_point(pertubation, pertubation_type="position"):
+	"""Maximum wander over orbit in the rotating frame for given initial point
+
+	Wander is the maximum deviation from the initial point 
+	(not the lagrange point) over this timespan
+	perturbation is the initial point in 2D position/velocity space
+	perturbation_type can be "position" or "velocity" 
+	"""
+	if pertubation_type == "position":
+		initial_cond = initial_cond_rot + np.array(
+			(pertubation[0], pertubation[1], pertubation[2], 0, 0, 0)
+		)  # add position pertubation
+	elif pertubation_type == "velocity":
+		initial_cond = initial_cond_rot + np.array(
+			(0, 0, 0, pertubation[0], pertubation[1], pertubation[2])
+		)  # add velocity pertubation
+	else:
+		raise ValueError(
+			"Unknown perturbation_type: Please use 'position' or 'velocity'"
+		)
+
+	orbit = orbits.rotating_frame(initial_cond)
+	wander_t = np.zeros((len(orbit.t)))
+	for i in range(len(orbit.t)):
+		wander_t[i] = np.linalg.norm(
+			orbit.y[0:3, i]
+			- initial_cond[0:3]
+			# - lagrange[0:3]
+		)  # deviation in pos only
+	return np.max(wander_t)
+
+\end{lstlisting}
 
@@
 
+\restoregeometry
 \end{document}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Can add my own style guide to listings, see suggestions on https://www.overleaf.com/learn/latex/code_listing

### 2020-04-26 constants.py

@@ -4,12 +4,12 @@ import numpy as np
 # USER DEFINED CONSTANTS
 G = 4 * np.pi ** 2  # gravitational constant in astronomical units
 M_S = 1  # solar mass
-M_P = 0.001  # planetary mass
-R = 5.2  # average planetary radius from sun
+M_P = 0.001  # planetary mass as a fraction of solar mass
+R = 5.2  # average planetary radius from sun in AU
 
 
 PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 30  # number of orbits
+ORBIT_NUM = 100  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-26 orbits.py

@@ -18,16 +18,12 @@ greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 greek_rad = math.sqrt(R ** 2 - solar_rad * planet_rad)
 greek_v = omega * (greek_rad)
 
-lagrange = (planet_rad - R / 2, R * math.sqrt(3) / 2, 0)
-# are these necessary as well as initial conditions?
-# decide later which are most useful to import etc
-
 # Initial conditions for Greeks, given as equillibrium state but can be changed in XXXXX
 cos, sin = np.cos(greek_theta), np.sin(greek_theta)
 rcos, rsin = greek_rad * cos, greek_rad * sin
 vcos, vsin = greek_v * cos, greek_v * sin
 
-# Positions of massize bodies
+# EXACT SOLUTIONS FOR POSITION
 def solar_pos(t=0):
     """Position of sun in 3-dimensions at time t
 
@@ -59,9 +55,7 @@ def lagrange_pos(t=0):
     )
 
 
-# assuming massive bodies do not deviate from xy plane
-
-
+# ODE SOLVER IN THE ROTATING FRAME
 def rot_derivatives(t, y):
     """Gives derivative of each term of y at time t in the rotating frame
 
@@ -70,7 +64,8 @@ def rot_derivatives(t, y):
 
     position, velocity = np.array(y[0:3]), np.array(y[3:6])
 
-    solar_dr3 = np.linalg.norm(solar_pos(0) - position) ** 3  # should this be zero?
+    # Factors defined for simplicity in acceleration term:
+    solar_dr3 = np.linalg.norm(solar_pos(0) - position) ** 3
     planet_dr3 = np.linalg.norm(planet_pos(0) - position) ** 3
 
     virtual_force = (
@@ -85,22 +80,22 @@ def rot_derivatives(t, y):
             M_S * (position[0] - solar_pos(0)[0]) / solar_dr3
             + M_P * (position[0] - planet_pos(0)[0]) / planet_dr3
         )
-        + virtual_force[0],
+        + virtual_force[0],  # x component
         -G * (M_S * position[1] / solar_dr3 + M_P * position[1] / planet_dr3)
-        + virtual_force[1],
+        + virtual_force[1],  # y component
         -G * (M_S * position[2] / solar_dr3 + M_P * position[2] / planet_dr3)
-        + virtual_force[2],
+        + virtual_force[2],  # z component
     )
[…]

### 2020-04-26 perturbations_2D.py

@@ -75,7 +75,7 @@ plt.show()
 # # # START OF MESH GRID PLOTS
 
 grid_size = 0.04
-sampling_points = 2
+sampling_points = 5
 # make these smaller to increase the resolution
 dx, dy = grid_size / sampling_points, grid_size / sampling_points
 

### 2020-04-26 planet_mass_variation.py

@@ -11,7 +11,7 @@ greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 cos = math.cos(greek_theta)
 sin = math.sin(greek_theta)
 
-mass = np.linspace(0.0005, 0.0425, 100)  # range of planetary masses
+mass = np.linspace(0.0005, 0.0425, 10)  # range of planetary masses
 max_wander = np.zeros_like(mass)
 for n in range(len(mass)):
     import constants
@@ -49,5 +49,5 @@ plt.plot(mass, max_wander, marker="x", linestyle="")
 plt.title("Variation of Wander with Planetary Mass")
 plt.xlabel("Planet Mass /Solar Masses")
 plt.ylabel("Wander /AU")
-plt.savefig("wanderwithplanetmass_p6.png")
+plt.savefig("wanderwithplanetmass_p7.png")
 plt.show()

### 2020-04-26 wander.py

@@ -3,24 +3,23 @@ import numpy as np
 import orbits
 from constants import omega, lagrange, greek_theta  # Derived constants
 
-# Initial Conditions
+# INITIAL CONDITIONS
 initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
 initial_cond_stat = np.array(
     (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
 )
 
-
-# Defined Functions
+# DEFINED FUNCTIONS
 def perturb(initial_cond, max_pertubation_size=0.01):
     """ Returns perturbed version of initial conditions array
 
     init_cond is the initial conditions, submitted as a numpy array
     pertubation_size is the relative magnitude of the perturbation, measured as 
     the percentage deivation from the lagrange point (for position)
-    Note that this will give a different value each time it is run
 
     Perturbations are chosen randomly within a circle of radius one, 
     to give a more uniform radial distribution than from a square.
+    Note that this will give a different value each time it is run
     """
     rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
     while np.linalg.norm(rand_array[0:2]) > 1:
@@ -53,9 +52,7 @@ def rand_sample(max_pertubation_size, samples):
 
         for i in range(len(sample_wander)):
             sample_wander[i] = np.linalg.norm(
-                orbit.y[0:3, i]
-                - initial_cond[0:3]
-                # - lagrange[0:3]
+                orbit.y[0:3, i] - initial_cond[0:3]
             )  # deviation in pos only
         output[n, 0] = np.linalg.norm((initial_cond[0:3] - lagrange[0:3])) * np.abs(
             np.cos(greek_theta - perturb_theta)
@@ -69,7 +66,6 @@ def rand_sample(max_pertubation_size, samples):
 
 def initial_point(pertubation, pertubation_type="position"):
     """Returns maximum wander over orbit in the rotating frame for given initial point 
-
     
     Wander is the maximum deviation from the initial point (not the lagrange point) over this timespan
     perturbation is the initial point in 2D position/velocity space

### 2020-04-27 Report/draft1.tex
adapted planet_mass_variation

@@
 	\centering
+	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_q7}
+	\caption{Variation of maximum wander with planetary mass, under a 0.001 AU radial perturbation and 0.001 AU/year tangential boost. Note the tendency }
+	\label{fig:planetmass}
+\end{figure}
+
+For small planetary masses (up to approximately $ M_{p} = 0.01 M_{s}$), the wander follows a $ M_{P}^{-\frac{1}{2}}$ dependence, tending towards the initial perturbation for larger masses.
+However, wander increases unbounded and orbits become unstable beyond $ M_{p} = 0.04 M_{s}$, as depicted in Figure \ref{fig:planetmass2}, in agreement with theoretical predictions by Darwin \cite{Darwin1897}.
+
+\begin{figure}
+	\centering
 	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p6}
@@
 Note the orbit stability up to $ M_{p} = 0.04 M_{s}$, and instability beyond this.}
-	\label{fig:planetmass}
+	\label{fig:planetmass2}
 \end{figure}
-
-An approximate quadratic trend is initially observed in the small mass region, with minimum wander occurring at $ M_{p} = 0.026 M_{s}$.
-However, the variation of wander below $ M_{p} = 0.04 M_{s}$ is minimal, while orbits are unstable beyond this point, in agreement with theoretical predictions by Darwin \cite{Darwin1897}.
 
@@
 It was also possible to replicate both tadpole and horseshoe orbits from radial perturbations, and frequency components were obtained for these orbits in agreement with previous literature.
-It was also found that orbits become unstable when the planetary\textendash solar mass ratio is greater than 0.04, in agreement with theoretical predictions, and there is little mass dependence below this point.
+It was also found that orbits become unstable when the planetary\textendash solar mass ratio is greater than 0.04, in agreement with theoretical predictions, and inverse root mass dependence below this point.
 

### 2020-04-27 planet_mass_variation.py

@@ -1,4 +1,5 @@
 import math
+import matplotlib
 import matplotlib.pyplot as plt
 import numpy as np
 
@@ -11,7 +12,7 @@ greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
 cos = math.cos(greek_theta)
 sin = math.sin(greek_theta)
 
-mass = np.linspace(0.0005, 0.0425, 10)  # range of planetary masses
+mass = np.linspace(1.0e-05, 1e-3, 150)  # range of planetary masses
 max_wander = np.zeros_like(mass)
 for n in range(len(mass)):
     import constants
@@ -34,7 +35,7 @@ for n in range(len(mass)):
 
     initial_cond = np.array((constants.lagrange[0], constants.lagrange[1], 0, 0, 0, 0))
     orbit = orbits.rotating_frame(
-        (initial_cond + 0.001 * np.array((cos, sin, 0, sin, -cos, 0)))
+        (initial_cond + 1e-3 * np.array((cos, sin, 0, sin, -cos, 0)))
     )  # radial perturbation
     wander_t = np.zeros((len(orbit.t)))
     for i in range(len(orbit.t)):
@@ -44,10 +45,32 @@ for n in range(len(mass)):
     max_wander[n] = np.max(wander_t)
     print(str(n) + ": " + str(orbits.M_P))
 
-plt.plot(mass, max_wander, marker="x", linestyle="")
 
+fig = plt.figure()
+ax = fig.add_subplot()
+
+coeff = np.polyfit(np.reciprocal(np.sqrt(mass)), max_wander, 1)  # for line of best fit
+order = np.argsort(max_wander)
+
+plt.plot(mass, max_wander, linestyle="None", marker="x", color="navy")
+plt.plot(
+    mass[order],
+    coeff[0] * np.reciprocal(np.sqrt(mass[order])) + coeff[1],
+    linewidth=0.75,
+)
+plt.ticklabel_format(axis="x", style="sci", scilimits=(0, 0))
+plt.text(
+    0.465,
+    0.95,
+    "Line of Best Fit: "
+    + str("{0:.2e}".format(coeff[0]))
+    + "/\u221AM"
+    + str("{0:+.2e}".format(coeff[1])),
+    transform=ax.transAxes,
+    fontsize=9.5,
+)
 plt.title("Variation of Wander with Planetary Mass")
[…]

### 2020-04-27 Report/draft1.tex
Final submission

@@
 	%\authorstyle{Christopher Gallagher}}
-\author{\authorstyle{Christopher Gallagher} %Candidate 6952R}
+\author{\authorstyle{Candidate 6952R}
 	\institution{University of Cambridge}}
@@
 Such orbits were observed to be stable up to a planetary mass of 0.04 \pmb{$M_{\odot}$}, also in agreement with theoretical predictions.
-Word Count: XXXX}
+Word Count: 2782}
 
@@
 \section{Methodology}
-This system of coupled first-order ordinary differential equations (ODEs) was solved using the scipy \textit{solve\_ivp} function.
+This system of coupled first-order ordinary differential equations (ODEs) was solved using the \textit{solve\_ivp} function from \textit{scipy}.
 The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years.
@@
 \subsection{Wander Analysis} \label{perturbed}
-Wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin.
-The perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively) were considered separately.
-By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).
 
-\begin{figure}[ht]
+\begin{figure}[hb!]
 	\centering
@@
 \end{figure}
+
+Wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin.
+The perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively) were considered separately.
+By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 	\centering

### 2020-04-27 Submission/code/constants.py

@@ -9,7 +9,7 @@ R = 5.2  # average planetary radius from sun in AU
 
 
 PRECISION = 100  # evaluation points per orbit
-ORBIT_NUM = 10  # number of orbits
+ORBIT_NUM = 100  # number of orbits
 
 
 # DERIVED CONSTANTS

### 2020-04-27 Submission/code/orbits.py

@@ -1,0 +1,168 @@
+import math
+import numpy as np
+from scipy import integrate
+import matplotlib.pyplot as plt
+
+from constants import G, M_S, M_P, R, ORBIT_NUM, PRECISION  # User defined constants
+from constants import (
+    solar_rad,
+    planet_rad,
+    period,
+    omega,
+    time_span,
+)  # Derived constants
+
+# DERIVED QUANTITIES
+# actually refers to greeks in this case, symmetric to trojans
+greek_theta = np.arctan((R * math.sqrt(3) / 2) / (R / 2 - solar_rad))
+greek_rad = math.sqrt(R ** 2 - solar_rad * planet_rad)
+greek_v = omega * (greek_rad)
+
+# Initial conditions for Greeks, given as equillibrium state but can be changed in XXXXX
+cos, sin = np.cos(greek_theta), np.sin(greek_theta)
+rcos, rsin = greek_rad * cos, greek_rad * sin
+vcos, vsin = greek_v * cos, greek_v * sin
+
+# EXACT SOLUTIONS FOR POSITION
+def solar_pos(t=0):
+    """Position of sun in 3-dimensions at time t
+
+    Note that z coordinate is fixed to 0, 
+    so movement is constrained to xy plane in this exact solution
+    """
+    return np.array([-solar_rad * np.cos(omega * t), -solar_rad * np.sin(omega * t), 0])
+
+
+def planet_pos(t=0):
+    """Position of planet in 3-dimensions at time t
+
+    Note that z coordinate is fixed to 0, 
+    so movement is constrained to xy plane in this exact solution"""
+    return np.array([planet_rad * np.cos(omega * t), planet_rad * np.sin(omega * t), 0])
+
+
+def lagrange_pos(t=0):
+    """Position of Greeks' lagrange point in 3-dimensions at time t
+
+    Note that z coordinate is fixed to 0, 
+    so movement is constrained to xy plane in this exact solution"""
+    return np.array(
+        [
+            greek_rad * np.cos(omega * t + greek_theta),
+            greek_rad * np.sin(omega * t + greek_theta),
+            0,
+        ]
+    )
+
+
+# ODE SOLVER IN THE ROTATING FRAME
+def rot_derivatives(t, y):
+    """Gives derivative of each term of y at time t in the rotating frame
+
+    y should be of the form (x_pos, y_pos, z_pos, x_vel, y_vel, z_vel)
+    """
+
[…]

### 2020-04-27 Submission/code/wander.py

@@ -1,0 +1,95 @@
+import numpy as np
+
+import orbits
+from constants import omega, lagrange, greek_theta  # Derived constants
+
+# INITIAL CONDITIONS
+initial_cond_rot = np.array((lagrange[0], lagrange[1], 0, 0, 0, 0))  # in rotating frame
+initial_cond_stat = np.array(
+    (lagrange[0], lagrange[1], 0, -omega * lagrange[1], omega * lagrange[0], 0)
+)
+
+# DEFINED FUNCTIONS
+def perturb(initial_cond, max_pertubation_size=0.01):
+    """ Returns perturbed version of initial conditions array
+
+    init_cond is the initial conditions, submitted as a numpy array
+    pertubation_size is the relative magnitude of the perturbation, measured as 
+    the percentage deivation from the lagrange point (for position)
+
+    Perturbations are chosen randomly within a circle of radius one, 
+    to give a more uniform radial distribution than from a square.
+    Note that this will give a different value each time it is run
+    """
+    rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+    while np.linalg.norm(rand_array[0:2]) > 1:
+        rand_array = np.random.uniform(-1, 1, (np.shape(initial_cond)))
+    return initial_cond + initial_cond * (
+        max_pertubation_size * rand_array
+    )  # random between -1 and 1, selects points in circle of radius 1
+
+
+def rand_sample(max_pertubation_size, samples):
+    """Returns maximum deviation over orbit in the rotating frame for given number of random samples
+
+    This calculates the distange from the initial point (not the lagrange point)
+    It returns a samples*3 array, giving the radial and tangential components
+    of the pertubation, and then the maximum wander over this timespan
+
+    pertubation_size determies the relative magnitude of the perturbation as in perturb()
+    Samples denotes the number of random pertubations sampled for given perturbation size
+    """
+    output = np.zeros((samples, 3))  # size of pertubation; size of wander
+    for n in range(samples):
+        initial_cond = perturb(initial_cond_rot, max_pertubation_size)
+        orbit = orbits.rotating_frame(initial_cond)
+        sample_wander = np.zeros((len(orbit.t)))
+
+        perturb_theta = np.arctan(
+            (initial_cond[1] - initial_cond_rot[1])
+            / (initial_cond[0] - initial_cond_rot[0])
+        )
+
+        for i in range(len(sample_wander)):
[…]
